---
output:
  html_document: default
  pdf_document: default
  word_document: default
  html_notebook: default
---
---
title: "Statistical re-analysis plan: Predicting body fat proportion using anthropometric measures"
author: "Mariana Nold" 
---

########  Run Task Bodyfat simplified first.

##### Load R packages
```{r load package, message = FALSE, warning = FALSE}
library("rstanarm")
library("projpred")
library("ggplot2")
library("loo")
library("bayesplot")
library("bayesrules")
library("DHARMa")
library("ggplot2")
library("dplyr")
library("tidyr")
```


```{r set path and load data (female and male)}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task Bodyfat/"
load("~/GitHub/Bayes_for_STRATOS/Task Bodyfat/Datafiles/bodyfatdata.Rdata")
#scale data sets
# Create copies of the original datasets
#woman <- women
#man <- men

# Standardize all variables except the first two in 'woman'
#woman[, -c(1,2)] <- scale(woman[, -c(1,2)])

# Standardize all variables except the first three in 'man'
#man[, -c(1,2,3)] <- scale(man[, -c(1,2,3)])

# Display summary statistics for validation
summary(women)
sd(women$Fat)
summary(men)
sd(men$Fat)
```

```{r fit full model with rstanarm, results='hide', cache=TRUE}

rmf <- as.formula(Fat ~ Waist + Height + Weight + Age + Neck + Chest +  
                          Biceps + Hips + Forearm 
                        +  Knee +  Ankle + BMI)

# Fit a Bayesian Gaussian regression model using stan_glm 
ref_fit <- stan_glm(rmf, data = women, family = gaussian())
# Check reference model, MCMC check
summary(ref_fit)
#shinystan::launch_shinystan(ref_fitf)

mcmc_areas(as.matrix(ref_fit)[,2:13])
```

##### Investigate quality of the reference model
```{r LOOPIT and bayesrules summary, cache=TRUE}

pp_check(ref_fit)
y_rep_rmf <- posterior_predict(ref_fit)
loo_rmf <- loo(ref_fit , save_psis = TRUE, cores = 4)
psis1_rmf <- loo_rmf$psis_object
lw_rmf <- weights(psis1_rmf)
color_scheme_set("red")
#pdf(file.path(PPath, "/LOOPITrmf.pdf"))
ppc_loo_pit_overlay(y= women$Fat,yrep=y_rep_rmf,lw=lw_rmf)
#dev.off()

# Summarize the posterior distributions simple model
summary_rmf_cv <- prediction_summary_cv(model = ref_fit , data = women, k = 3)

# Print the summary simple model
print(summary_rmf_cv)

```
##### Use projpred to downsize model even more
```{r predictive projection, results='hide', cache=TRUE}
set.seed(2341)
vs_cvf <- cv_varsel(ref_fit)

save(vs_cvf,file = file.path(PPath,"vs_cvf"))
#load(file = file.path(PPath,"vs_cvf"))

# Plot the variable selection 
#pdf(file.path(PPath, "cv_valsel_plotf.pdf"))
plot(vs_cvf, deltas = TRUE)
#dev.off()
# 

```

```{r best five predictors}
nsel <- suggest_size(vs_cvf)
nsel

rank_vs_cvf <- ranking(vs_cvf)
rank_vs_cvf$fulldata

solterms_final_vs_cvf <- head(rank_vs_cvf$fulldata ,suggest_size(vs_cvf))

solterms_final_vs_cvf

# Print the summary of the variable selection 
summary(vs_cvf) 
solterms_final_fivef <- head(rank_vs_cvf$fulldata, 5)
solterms_final_fivef 
```

##### Projected posterior for the selected model
```{r post-selection inference, results='hide'}
# https://cran.r-project.org/web/packages/projpred/vignettes/projpred.html#post-selection-inference
proj_ref_fit <- project(vs_cvf, predictor_terms = solterms_final_fivef, ns = 4000)

```
##### Show results
```{r Show results}
color_scheme_set(scheme = "green")
mcmc_areas(as.matrix(proj_ref_fit), pars = solterms_final_fivef)# 
```
#### Robustness check based on Bayesian LASSO and compuation of prob. unequal zero
```{r compute Bayesian LASSO, results='hide'}
# Bayesian Lasso using rstanarm
lasso <- stan_glm(Fat ~ Waist + Height + Weight + Age + Neck + Chest +  
                    Biceps + Hips + Forearm + Knee + Ankle + BMI, 
                  data = women, family = gaussian(),
                  prior =  lasso(df = 1, location = 0, scale = NULL, autoscale = TRUE), prior_intercept = normal())
```

```{r robustness check based on Bayesian LASSO and compuation of prob. unequal zero}
prior_summary(lasso)
# show results
summary(lasso)

color_scheme_set(scheme = "pink")
mcmc_areas(as.matrix(lasso))


mcmc_areas(as.matrix(lasso), pars = setdiff(names(lasso$coefficients), "(Intercept)"))
pred <- setdiff(names(lasso$coefficients), "(Intercept)")
posterior_samples <- as.matrix(lasso, pars = pred)

# Define the range
lower_bound <- -0.05
upper_bound <- 0.05

# Apply the condition to each element of the matrix
prob_between <- apply(posterior_samples, 2, function(par) mean(par >= lower_bound & par <= upper_bound))

# Print the result
print(prob_between)

sort(prob_between)
sort(1 - prob_between)
```
##### The final model, which includes five predictors, is  identical
```{r define formula}
formula_male <- as.formula(Fat ~ Waist  + Hips + BMI + Knee + Ankle)
```

###### Model diagnostics (1.3.3)
```{r Model diagnostics}
y_linpred_female <- proj_linpred(proj_ref_fit, solution_terms= solterms_final_fivef, newdata = women)
women2 <- cbind(women, linpred = colMeans(y_linpred_female$pred))
plotLocBias <- ggplot(data.frame(fitted.values = women2$linpred, y = women2$Fat), aes(x=fitted.values, y=y)) + geom_point() + geom_smooth() + geom_abline(slope=1, intercept=0)
plotLocBias
# Compute residuals
prj_predictW <- proj_predict(proj_ref_fit)

qres.m1=createDHARMa(simulatedResponse = t(prj_predictW ), 
                          observedResponse = women2$Fat, 
                          fittedPredictedResponse = apply(prj_predictW , 2, median), integerResponse = F, seed = 123, method = "PIT") 
summary(residuals(qres.m1))
# QQ-Plot 
res <- qnorm(residuals(qres.m1))

# QQ-Plot mit ggplot2
res.plot <- ggplot(data.frame(res), aes(sample = res)) +
  stat_qq(color = "steelblue", size = 2) +
  stat_qq_line(color = "firebrick", linetype = "dashed") +
  labs(
     title = "QQ Plot of Transformed Residuals",
    x = "Theoretical Quantiles",
    y = "Observed Quantiles"
  ) +
  theme_minimal(base_size = 14)
res.plot
# Check heteroskedasticity (1.3.3.2)
plotHet <- ggplot(data.frame(fitted.values = women2$linpred, y=sqrt(abs(res))), aes(x=fitted.values, y=y)) + geom_point() + geom_smooth() 
plotHet 
# Check association with excluded predictors
plotExcPred <- ggplot(data.frame(Weight=women2$Weight, Residuals=res), aes(x=Weight, y=Residuals)) + geom_point() + geom_smooth() 
plotExcPred

ggplot(data=data.frame(residuals=res), aes(x=res)) + geom_histogram(bins=20)
# Apparent performance of the model (1.3.4)
R.square <- cor(women2$linpred, women2$Fat, method = "pearson")
R.square
color_scheme_set(scheme = "purple")
mcmc_areas(as.matrix(proj_ref_fit), pars = "sigma")
summary(as.matrix(proj_ref_fit)[, "sigma"])
```
##### Internal validation
##### Calibration intercept and slope (internal validation)
```{r compute Calibration intercept and slope (internal), results='hide'}
observed <- women2$Fat
predicted <- women2$linpred
calib_dataW <- data.frame(observed = observed, predicted = predicted)
# Compute calibration intercept and slope
calibration_modelW <- stan_glm(observed ~ predicted, family = gaussian(), data = calib_dataW)
```

```{r Calibration intercept and slope (internal)}


summary(calibration_modelW)

posterior_samplesW <- as.matrix(calibration_modelW)
intercepts <- posterior_samplesW[, "(Intercept)"]
slopes <- posterior_samplesW[, "predicted"]

# Dataframe  for regression lines
regression_linesW <- data.frame(
  intercept = intercepts,
  slope = slopes
)

n_lines <- nrow(as.matrix(calibration_modelW))
print(n_lines)
sampled_lines <- regression_linesW[sample(nrow(regression_linesW), size = 100), ]
# plot
ggplot(calib_dataW, aes(x = predicted, y = observed)) +
  geom_point(alpha = 0.9) +
  geom_abline(data = sampled_lines, aes(intercept = intercept, slope = slope), alpha = 0.1, color = "darkgreen") +
  labs(title = "Calibration Plot with regression lines (internal)",
       x = "Predicted",
       y = "Observed") +
  theme_minimal()
color_scheme_set("gray")
summary(as.matrix(calibration_modelW))

mcmc_areas(as.matrix(calibration_modelW),
           pars = c("(Intercept)", "predicted"),
           prob = 0.95) + 
  ggtitle("Calibration Intercept & Calibration slope (internal)")

library(bayesplot)

# Intercept
mcmc_areas(
  as.matrix(calibration_modelW),
  pars = "(Intercept)",
  prob = 0.95
) +
  ggtitle("Calibration Intercept (internal)")

# Slope
mcmc_areas(
  as.matrix(calibration_modelW),
  pars = "predicted",
  prob = 0.95
) +
  ggtitle("Calibration Slope (internal)")



mcmc_areas(as.matrix(calibration_modelW),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model (internal)")
```
##### External validation
```{r predictions for males based on female model}
y_linpred_male <- proj_linpred(proj_ref_fit, solution_terms= solterms_final_fivef, newdata = men)


color_scheme_set("blue")
prj_predictM <- proj_predict(proj_ref_fit, newdata = men)
# Using the 'bayesplot' package:
ppc_dens_overlay(y = men$Fat, yrep = prj_predictM)

# 3.1
eval_men <- cbind(men, linpred = colMeans(y_linpred_male$pred))
# performance measures at external validation
rmspe_ext <- sqrt(mean((eval_men$Fat - eval_men$linpred)^2))
rmspe_ext
Rsqu_ext <- 1 - (mean((eval_men$Fat - eval_men$linpred)^2))/var(eval_men$Fat)
Rsqu_ext
R.squareM <- cor(eval_men$linpred, eval_men$Fat, method = "pearson")
R.squareM

calPlot <- ggplot(eval_men, aes(linpred, Fat)) + geom_point() + geom_smooth() + geom_smooth(method="lm", col="red") + geom_abline(intercept=0, slope=1) 
calPlot
calResPlot <- ggplot(eval_men, aes(x=linpred, y=Fat-linpred)) + geom_point() + geom_smooth()
calResPlot
hetPlot <- ggplot(eval_men, aes(x=linpred, y=sqrt(abs(Fat-linpred)))) + geom_point() + geom_smooth()
hetPlot
# Predictors: "BMI"   "Hips"  "Waist" "Ankle" "Knee"
nonpredictors <- c("Height", "Weight", "Age", "Neck", "Chest", "Biceps",  "Forearm") 

eval_men_long <- eval_men %>% pivot_longer(cols = all_of(nonpredictors), names_to = "NonPredictor", values_to = "Value") 

plotAll <- ggplot(eval_men_long, aes(x = Value, y = Fat - linpred)) + geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE, color = "blue") + facet_wrap(~ NonPredictor, scales = "free_x") + labs(title = "Non-Predictors versue residuals (Fat - linpred)", x = "Non-Predictor", y = "Residuals (Fat - linpred)") + theme_minimal() 

plotAll
```



##### Calibration intercept and slope (external validation)
```{r compute Calibration intercept and slope (external), results='hide'}
observed <- eval_men$Fat
predicted <- eval_men$linpred
calib_data <- data.frame(observed = observed, predicted = predicted)
# Compute calibration intercept and slope
calibration_model <- stan_glm(observed ~ predicted, family = gaussian(), data = calib_data)
```

```{r Calibration intercept and slope (external)}
summary(calibration_model)

posterior_samples <- as.matrix(calibration_model)
intercepts <- posterior_samples[, "(Intercept)"]
slopes <- posterior_samples[, "predicted"]

# Dataframe  for regression lines
regression_lines <- data.frame(
  intercept = intercepts,
  slope = slopes
)

n_lines <- nrow(as.matrix(calibration_model))
print(n_lines)
sampled_lines <- regression_lines[sample(nrow(regression_lines), size = 100), ]
# plot
ggplot(calib_data, aes(x = predicted, y = observed)) +
  geom_point(alpha = 0.9) +
  geom_abline(data = sampled_lines, aes(intercept = intercept, slope = slope), alpha = 0.1, color = "blue") +
  labs(title = "Calibration Plot with regression lines",
       x = "Predicted",
       y = "Observed") +
  theme_minimal()
color_scheme_set("orange")
summary(as.matrix(calibration_model))


mcmc_areas(as.matrix(calibration_model),
           pars = c("(Intercept)", "predicted"),
           prob = 0.95) + 
  ggtitle("Calibration Intercept & Calibration slope (external)")

library(bayesplot)

# Intercept
mcmc_areas(
  as.matrix(calibration_model),
  pars = "(Intercept)",
  prob = 0.95
) +
  ggtitle("Calibration Intercept (external)")

# Slope
mcmc_areas(
  as.matrix(calibration_model),
  pars = "predicted",
  prob = 0.95
) +
  ggtitle("Calibration Slope (external)")



mcmc_areas(as.matrix(calibration_model),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model (external)")
```







