---
output:
  pdf_document: default
  html_document: default
  word_document: default
  html_notebook: default
---
---
title: "Statistical re-analysis plan: Predicting body fat proportion using anthropometric measures"
author: "Mariana Nold" 
---


##### Load R packages
```{r load package, message = FALSE, warning = FALSE}

library(mfp2)
library(rstanarm)
library(projpred)
library(ggplot2)
library(loo)
library(bayesplot)
library(bayesrules)
library(FBMS)
```


```{r read path and load data}
women <- read.delim("C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task Bodyfat/Datafiles/BodyFat.txt", header = TRUE)
names(women)
PPath2 <- "C:/Users/zo95yup/Nextcloud/Statistik_SoSe24/Forschung/STRATOS/P6/Code_A/Results"

```

##### Prepare data
```{r prepare data, message = FALSE, warning = FALSE}
fat_density <- density(women$Fat, na.rm = TRUE)
ggplot(women, aes(x = Fat)) + geom_density() + 
  labs(title = "Kernel-density plot for outcome fat", x = "Fat", y = "Density")
summary(women$Fat)
summary(women$Age)
women<- women[women$Age != 1, ]
women[,3:20] <- scale(women[,3:20])
head(women)
```

##### Derive reference model

```{r derive reference model based on original SAP}
model1_f <- formula(Fat ~ fp(Waist, df = 2) + fp(Height, df = 2) + fp(Weight, df = 2))
model1 <- mfp2(model1_f, data = women, verbose=TRUE)

# Summarize the model
summary(model1)
print(model1)
# Estimate the model
model2 <- mfp2(Fat ~ fp(Waist, df = 2) + fp(Height, df = 2) + fp(Weight, df = 2) 
               + fp(BMI, df = 1), data = women, verbose=TRUE)

# Summarize the model
summary(model2)
print(model2)
#library(rstanarm)

```


```{r refit model with rstanarm, results='hide' }
# Use the full model after stept 3 in Georgs task, thus model 1 and all predictors but not BMI
# Only use predictors that are available in the man dataset
# Fat = siri, Age = age, Height = height, Weight = weight, Neck = neck, Chest = chest, Waist = abdomen, Hips = hip, Knee = knee, Ankle = ankle, Forearm = forearm, Biceps = biceps
rmf <- as.formula(Fat ~ Waist + Height + Weight + Age + Neck + Chest +  
                          Biceps + Hips + Forearm +
                        Knee +  Ankle)

# Fit a Bayesian Gaussian regression model using stan_glm 
ref_fitf <- stan_glm(rmf, data = women, family = gaussian())
# Check reference model, MCMC check
summary(ref_fitf)
#shinystan::launch_shinystan(ref_fitf)

mcmc_areas(as.matrix(ref_fitf)[,2:12])
```


```{r LOOPIT and bayesrules summary}

pp_check(ref_fitf)
y_rep_rmf <- posterior_predict(ref_fitf)
loo_rmf <- loo(ref_fitf , save_psis = TRUE, cores = 4)
psis1_rmf <- loo_rmf$psis_object
lw_rmf <- weights(psis1_rmf)
color_scheme_set("red")
#pdf(file.path(PPath2, "/LOOPITrmf.pdf"))
ppc_loo_pit_overlay(y= women$Fat,yrep=y_rep_rmf,lw=lw_rmf)
#dev.off()

# Summarize the posterior distributions simple model
summary_rmf_cv <- prediction_summary_cv(model = ref_fitf , data = women, k = 3)

# Print the summary simple model
print(summary_rmf_cv)

```

##### Compute the probabilty for each estimator that it is close to zero 
```{r close to zero}

pred <- c("Waist" ,"Height" ,"Weight" ,"Age", "Neck" ,"Chest" , 
                          "Biceps" ,"Hips" , "Forearm",
                        "Knee", "Ankle")

# Extract the posterior samples for the estimator
posterior_samples <- as.matrix(ref_fitf, pars = pred)

# Define the range
lower_bound <- -0.05
upper_bound <- 0.05

# Apply the condition to each element of the matrix
prob_between <- apply(posterior_samples, 2, function(par) mean(par >= lower_bound & par <= upper_bound))

# Print the result
print(prob_between)
print(prob_between[prob_between > 0.25])

```

##### Use projpred to downsize model even more
```{r predictive projection, results='hide'}
set.seed(2341)
vs_cvf <- cv_varsel(ref_fitf)

save(vs_cvf,file = file.path(PPath2,"vs_cvf"))
#load(file = file.path(PPath2,"vs_cv"))
nsel <- suggest_size(vs_cvf)
nsel

rank_vs_cvf <- ranking(vs_cvf)
rank_vs_cvf
solterms_final_vs_cvf <- head(rank_vs_cvf$fulldata ,suggest_size(vs_cvf))

solterms_final_vs_cvf

# Print the summary of the variable selection 
summary(vs_cvf) 

# Plot the variable selection 
#pdf(file.path(PPath2, "cv_valsel_plotf.pdf"))
plot(vs_cvf, deltas = TRUE)
#dev.off()
# 


```

```{r best five predictors}
solterms_final_fivef <- head(rank_vs_cvf$fulldata, 5)
solterms_final_fivef 
```



##### Projected posterior for the selected model
```{r post-selection inference, results='hide'}
# https://cran.r-project.org/web/packages/projpred/vignettes/projpred.html#post-selection-inference
proj_ref_fitf <- project(vs_cvf, predictor_terms = solterms_final_fivef, ns = 4000)

```
##### Show results
```{r Show results}
color_scheme_set(scheme = "green")
mcmc_areas(as.matrix(proj_ref_fitf), pars = solterms_final_fivef)# 
```

###### Derive second reference model based on FBMS: Using BMA, marginal likelihoods inherently account for model complexity, and PIPs derived from them reflect this balance.
```{r second reference model based on FBMS, results='hide', message=FALSE}
#Marginal likelihoods inherently account for model complexity by integrating over all possible parameter values of #a model, rather than focusing on specific parameter estimates. This process naturally penalizes more complex #models, as they tend to spread their likelihood across a larger parameter space. Simpler models, with fewer #parameters, often concentrate their likelihood more narrowly, giving them an advantage in terms of marginal #likelihood.

transforms <- c("p2","p3")#  transforms planed in my re-sap

probs <- gen.probs.gmjmcmc(transforms)
probs$gen <- c(0,1,0,1) # Only modifications!
params <- gen.params.gmjmcmc(women[,c("Waist","Height","Weight","Age","Neck","Chest","Biceps","Hips","Forearm","Knee","Ankle","BMI")])
params$feat$pop.max <- 20 #set the population size to be 20
params$feat$D <- 1   # Set depth of features to 1
#@ Aliaksandr(1): please decide about this options
params$loglik$r <- exp(-2) # to correspond to AIC complexity that Georg uses in mfp
params$loglik$var <- "unknown"
modelFBMS <- as.formula(Fat ~ Waist + Height + Weight + Age + Neck + Chest +  
                          Biceps + Hips + Forearm 
                        +  Knee +  Ankle + BMI)

set.seed(1234)

result <- fbms(modelFBMS, data = women, method = "gmjmcmc.parallel", 
                family = "gaussian",
               transforms = transforms, probs = probs, params = params, P = 50,
              runs = 40, cores = 10)

```
##### The final model, which includes five predictors, is  identical
```{r formula_male}

summary(result, labels = c("Waist","Height","Weight","Age","Neck","Chest","Biceps","Hips","Forearm","Knee","Ankle","BMI"),tol = 0.1)

formula_male <- as.formula(Fat ~ Waist  + Hips + Height + Knee + Ankle)
```

##### Fit model to male data set and compare the posterior predictive distributions
##### See Task Bodyfat.qmd
```{r read and prepare male data}
library(mfp)

data(bodyfat)

bodyfat_corr <- bodyfat 

bodyfat_corr[42,"height"] <- 69.5       # one-digit correction of height
bodyfat_corr[48,"density"] <- 1.0865    # one-digit correction of density
bodyfat_corr[76,"density"] <- 1.0566    # one-digit correction of density
bodyfat_corr[96,"density"] <- 1.0591    # one-digit correction of density

# recompute siri formula based on corrected density

bodyfat_corr[,"siri"]   <- pmax(round(495 / bodyfat_corr$density - 450, 1) , 0)
bodyfat_corr[,"brozek"] <- pmax(round(457/ bodyfat_corr$density - 414.2, 1) , 0)
```
###### Aligning the two data sets
```{r Aligning the two data sets}
bodyfat_corr$height <- bodyfat_corr$height * 2.54/100
bodyfat_corr$weight <- bodyfat_corr$weight * 0.454592

library(tidyverse)

men <- bodyfat_corr %>%
  rename(Fat = siri, Age = age, Height = height, Weight = weight, 
         Neck = neck, Chest = chest, Waist = abdomen, Hips = hip, 
         Knee = knee, Ankle = ankle, Forearm = forearm, 
         Biceps = biceps, Thigh = thigh)
names(men)
# Add BMI
men$BMI <- men$Weight / (men$Height^2)
men[,4:18] <- scale(men[,4:18])

```

##### Prepare data for males
```{r prepare data for males, message = FALSE, warning = FALSE}
fat_density_male <- density(men$Fat, na.rm = TRUE)
ggplot(men, aes(x = Fat)) + geom_density() + 
  labs(title = "Kernel-density plot for outcome fat for males", x = "Fat", y = "Density")
summary(men$Fat)

```


```{r LOOPIT and bayesrules summary for male data,results='hide'}
# LOOPIT and bayesrules summary for male data based on female model
# Prediction for male data based on female model
fit_female <- stan_glm(formula_male, data = women, family = gaussian())
#proj_ref_fitf <- project(vs_cvf, predictor_terms = solterms_final_fivef, ns = 4000)
# https://www.rdocumentation.org/packages/projpred/versions/2.0.2/topics/proj-pred
#y_rep_male <- proj_linpred(vs_cvf, solution_terms= solterms_final_fivef, newdata = men)#
y_rep_male <- posterior_predict(fit_female, newdata = men)
color_scheme_set("blue")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male[1:50,],fun = ppc_dens_overlay)

binary_waist_male <- ifelse(men$Waist <= 0, 0, 1)
binary_height_male <- ifelse(men$Height <= 0, 0, 1)
binary_hips_male <- ifelse(men$Hips <= 0, 0, 1)
binary_knee_male <- ifelse(men$Knee <= 0, 0, 1)
binary_ankle_male <- ifelse(men$Ankle <= 0, 0, 1)
color_scheme_set("pink")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_waist_male, stat = "median")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_height_male, stat = "median")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_hips_male, stat = "median")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,fun = "stat_grouped", 
                    group = binary_knee_male, stat = "median")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,fun = "stat_grouped", 
                    group = binary_ankle_male, stat = "median")
color_scheme_set("gray")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_waist_male, stat = "mean")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_height_male, stat = "mean")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = binary_hips_male, stat = "mean")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,fun = "stat_grouped", 
                    group = binary_knee_male, stat = "mean")

bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,fun = "stat_grouped", 
                    group = binary_ankle_male, stat = "mean")


```

##### Summarize the posterior distributions
```{r Summarize the posterior distributions}
# Summarize the posterior distributions 
summary_male_cv <- prediction_summary_cv(model = fit_female, data = men, k = 3)

# Print the summary 
print(summary_male_cv)

summary_male <- prediction_summary(model = fit_female, data = men)

# Print the summary 
print(summary_male)
```



##### Identify the groups where the prediction performs effectively

```{r define different groups}
age_group <- cut(men$Age, breaks = quantile(men$Age, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)
table(age_group)
fat_group <- cut(men$Fat, breaks = quantile(men$Fat, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)
table(fat_group)
waist_group <- cut(men$Waist, breaks = quantile(men$Waist, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE) # table(waist_group)

color_scheme_set(scheme = "purple")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = age_group, stat = "median")
color_scheme_set(scheme = "yellow")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = fat_group, stat = "median")
color_scheme_set(scheme = "teal")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = waist_group, stat = "median")


color_scheme_set(scheme = "purple")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = age_group, stat = "mean")
color_scheme_set(scheme = "yellow")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = fat_group, stat = "mean")
color_scheme_set(scheme = "teal")
bayesplot::pp_check(object = men$Fat, yrep = y_rep_male,
                    fun = "stat_grouped", group = waist_group, stat = "mean")
```

