---
output:
  html_document: default
  pdf_document: default
  word_document: default
  html_notebook: default
---
---
title: "Statistical re-analysis plan: Predicting body fat proportion using anthropometric measures"
author: "Mariana Nold" 
---

########  Run Task Bodyfat simplified first.

##### Load R packages

```{r load package, message = FALSE, warning = FALSE}
library("rstanarm")
library("projpred")
library("ggplot2")
library("loo")
library("bayesplot")
library("bayesrules")
library("DHARMa")
library("ggplot2")
library("dplyr")
library("tidyr")
library("patchwork")
library("posterior")
```


```{r set path and load data (female and male)}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task Bodyfat/"
load("~/GitHub/Bayes_for_STRATOS/Task Bodyfat/Datafiles/bodyfatdata.Rdata")
#scale data sets


# Save standard deviations before scaling (excluding first two columns)
sd_women <- apply(women[, -c(1,2)], 2, sd)
women[, -c(1,2)] <- scale(women[, -c(1,2)])

# Save standard deviations before scaling (excluding first three columns)
sd_men <- apply(men[, -c(1,2,3)], 2, sd)
men[, -c(1,2,3)] <- scale(men[, -c(1,2,3)])


# Display summary statistics for validation
#summary(women)
sd_women
sd(women$Fat)
#summary(men)
sd_men
sd(men$Fat)
```

```{r fit full model with rstanarm, results='hide', cache=TRUE}

rmf <- as.formula(Fat ~ Waist + Height + Weight + Age + Neck + Chest +  
                          Biceps + Hips + Forearm 
                        +  Knee +  Ankle + BMI)

# Fit a Bayesian Gaussian regression model using stan_glm 
ref_fit <- stan_glm(rmf, data = women, family = gaussian())
# Check reference model, MCMC check
summary(ref_fit)
#shinystan::launch_shinystan(ref_fitf)

mcmc_areas(as.matrix(ref_fit)[,2:13])
```

##### Investigate quality of the reference model

```{r LOOPIT and bayesrules summary, cache=TRUE}
prior_summary(ref_fit) # https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html
pp_check(ref_fit)
y_rep_rmf <- posterior_predict(ref_fit)
loo_rmf <- loo(ref_fit , save_psis = TRUE, cores = 4)
psis1_rmf <- loo_rmf$psis_object
lw_rmf <- weights(psis1_rmf)
color_scheme_set("red")

ppc_loo_pit_overlay(y = women$Fat,yrep=y_rep_rmf,lw=lw_rmf)


# Summarize the posterior distributions simple model
summary_rmf_cv <- prediction_summary_cv(model = ref_fit , data = women, k = 3)

# Print the summary simple model
print(summary_rmf_cv)

```

##### Use projpred to downsize model even more

```{r predictive projection, results='hide', cache=TRUE}
set.seed(2341)
vs_cvf <- cv_varsel(ref_fit, method = 'forward', cv_method = 'loo')

save(vs_cvf,file = file.path(PPath,"vs_cvf"))
#load(file = file.path(PPath,"vs_cvf"))


plot(vs_cvf, deltas = TRUE, stats =  c('elpd','rmse'))
```


```{r define functions, include = FALSE}

visualize_variable_selection <- function(vs_cvf) {

  # Extract the summary and store as return object
  model_summary <- summary(vs_cvf, deltas = FALSE, cumulate = FALSE)

  # Extract predictor ranking
  rk <- ranking(vs_cvf)

  # Compute variable ranking proportions
  pr_rk <- cv_proportions(rk)
  pr_rk2 <- cv_proportions(rk, cumulate = TRUE)
  # Generate plots
  gg_pr_rk <- plot(pr_rk) + theme(legend.position = "none")  
  gg_pr_rk2 <- plot(pr_rk2) + theme(legend.position = "none")
  combined_plot <- gg_pr_rk / gg_pr_rk2 


  # Return both the summary and the plot
  return(list(summary = model_summary, plot = combined_plot))
}

#' Extract and round a single model size row (e.g. "<=5") from cv_proportions
#'
#' @param cumprop Cumulative proportions matrix from cv_proportions(..., cumulate = TRUE)
#' @param size Integer value, the exact model size to extract (default: 5)
#' @param digits Number of decimal places to round to (default: 3)
#' @return A 1-row matrix with rounded values
get_cumprop_row_rounded <- function(cumprop, size = 5, digits = 3) {
  rowname <- paste0("<=", size)
  if (!(rowname %in% rownames(cumprop))) {
    stop(paste0("Model size '<=", size, "' not found in row names."))
  }
  row <- cumprop[rowname, , drop = FALSE]
  row[] <- round(row, digits = digits)
  return(row)
}

# Function to back-transform standardized regression coefficients to unstandardized ones
# when only predictors (x) were standardized and y remained in the original scale

compute_unstandardized_coefs <- function(proj_ref_fit, sd_x) {
  # proj_ref_fit: object (e.g. matrix) containing standardized regression coefficients
  # sd_x: numeric vector of standard deviations of the predictor variables (x)

  # Extract coefficients as a matrix
  
  # Convert to matrix (if not already)
    mat <- as.matrix(proj_ref_fit)

  # Remove first and last columns
    beta_std  <- mat[, 2:(ncol(mat) - 1)]

  # Check if the number of standard deviations matches the number of coefficients
  if (length(sd_x) != ncol(beta_std)) {
    stop("Length of 'sd_x' must match the number of coefficients in 'proj_ref_fit'.")
  }

  # Back-transform: unstandardized beta = standardized beta / SD(x)
  beta_unstd <- sweep(beta_std, 2, sd_x, "/")


  return(beta_unstd)
}


```


```{r stability of selection process }

result <- visualize_variable_selection(vs_cvf)
result$summary    # inspect summary
print(result$plot)  # view the plot


cumprop <- cv_proportions(ranking(vs_cvf), cumulate = TRUE) 
rounded_result <- get_cumprop_row_rounded(cumprop, size = 5)
print(rounded_result)


```


```{r best five predictors}
nsel <- suggest_size(vs_cvf)
nsel

rank_vs_cvf <- ranking(vs_cvf)
rank_vs_cvf$fulldata

solterms_final_vs_cvf <- head(rank_vs_cvf$fulldata ,suggest_size(vs_cvf))

solterms_final_vs_cvf


solterms_final_fivef <- head(rank_vs_cvf$fulldata, 5)
solterms_final_fivef 
```

##### Use 5-fold cross validation instead

```{r 5-fold cross validation, results='hide', cache=TRUE}
set.seed(2341)
vs_cvs <- cv_varsel(
  object = ref_fit,         
  cv_method = "kfold",      
  K = 5,                   
  method = "forward",        
  validate_search = TRUE, verbose = FALSE  
)

save(vs_cvs,file = file.path(PPath,"vs_cvs"))
#load(file = file.path(PPath,"vs_cvs"))


plot(vs_cvs, deltas = TRUE, stats =  c('elpd','rmse'))

```

##### Stability of 5-fold cross validation (forward selection)

```{r stability of 5-fold cross validation}
results <- visualize_variable_selection(vs_cvs)
results$summary    # inspect summary
print(results$plot)  # view the plot


cumprops <- cv_proportions(ranking(vs_cvs), cumulate = TRUE) 
rounded_results <- get_cumprop_row_rounded(cumprops, size = 5)
print(rounded_results)

```

##### Best five predictors with 5-fold cv

```{r best five predictors with 5-fold cv}
nsel <- suggest_size(vs_cvs)
nsel

rank_vs_cvs <- ranking(vs_cvs)
rank_vs_cvs$fulldata

solterms_final_vs_cvs <- head(rank_vs_cvs$fulldata ,suggest_size(vs_cvs))

solterms_final_vs_cvs


solterms_final_fives <- head(rank_vs_cvs$fulldata, 5)
solterms_final_fives 
```

#### L1 serach (LOO)

```{r predictive lasso projection, results='hide', cache=TRUE}
# Generate model path: By varying lambda, a sequence of submodels with an increasing number of (non-zero) coefficients is created. ➤ The order in which coefficients “appear” serves as a ranking of the variables.

set.seed(2341)
vs_cvl <- cv_varsel(ref_fit, method = 'L1', verbose = FALSE)

save(vs_cvl,file = file.path(PPath,"vs_cvl"))
#load(file = file.path(PPath,"vs_cvl"))


```

##### Summarize stability lasso cv

```{r Summarize stability lasso}
# Plot the variable selection 

plot(vs_cvl, deltas = TRUE, stats =  c('elpd','rmse'))

# 
resultl <- visualize_variable_selection(vs_cvl)
resultl$summary    # inspect summary
print(resultl$plot)  # view the plot


cumpropl <- cv_proportions(ranking(vs_cvl), cumulate = TRUE) 
rounded_resultl <- get_cumprop_row_rounded(cumpropl, size = 5)
print(rounded_resultl)

```


##### Solution terms Lasso

```{r solution terms Lasso serach}
# Print the summary of the variable selection 
summary(vs_cvl) 

nsel <- suggest_size(vs_cvl)
nsel

rank_vs_cvl <- ranking(vs_cvl)
rank_vs_cvl$fulldata

solterms_final_vs_cvl <- head(rank_vs_cvl$fulldata ,suggest_size(vs_cvl))

solterms_final_vs_cvl


solterms_final_fivel <- head(rank_vs_cvl$fulldata, 5)
solterms_final_fivel
```


#### L1 serach (K-fold)

```{r predictive lasso projection k-fold, results='hide', cache=TRUE}
# Generate model path: By varying lambda, a sequence of submodels with an increasing number of (non-zero) coefficients is created. ➤ The order in which coefficients “appear” serves as a ranking of the variables.

set.seed(2341)
vs_cvl5 <- cv_varsel(
  object = ref_fit,
  method = 'L1',
  cv_method = "kfold",
  K = 5, verbose = FALSE
)#cv_varsel(ref_fit, method = 'L1', verbose = FALSE)

save(vs_cvl5,file = file.path(PPath,"vs_cvl5"))
#load(file = file.path(PPath,"vs_cvf"))

# Plot the variable selection 

plot(vs_cvl5, deltas = TRUE, stats =  c('elpd','rmse'))

```


##### Summarize stability lasso 5-fold

```{r Summarize stability lasso 5-fold}

resultl5 <- visualize_variable_selection(vs_cvl5)
resultl5$summary    # inspect summary
print(resultl5$plot)  # view the plot


cumpropl5 <- cv_proportions(ranking(vs_cvl5), cumulate = TRUE) 
rounded_resultl5 <- get_cumprop_row_rounded(cumpropl5, size = 5)
print(rounded_resultl5)

```


##### Solution terms Lasso

```{r solution terms Lasso serach, 5-fold}
# Print the summary of the variable selection 
summary(vs_cvl5) 

nsel <- suggest_size(vs_cvl5)
nsel

rank_vs_cvl5 <- ranking(vs_cvl5)
rank_vs_cvl5$fulldata

solterms_final_vs_cvl5 <- head(rank_vs_cvl5$fulldata ,suggest_size(vs_cvl5))

solterms_final_vs_cvl5


solterms_final_fivel5 <- head(rank_vs_cvl5$fulldata, 5)
solterms_final_fivel5
```


##### Projected posterior for the selected model (L1 serach cv_method = LOO)

```{r post-selection inference, results='hide'}
# https://cran.r-project.org/web/packages/projpred/vignettes/projpred.html#post-selection-inference
proj_ref_fit <- project(vs_cvl, predictor_terms = solterms_final_fivel, ns = 4000)

```

##### Show results

```{r Show results}
color_scheme_set(scheme = "green")
mat <- as.matrix(proj_ref_fit)
beta_std <- mat[, 2:(ncol(mat) - 1)]

# Transform matrix to draws format (optional, for posterior-friendly summaries)
draws1 <- posterior::as_draws_matrix(beta_std)

summarise_draws(draws1, mean, sd, median, ~quantile(.x, probs = c(0.05, 0.95)))
mcmc_areas(as.matrix(proj_ref_fit), pars = solterms_final_fivel)# 
```

##### unstandardized coefs

```{r unstandardized_coefs}
vars_to_select <- c("BMI", "Waist", "Hips", "Knee", "Biceps")
sd_selected <- sd_women[intersect(vars_to_select, names(sd_women))]
color_scheme_set(scheme = "purple")


beta_unstd <- compute_unstandardized_coefs(proj_ref_fit, sd_selected)

# Transform matrix to draws format (optional, for posterior-friendly summaries)
draws <- posterior::as_draws_matrix(beta_unstd)

summarise_draws(draws, mean, sd, median, ~quantile(.x, probs = c(0.05, 0.95)))


mcmc_areas(beta_unstd)
```


```{r define formula}
formula_male <- as.formula(Fat ~ Waist  + Hips + BMI + Knee + Biceps)
```

###### Model diagnostics (1.3.3)

```{r Model diagnostics}
y_linpred_female <- proj_linpred(proj_ref_fit, solution_terms= solterms_final_fivel, newdata = women)
women2 <- cbind(women, linpred = colMeans(y_linpred_female$pred))
plotLocBias <- ggplot(data.frame(fitted.values = women2$linpred, y = women2$Fat), aes(x=fitted.values, y=y)) + geom_point() + geom_smooth() + geom_abline(slope=1, intercept=0)
plotLocBias
# Compute residuals
prj_predictW <- proj_predict(proj_ref_fit)

qres.m1=createDHARMa(simulatedResponse = t(prj_predictW ), 
                          observedResponse = women2$Fat, 
                          fittedPredictedResponse = apply(prj_predictW , 2, median), integerResponse = F, seed = 123, method = "PIT") 
summary(residuals(qres.m1))
# QQ-Plot 
res <- qnorm(residuals(qres.m1))

# QQ-Plot mit ggplot2
res.plot <- ggplot(data.frame(res), aes(sample = res)) +
  stat_qq(color = "steelblue", size = 2) +
  stat_qq_line(color = "firebrick", linetype = "dashed") +
  labs(
     title = "QQ Plot of Transformed Residuals",
    x = "Theoretical Quantiles",
    y = "Observed Quantiles"
  ) +
  theme_minimal(base_size = 14)
res.plot



# Check heteroskedasticity (1.3.3.2)
plotHet <- ggplot(data.frame(fitted.values = women2$linpred, y=sqrt(abs(res))), aes(x=fitted.values, y=y)) + geom_point() + geom_smooth() 
plotHet 
# Check association with excluded predictors
plotExcPred <- ggplot(data.frame(Weight=women2$Weight, Residuals=res), aes(x=Weight, y=Residuals)) + geom_point() + geom_smooth() 
plotExcPred

ggplot(data=data.frame(residuals=res), aes(x=res)) + geom_histogram(bins=20)
# Apparent performance of the model (1.3.4)
R.square <- (cor(women2$linpred, women2$Fat, method = "pearson"))^2
R.square
color_scheme_set(scheme = "purple")
mcmc_areas(as.matrix(proj_ref_fit), pars = "sigma")
summary(as.matrix(proj_ref_fit)[, "sigma"])
```
##### Internal validation
##### Calibration intercept and slope (internal validation)

```{r compute Calibration intercept and slope (internal), results='hide'}
observed <- women2$Fat
predicted <- women2$linpred
calib_dataW <- data.frame(observed = observed, predicted = predicted)
# Compute calibration intercept and slope
calibration_modelW <- stan_glm(observed ~ predicted, family = gaussian(), data = calib_dataW)
```

```{r Calibration intercept and slope (internal)}


summary(calibration_modelW)

posterior_samplesW <- as.matrix(calibration_modelW)
intercepts <- posterior_samplesW[, "(Intercept)"]
slopes <- posterior_samplesW[, "predicted"]

# Dataframe  for regression lines
regression_linesW <- data.frame(
  intercept = intercepts,
  slope = slopes
)

n_lines <- nrow(as.matrix(calibration_modelW))
print(n_lines)
sampled_lines <- regression_linesW[sample(nrow(regression_linesW), size = 100), ]
# plot
ggplot(calib_dataW, aes(x = predicted, y = observed)) +
  geom_point(alpha = 0.9) +
  geom_abline(data = sampled_lines, aes(intercept = intercept, slope = slope), alpha = 0.1, color = "darkgreen") +
  labs(title = "Calibration Plot with regression lines (internal)",
       x = "Predicted",
       y = "Observed") +
  theme_minimal()
color_scheme_set("gray")
summary(as.matrix(calibration_modelW))

mcmc_areas(as.matrix(calibration_modelW),
           pars = c("(Intercept)", "predicted"),
           prob = 0.95) + 
  ggtitle("Calibration Intercept & Calibration slope (internal)")



# Intercept
mcmc_areas(
  as.matrix(calibration_modelW),
  pars = "(Intercept)",
  prob = 0.95
) +
  ggtitle("Calibration Intercept (internal)")

# Slope
mcmc_areas(
  as.matrix(calibration_modelW),
  pars = "predicted",
  prob = 0.95
) +
  ggtitle("Calibration Slope (internal)")



mcmc_areas(as.matrix(calibration_modelW),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model (internal)")
```

##### External validation

```{r predictions for males based on female model}
y_linpred_male <- proj_linpred(proj_ref_fit, solution_terms= solterms_final_fivel, newdata = men)


color_scheme_set("blue")
prj_predictM <- proj_predict(proj_ref_fit, newdata = men)
# Using the 'bayesplot' package:
ppc_dens_overlay(y = men$Fat, yrep = prj_predictM)

# 3.1
eval_men <- cbind(men, linpred = colMeans(y_linpred_male$pred))
# performance measures at external validation
rmspe_ext <- sqrt(mean((eval_men$Fat - eval_men$linpred)^2))
rmspe_ext
Rsqu_ext <- 1 - (mean((eval_men$Fat - eval_men$linpred)^2))/var(eval_men$Fat)
Rsqu_ext
R.squareM <- (cor(eval_men$linpred, eval_men$Fat, method = "pearson"))^2
R.squareM

calPlot <- ggplot(eval_men, aes(linpred, Fat)) + geom_point() + geom_smooth() + geom_smooth(method="lm", col="red") + geom_abline(intercept=0, slope=1) 
calPlot
calResPlot <- ggplot(eval_men, aes(x=linpred, y=Fat-linpred)) + geom_point() + geom_smooth()
calResPlot
hetPlot <- ggplot(eval_men, aes(x=linpred, y=sqrt(abs(Fat-linpred)))) + geom_point() + geom_smooth()
hetPlot
# Predictors: "BMI"   "Hips"  "Waist"  "Knee"
nonpredictors <- c("Height", "Weight", "Age", "Neck", "Chest", "Ankle",  "Forearm") 

eval_men_long <- eval_men %>% pivot_longer(cols = all_of(nonpredictors), names_to = "NonPredictor", values_to = "Value") 

plotAll <- ggplot(eval_men_long, aes(x = Value, y = Fat - linpred)) + geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE, color = "blue") + facet_wrap(~ NonPredictor, scales = "free_x") + labs(title = "Non-Predictors versue residuals (Fat - linpred)", x = "Non-Predictor", y = "Residuals (Fat - linpred)") + theme_minimal() 

plotAll
```



##### Calibration intercept and slope (external validation)

```{r compute Calibration intercept and slope (external), results='hide'}
observed <- eval_men$Fat
predicted <- eval_men$linpred
calib_data <- data.frame(observed = observed, predicted = predicted)
# Compute calibration intercept and slope
calibration_model <- stan_glm(observed ~ predicted, family = gaussian(), data = calib_data)
```

```{r Calibration intercept and slope (external)}
summary(calibration_model)

posterior_samples <- as.matrix(calibration_model)
intercepts <- posterior_samples[, "(Intercept)"]
slopes <- posterior_samples[, "predicted"]

# Dataframe  for regression lines
regression_lines <- data.frame(
  intercept = intercepts,
  slope = slopes
)

n_lines <- nrow(as.matrix(calibration_model))
print(n_lines)
sampled_lines <- regression_lines[sample(nrow(regression_lines), size = 100), ]
# plot
ggplot(calib_data, aes(x = predicted, y = observed)) +
  geom_point(alpha = 0.9) +
  geom_abline(data = sampled_lines, aes(intercept = intercept, slope = slope), alpha = 0.1, color = "blue") +
  labs(title = "Calibration Plot with regression lines",
       x = "Predicted",
       y = "Observed") +
  theme_minimal()
color_scheme_set("orange")
summary(as.matrix(calibration_model))


mcmc_areas(as.matrix(calibration_model),
           pars = c("(Intercept)", "predicted"),
           prob = 0.95) + 
  ggtitle("Calibration Intercept & Calibration slope (external)")



# Intercept
mcmc_areas(
  as.matrix(calibration_model),
  pars = "(Intercept)",
  prob = 0.95
) +
  ggtitle("Calibration Intercept (external)")

# Slope
mcmc_areas(
  as.matrix(calibration_model),
  pars = "predicted",
  prob = 0.95
) +
  ggtitle("Calibration Slope (external)")



mcmc_areas(as.matrix(calibration_model),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model (external)")
```







