---
title: "Analysis (simplified): bodyfat prediction"
author:
  - name: Georg Heinze
    affiliation: Medical University of Vienna, Center for Medical Data Science
date: last-modified                            # "`r Sys.Date()`" for system date
categories: [2024, Contrast-Bayes-Frequentist, STRATOS, R]
description: "Classification: confidential"
editor: visual
execute:
  warning: false
theme: cosmo                                   # cosmo bootswatch theme
toc: true                                      # table of contents  
number-sections: true
colorlinks: true                               # ??
highlight-style: pygments
#bibliography: references.bib
format:
  html: 
    code-fold: true
    code-tools: true
    html-math-method: katex
    self-contained: true                      # html is self-contained
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
  docx: default                               
---

## Preliminaries

This analysis report relates to SAPI V1.1 for the Analysis of bodyfat example (simplified).


## Load libraries

```{r}
library(mfp2)

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(gt)
library(gtExtras)
library(patchwork)
library(Hmisc)
library(sjlabelled)
```

# Model estimation

## Load data

```{r}
load("../Datafiles/bodyfatdata.Rdata")
```

## Main data analysis

### Description of observational units

```{r}
library(dplyr)
library(table1)


# Assume these are your two existing datasets
# data_women <- ...
# data_men <- ...

# Add Gender variable to each
data_women <- women %>% mutate(Gender = "1-Women")
data_men   <- men %>% mutate(Gender = "2-Men")

# Combine them into one dataset
data <- bind_rows(data_women, data_men)


my.render.cont <- function(x) {
        sprintf("%0.1f (%0.1f)\n[%0.1f – %0.1f]", 
               mean(x, na.rm=TRUE), sd(x, na.rm=TRUE), min(x, na.rm=TRUE), max(x, na.rm=TRUE))
}

# Label variables
label(data$Fat) <- "Body Fat (%)"
label(data$Weight) <- "Weight (kg)"
label(data$Height) <- "Height (m)"
label(data$BMI) <- "BMI (kg/m^2)"
label(data$Age) <- "Age (years)"
label(data$Neck) <- "Neck (cm)"
label(data$Chest) <- "Chest (cm)"
label(data$Calf) <- "Calf (cm)"
label(data$Biceps) <- "Biceps (cm)"
label(data$Hips) <- "Hips (cm)"
label(data$Waist) <- "Waist (cm)"
label(data$Forearm) <- "Forearm (cm)"
label(data$PThigh) <- "Proximal Thigh (cm)"
label(data$MThigh) <- "Middle Thigh (cm)"
label(data$DThigh) <- "Distal Thigh (cm)"
label(data$Wrist) <- "Wrist (cm)"
label(data$Knee) <- "Knee (cm)"
label(data$Elbow) <- "Elbow (cm)"
label(data$Ankle) <- "Ankle (cm)"

# Generate Table 1


table1(~ Fat + Weight + Height + BMI + Age + Neck + Chest  + Biceps + 
         Hips + Waist + Forearm +  MThigh  + Wrist + Knee + 
          Ankle | Gender, data = data, render.continuous = my.render.cont, overall = NULL)
```


### Objective 1: Model development and internal assessment

#### Model estimation

Define a function that fits the model.

```{r}
# Function to derive the model with p predictors

# x = predictor matrix with column names being the labels



fit_rlasso <- function(x, y, p_final){
  n <- nrow(x)
  fit_glmnet <- glmnet(x=x, y=y, family="gaussian", relax=T)
  coef_mat <- coef(fit_glmnet$relaxed)
  num_nonzero <- apply(coef_mat, 2, function(v) sum(v != 0))-1
  target_lambda_index <- max(which(num_nonzero <= p_final))
  target_s <- min(fit_glmnet$lambda[num_nonzero <=p_final])
  active_coef <- coef_mat[, target_lambda_index]
  active_predictors <- rownames(coef_mat)[which(active_coef != 0)]
  fitted <- cbind(1,x) %*% active_coef
  residuals <- y - fitted
  residual_se <- sqrt(sum(residuals^2)/(n-1-p_final))
  return(list(coef=active_coef, residual_se=residual_se, residuals=residuals, fitted=fitted, relaxed=fit_glmnet$relaxed, fit=fit_glmnet, s=target_s))
}
```


Fit the model:


```{r, message=TRUE}
library(glmnet)

predictors <- c("Weight", "Height", "BMI", "Age",
  "Neck", "Chest",  "Biceps", "Hips", "Waist",
  "Forearm",  "MThigh",  "Wrist",
  "Knee", "Ankle")
p <- length(predictors)

x <- as.matrix(women[, predictors])
y <- as.numeric(women$Fat)
n <- nrow(x)

# fit the coefficient path
fit_glmnet <- fit_rlasso(x=x, y=y, p_final=5)


ypred_orig <- predict(fit_glmnet$relaxed, newx=x, s=fit_glmnet$s)
```

#### Model stability

Repeat the model building process 1000 times on bootstrap samples and store the coefficients and the predictions from the model on the original data:



```{r}
p_final <- 5
sel_orig <- fit_glmnet$coef !=0

B <- 1000

coef_b <- matrix(NA, nrow=B, ncol=ncol(x)+1)
colnames(coef_b) <- names(fit_glmnet$coef)

pred_b <- matrix(NA, nrow=B, ncol=nrow(x))

rsq_app <- rsq_test <- se_b <- model_sel <- (1:B)*0

set.seed(7123981)

for(b in 1:B){
  indb <- sample(1:nrow(x), repl=T)
  xb <- x[indb,]
  yb <- y[indb]
  
  fitb <- fit_rlasso(x=xb, y=yb, p_final=p_final)
  
  pred_b[b,] <- predict(fitb$relaxed, s=fitb$s, newx=x)
  coef_b[b,][names(fitb$coef)] <- fitb$coef
  model_sel <- prod((fitb$coef!=0) == sel_orig) # 1 if exactly equal to originally selected model, 0 else
  se_b[b] <- fitb$residual_se
  
  rsq_app[b] <- 1 - fitb$residual_se^2/var(yb)
  rsq_test[b] <- 1- var(predict(fitb$relaxed, s=fitb$s, newx=x) - y)/var(y)
}


sel_prob <- apply(coef_b, 2, function(X) mean(X!=0))   # selection probabilities for predictors
model_prob <- mean(model_sel)   # model selection probability

```

Evaluate stability results:

Instability plot of predictions

```{r}
matplot(fit_glmnet$fitted, t(pred_b[1:100,]), 
        pch = 16,              # solid dots
        cex = 0.3,             # small point size
        col = rgb(0, 0, 0, 0.1), # semi-transparent black
        main = "Prediction instability plot",
        xlab = "Original prediction", ylab = "Bootstrapped prediction")
```


Bootstrap RMSE of prediction vs. model-based standard error of fitted values

```{r}
# bootstrap RMSE of prediction
rmse_pred <- apply(pred_b, 1, function(X) sqrt(mean((fit_glmnet$fitted - X)^2)))

cat("Bootstrap RMSE of fitted values (selected):", mean(rmse_pred), "\n")

# model-based RMSE of prediction
xmod <- cbind(1,x[,fit_glmnet$coef[-1]!=0])
invXtX <-solve(t(xmod) %*% xmod)

mse_mod <- sapply(1:nrow(x), function(X) fit_glmnet$residual_se^2*t(xmod[X,]) %*% invXtX %*% xmod[X,])
cat("(Model-based RMSE of fitted values:", sqrt(mean(mse_mod)), ")\n")

```
```{r}
# Pairwise selection probabilities

sel_b <- (coef_b!=0)[,-1]
ncoef <- ncol(sel_b)

exp_pw_sel <- pw_sel <- zscore <- matrix(0,nrow=ncoef, ncol=ncoef)
rownames(pw_sel) <- colnames(pw_sel) <- rownames(zscore) <- colnames(zscore) <- predictors

diag(pw_sel) <- sel_prob[-1]



for(j in 1:(ncoef-1)){
  for(jj in (j+1):ncoef){
    exp_pw_sel[j,jj] <- sel_prob[j+1]*sel_prob[jj+1]
    pw_sel[jj,j] <- pw_sel[j,jj] <- mean(sel_b[,j]*sel_b[,jj])
    zscore[jj,j] <- zscore[j,jj] <- sign(pw_sel[j,jj]-exp_pw_sel[j,jj])*sqrt((pw_sel[j,jj]-exp_pw_sel[j,jj])^2/exp_pw_sel[j,jj]*B)
  }
}




```



#### Model description


```{r}
library(knitr)
# assemble data

results_beta <- matrix(0, nrow=length(fit_glmnet$coef), ncol=5)
rownames(results_beta) <- names(fit_glmnet$coef)
colnames(results_beta) <- c("Regression coefficient", "Standardized beta", "Variable inclusion frequency", "Lower 95% confidence limit", "Upper 95% confidence limit")

results_beta[,1] <- fit_glmnet$coef
results_beta[,2] <- fit_glmnet$coef*c(1,apply(x, 2, sd))
results_beta[,3] <- sel_prob
results_beta[,4] <- apply(coef_b,2,quantile, 0.025)
results_beta[,5] <- apply(coef_b,2,quantile, 0.975)

results_beta<-results_beta[order(-abs(results_beta[,3])),]
kable(results_beta, format="html", caption="Regression coefficients and stability measures")
```

Global model descriptors:

```{r}
# Apparent adjusted R-squared:
Rsq_adj <- (1 - (fit_glmnet$residual_se^2)/var(y))
Rsq_corr <- Rsq_adj - (mean(rsq_app)-mean(rsq_test))

global_desc <- c(residual_error = fit_glmnet$residual_se, boot_rmse=mean(rmse_pred), mod_rmse=sqrt(mean(mse_mod)), Rsq_adj=Rsq_adj, Rsq_corr=Rsq_corr)
names(global_desc) <- c("Residual error", "Bootstrap RMSE of fitted", "(Model-based RMSE of fitted)", "R-squared adj.", "Corrected R-squared")

kable(global_desc, caption="Global model descriptors")

```

Pairwise selection probabilities: 

```{r}
kable(pw_sel*100, caption="Pairwise selection probabilities x 100", digits=0)
```

The following table of z-scores indicates if some pairs of variables were selected overly together (z-score is positive), or alternatively to each other (z-score negative):

```{r}
kable(zscore, caption="Z-scores for nonrandom pairwise selection", digits=1)
```


#### Assumptions and diagnostics: residual analysis

```{r}
# first, find the extended set of active predictors (8 predictors)

fit_ext <- fit_rlasso(x=x, y=y, p_final=8)
active_predictors <- names(fit_glmnet$coef)[fit_glmnet$coef!=0][-1]
active_predictors_ext <- names(fit_ext$coef)[fit_ext$coef!=0][-1]
```

```{r}
library(ggplot2)

predictions <- fit_glmnet$fitted

residuals <- fit_glmnet$residuals

ggplot(data=data.frame(residuals), aes(x=residuals)) + geom_histogram() + ggtitle(label="Histogram of residuals")

sqrt_abs_res <- sqrt(abs(residuals))

ggplot(data=data.frame(residuals, predictions, y), aes(x=predictions, y=residuals)) + geom_point() + geom_smooth() + ggtitle(label="Residals vs. predictions")

ggplot(data=data.frame(sqrt_abs_res, predictions, y), aes(x=predictions, y=sqrt_abs_res)) + geom_point() + geom_smooth() + ggtitle(label="Assessment of heteroskedasticity: square root of absolute residuals vs. predictions")
```


In the following residual plots, the nearly-selected predictors are denoted by red dots, the selected predictors by blue dots.


```{r}

for(v in active_predictors_ext){
  cols=ifelse(v %in% active_predictors, "blue", "red")
  print(ggplot(data=data.frame(residuals, variable=x[,v]), aes(x=variable, y=residuals), xlab(v)) + geom_point(color=cols) + geom_smooth() + xlab(v) + ggtitle(label=paste("Residuals vs ",v, collapse=""))) 
  
}
```

### Objective 2: Model validation

```{r}
# Computing predictions in the men's data set

xmen <- as.matrix(men[, predictors])
ymen <- as.numeric(men$Fat)
ymean_men <- mean(ymen)
nmen <- nrow(xmen)

ypred_men <- predict(fit_glmnet$relaxed, s=fit_glmnet$s, newx=xmen)
res_men <- ymen - ypred_men

r2_val <- 1 - sum((ymen - ypred_men)^2)/sum((ymen - ymean_men)^2)
   #negative r2 because of miscalibration

#se_r2_val <- sqrt((4*r2_val*(1-r2_val)^2)/nmen)
#se_r2_val


rmspe <- sqrt(mean(res_men^2))

glob_pm <- c(r2_val, rmspe)
names(glob_pm) <- c("Validation R-squared", "Root mean squared prediction error")

kable(glob_pm, caption="Overall performance measures")

```

Calibration plot:

```{r}
# Calibration plot

ggplot(data.frame(ypred_men, ymen), aes(x=ypred_men, y=ymen)) + geom_point() + geom_smooth(col="red") + geom_abline(intercept=0, slope=1, linetype="dashed")
```

Calibration model:

```{r}
fit_cal <- lm(ymen ~ ypred_men)
summary(fit_cal)

cal_slope <- coef(fit_cal)[2]

fit_cal0 <- lm(ymen ~ offset(ypred_men))
cal_int <- coef(fit_cal0)[1]


cal_stats <- c(cal_int, confint(fit_cal0), cal_slope, confint(fit_cal)[2,])

names(cal_stats) <- c("Calibration intercept", "Calibration intercept, lower 95% CL", "Calibration intercept, upper 95% CL", "Calibration slope", "Calibration slope, lower 95% CL", "Calibration slope, upper 95% CL")

kable(cal_stats, caption="Calibration statistics")

```
It is easily seen that the calibration slope is very close to 1, but the intercept points towards a systematic overestimation of Fat in men by 11.9228 percentage point. As a remedy, one could propose a recalibrated model for men where the intercept is corrected by `r cal_int`. This recalibrated model would achieve an apparent $R^2$ of `r 1 - sum((ymen-ypred_men-cal_int)^2)/sum((ymen-ymean_men)^2)`.
