---
title: "Analysis (simplified): bodyfat prediction (unfinished!)"
author:
  - name: Georg Heinze
    affiliation: Medical University of Vienna, Center for Medical Data Science
date: last-modified                            # "`r Sys.Date()`" for system date
categories: [2024, Contrast-Bayes-Frequentist, STRATOS, R]
description: "Classification: confidential"
editor: visual
execute:
  warning: false
theme: cosmo                                   # cosmo bootswatch theme
toc: true                                      # table of contents  
number-sections: true
colorlinks: true                               # ??
highlight-style: pygments
#bibliography: references.bib
format:
  html: 
    code-fold: true
    code-tools: true
    html-math-method: katex
    self-contained: true                      # html is self-contained
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
  docx: default                               
---

## Preliminaries

This analysis report relates to SAPI V1 for the Analysis of bodyfat example (simplified).


## Load libraries

```{r}
library(mfp2)

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyselect)
library(gt)
library(gtExtras)
library(patchwork)
library(Hmisc)
library(sjlabelled)
```

# Model estimation

## Load data

```{r}
load("../Datafiles/bodyfatdata.Rdata")
```

## Main analysis

### Description of observational units

```{r}
library(dplyr)
library(table1)


# Assume these are your two existing datasets
# data_women <- ...
# data_men <- ...

# Add Gender variable to each
data_women <- women %>% mutate(Gender = "1-Women")
data_men   <- men %>% mutate(Gender = "2-Men")

# Combine them into one dataset
data <- bind_rows(data_women, data_men)


my.render.cont <- function(x) {
        sprintf("%0.1f (%0.1f)\n[%0.1f – %0.1f]", 
               mean(x, na.rm=TRUE), sd(x, na.rm=TRUE), min(x, na.rm=TRUE), max(x, na.rm=TRUE))
}

# Label variables
label(data$Fat) <- "Body Fat (%)"
label(data$Weight) <- "Weight (kg)"
label(data$Height) <- "Height (m)"
label(data$BMI) <- "BMI (kg/m^2)"
label(data$Age) <- "Age (years)"
label(data$Neck) <- "Neck (cm)"
label(data$Chest) <- "Chest (cm)"
label(data$Calf) <- "Calf (cm)"
label(data$Biceps) <- "Biceps (cm)"
label(data$Hips) <- "Hips (cm)"
label(data$Waist) <- "Waist (cm)"
label(data$Forearm) <- "Forearm (cm)"
label(data$PThigh) <- "Proximal Thigh (cm)"
label(data$MThigh) <- "Middle Thigh (cm)"
label(data$DThigh) <- "Distal Thigh (cm)"
label(data$Wrist) <- "Wrist (cm)"
label(data$Knee) <- "Knee (cm)"
label(data$Elbow) <- "Elbow (cm)"
label(data$Ankle) <- "Ankle (cm)"

# Generate Table 1


table1(~ Fat + Weight + Height + BMI + Age + Neck + Chest  + Biceps + 
         Hips + Waist + Forearm +  MThigh  + Wrist + Knee + 
          Ankle | Gender, data = data, render.continuous = my.render.cont)
```

### Main data analysis

#### Model estimation

```{r, message=TRUE}
library(glmnet)

predictors <- c("Weight", "Height", "BMI", "Age",
  "Neck", "Chest",  "Biceps", "Hips", "Waist",
  "Forearm",  "MThigh",  "Wrist",
  "Knee", "Ankle")
p <- length(predictors)

x <- as.matrix(women[, predictors])
y <- as.numeric(women$Fat)
n <- nrow(x)

# fit the coefficient path
fit_glmnet <- glmnet(x=x, y=y, family="gaussian", relax=T)

# plot the coefficient path vs. log(lambda), y axis limited as intercept breaks out
plot(fit_glmnet$relaxed, xvar="lambda", ylim=c(-2,2), xlim=c(-2,2), label=TRUE)


# find lambda where p_active=5


# Get coefficient matrix (excluding intercept)
coef_mat <- coef(fit_glmnet)[-1, ]  # remove intercept row
# Each column corresponds to a different lambda

# Count how many non-zero coefficients for each lambda
num_nonzero <- apply(coef_mat, 2, function(v) sum(v != 0))

# Find the largest lambda where the number of nonzero coefficients equals p
p <- 5
target_lambda_index <- max(which(num_nonzero == p))
target_s <- max(fit_glmnet$lambda[num_nonzero ==p])

# Extract names of active predictors at that lambda
active_coef <- coef_mat[, target_lambda_index]
active_predictors <- rownames(coef_mat)[which(active_coef != 0)]


## find extended set of active predictors (5+3)
p_ext <- 8
target_lambda_index_ext <- max(which(num_nonzero == p_ext))

# Extract names of active predictors at that lambda
active_coef_ext <- coef_mat[, target_lambda_index_ext]
active_predictors_ext <- rownames(coef_mat)[which(active_coef_ext != 0)]




# Result: character vector of predictor names
cat("\n\nSelected predictors:\n")
active_predictors


cat("\n\nExtended set of selected predictors:\n")
active_predictors_ext



#se_residual
se_residual <- ((1-fit_glmnet$relaxed$dev.ratio)*fit_glmnet$relaxed$nulldev)/(n-fit_glmnet$relaxed$df-1)
# at selected lambda
#se_residual[num_nonzero == p]
cat("\n\nResidual standard error (model based):\n")
se_residual[target_lambda_index]

cat("\n\nModel coefficients:\n")
coef(fit_glmnet$relax, s=target_s)
# predict(fit_glmnet$relax, s=target_s, newx=x)


```

### Residual analysis

```{}
```

```{r}
library(ggplot2)

predictions <- predict(fit_glmnet$relaxed, s=target_s, newx=x)

residuals <- y - predictions

ggplot(data=data.frame(residuals), aes(x=residuals)) + geom_histogram()

sqrt_abs_res <- sqrt(abs(residuals))

ggplot(data=data.frame(residuals, predictions, y), aes(x=predictions, y=residuals)) + geom_point() + geom_smooth()

ggplot(data=data.frame(sqrt_abs_res, predictions, y), aes(x=predictions, y=sqrt_abs_res)) + geom_point() + geom_smooth()
```


In the following residual plots, the nearly-selected predictors are denoted by red dots, the selected predictors by blue dots.


```{r}

for(v in active_predictors_ext){
  cols=ifelse(v %in% active_predictors, "blue", "red")
  print(ggplot(data=data.frame(residuals, variable=x[,v]), aes(x=variable, y=residuals), xlab(v)) + geom_point(color=cols) + geom_smooth() + xlab(v)) 
  
}
```



