---
title: "Re-Analysis (simplified): bodyfat prediction"
author:
  - name: Mariana Nold
    affiliation: University of Jena, Institute of Sociology
date: last-modified                            # "`r Sys.Date()`" for system date
categories: [2024, Contrast-Bayes-Frequentist, STRATOS, R]
description: "Classification: confidential"
editor: visual
execute:
  warning: false
theme: cosmo                                   # cosmo bootswatch theme
toc: true                                      # table of contents  
number-sections: true
colorlinks: true                               # ??
highlight-style: pygments
#bibliography: references.bib
format:
  html: 
    code-fold: true
    code-tools: true
    html-math-method: katex
    self-contained: true                      # html is self-contained
    code-line-numbers: true
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
  docx: default                               
---

## Preliminaries

This analysis report relates to SRe-API for the Analysis of bodyfat example (simplified).

## Load libraries

```{r load package, message = FALSE, warning = FALSE}
library("rstanarm")
library("projpred")
library("ggplot2")
library("loo")
library("bayesplot")
library("bayesrules")
library("DHARMa")
library("ggplot2")
library("dplyr")
library("tidyr")
library("patchwork")
library("posterior")
library("knitr")
library("kableExtra")
```

# Model estimation

## Load data

```{r define functions, include = FALSE}

visualize_variable_selection <- function(vs_cvf) {

  # Extract the summary and store as return object
  model_summary <- summary(vs_cvf, deltas = FALSE, cumulate = FALSE)

  # Extract predictor ranking
  rk <- ranking(vs_cvf)

  # Compute variable ranking proportions
  pr_rk <- cv_proportions(rk)
  pr_rk2 <- cv_proportions(rk, cumulate = TRUE)
  # Generate plots
  gg_pr_rk <- plot(pr_rk) + theme(legend.position = "none")  
  gg_pr_rk2 <- plot(pr_rk2) + theme(legend.position = "none")
  combined_plot <- gg_pr_rk / gg_pr_rk2 


  # Return both the summary and the plot
  return(list(summary = model_summary, plot = combined_plot))
}

#' Extract and round a single model size row (e.g. "<=5") from cv_proportions
#'
#' @param cumprop Cumulative proportions matrix from cv_proportions(..., cumulate = TRUE)
#' @param size Integer value, the exact model size to extract (default: 5)
#' @param digits Number of decimal places to round to (default: 3)
#' @return A 1-row matrix with rounded values
get_cumprop_row_rounded <- function(cumprop, size = 5, digits = 3) {
  rowname <- paste0("<=", size)
  if (!(rowname %in% rownames(cumprop))) {
    stop(paste0("Model size '<=", size, "' not found in row names."))
  }
  row <- cumprop[rowname, , drop = FALSE]
  row[] <- round(row, digits = digits)
  return(row)
}


#' Inspect and summarize variable selection results
#'
#' This function provides a structured summary of variable selection output from 
#' both Lasso and forward selection methods (e.g. via projpred). It visualizes 
#' selection stability, shows cumulative selection proportions, and identifies 
#' top predictors based on ranking and suggested model size.
#'
#' @param vs_object A variable selection object (e.g. result from cv_varsel())
#' @param top_n Number of top variables to display (default is 5)
#' @return A list containing summary, plot, ranking, top variables, and suggested terms

inspect_variable_selection <- function(vs_object, top_n = 5) {
  # Run visualization and show textual summary
  cat("\n Summary of variable selection:\n")
  result <- visualize_variable_selection(vs_object)
  print(result$summary)
  
  # Plot selection trace
  cat("\n️ Visualization of selection process:\n")
  print(result$plot)
  
  # Cumulative selection proportions for top_n terms
  cat(paste0("\n Cumulative selection proportions (Top ", top_n, "):\n"))
  cumprop <- cv_proportions(ranking(vs_object), cumulate = TRUE)
  rounded <- get_cumprop_row_rounded(cumprop, size = top_n)
  print(rounded)
  
  # Suggested model size based on selection criteria
  suggested <- suggest_size(vs_object)
  cat("\n Suggested model size:\n")
  print(suggested)
  
  # Full ranking table of all variables
  ranking_data <- ranking(vs_object)$fulldata
  cat("\n Full ranking table:\n")
  print(ranking_data)
  
  # Final selected terms based on suggested model size
  cat(paste0("\n Terms selected (suggested size = ", suggested, "):\n"))
  final_terms <- head(ranking_data, suggested)
  print(final_terms)
  
  # Display top_n terms (fixed number)
  cat(paste0("\n Top ", top_n, " ranked terms:\n"))
  top_terms <- head(ranking_data, top_n)
  print(top_terms)
  
  # Return all outputs invisibly for further use
  invisible(list(
    result = result,
    cumprop = rounded,
    suggested_size = suggested,
    full_ranking = ranking_data,
    final_terms = final_terms,
    top_terms = top_terms
  ))
}

# Computes Bayesian R² as defined by Gelman et al. (2019).
# Measures the proportion of variance in the predictions (ypred) relative to total variance (predictions + residuals).
# The function returns a vector of R² values across posterior draws.
# Input: 
#   - ypred: matrix of posterior linear predictions (rows = draws, columns = observations)
#   - y: vector of observed outcome values
# Output:
#   - r2: numeric vector of Bayesian R² values (one per draw)

bayes_r2 <- function(ypred, y) {
  resid <- sweep(ypred, 2, y)
  var_ypred <- apply(ypred, 1, var)
  var_resid <- apply(resid, 1, var)
  r2 <- var_ypred / (var_ypred + var_resid)
  return(r2)
}

#r2_val <- 1 - sum((ymen - ypred_men)^2)/sum((ymen - ymean_men)^2)

bayes_r2_ext <- function(ypred, y) {
  # compute mean
   ymean_men <- mean(y)

  # R² for each row
   r2_val <- apply(ypred, 1, function(pred_row) {
  1 - sum((ymen - pred_row)^2) / sum((ymen - ymean_men)^2)
})

  return(r2_val)
}
```

```{r  load data and set path}
load("../Datafiles/bodyfatdata.Rdata")
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task Bodyfat/"
```

## Main data analysis

### Define and fit reference model

```{r define and fit refernce model, results='hide', cache=TRUE}

# Create the vector of predictor names
predictors <- c("Weight", "Height", "BMI", "Age",
  "Neck", "Chest", "Biceps", "Hips", "Waist",
  "Forearm", "MThigh", "Wrist",
  "Knee", "Ankle")

# Collapse the predictor names into a single string joined by +
rhs <- paste(predictors, collapse = " + ")

# Combine with the response variable to create a formula
rmf <- as.formula(paste("Fat ~", rhs))

# Print it to check
rmf

# Fit a Bayesian Gaussian regression model using stan_glm 
ref_fit <- stan_glm(rmf, data = women, family = gaussian())
```

```{r MCMC diagnostic and prior summary}
draws <- as.matrix(ref_fit)
# This function returns Rhat and n_eff diagnostics for each model parameter.
# Use it to evaluate convergence and sampling efficiency in Bayesian regression.
# Recommended threshold: Rhat < 1.01 and n_eff_bulk and n_eff_tail > 100 per chain for # reliable inference. see: https://mc-stan.org/rstan/reference/Rhat.html


diagnostics <- summarise_draws(draws)
diagnostics_tbl <- diagnostics[, c("variable", "rhat", "ess_bulk" , "ess_tail")]

kable(diagnostics_tbl, digits = 3, caption = "MCMC diagnostics") %>%
  kable_styling(full_width = FALSE)


prior_summary(ref_fit) # https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html
```

```{r summarize estimates of refernce model}
mat_ref_fit <- as.matrix(ref_fit)
draws_ref_fit <- posterior::as_draws_matrix(mat_ref_fit)
mcmc_areas(mat_ref_fit[,2:13])
summary_tbl_ref_fit <- summarise_draws(draws_ref_fit, mean, sd, median, ~quantile(.x, probs = c(0.05, 0.95)))
kable(summary_tbl_ref_fit, digits = 3, caption = "Posterior summary statistics")
```

Investigate quality of the reference model:

```{r LOOPIT and bayesrules summary, cache=TRUE}

pp_check(ref_fit)
y_rep_rmf <- posterior_predict(ref_fit)
loo_rmf <- loo(ref_fit , save_psis = TRUE, cores = 4)
psis1_rmf <- loo_rmf$psis_object
lw_rmf <- weights(psis1_rmf)
color_scheme_set("red")

ppc_loo_pit_overlay(y = women$Fat,yrep=y_rep_rmf,lw=lw_rmf)

# Summarize the posterior distributions simple model
summary_rmf_cv <- prediction_summary_cv(model = ref_fit , data = women, k = 3)

# Print the summary simple model
print(summary_rmf_cv)
```

The reference model passes the checks and therefore I decide to use this model, because it predicts well. In a next step I use the `projpred`-algorithms to find a simpler model that gives similar predictions as the reference model. In `projpred`, variable selection is performed by evaluating submodels of increasing size, starting from zero predictors up to a predefined maximum (`nterms_max`) in case of of **forward excursion.** When fixing the number of variables (which is not possible for L1 search), the best model of that size is chosen by projecting the reference model onto all candidate submodels and selecting the one with minimal expected Kullback-Leibler divergence (KLD)---this ensures the reduced model preserves the predictive distribution of the full model as closely as possible.

In the case of **L1-regularized search**, coefficients are estimated over a grid of penalization values using the reference model's fit in place of the observed response. This produces a sequence of sparse models ranked by the order in which variables enter with non-zero coefficients as the penalty decreases. While computationally efficient, L1 search does not support fixing model size in advance, making it less flexible than forward search in that regard. Importantly, penalization is only used for variable ranking; submodels are later refit using projections without penalization.

To evaluate predictor stability, the search procedure is repeated across data subsets. **In this re-analysis,** **leave-one-out cross-validation (LOO)** **is used in combination with** **L1 search** **as the primary strategy (`LOO-L1`).** Additionally, 5-fold cross-validation is implemented to assess the **robustness** of variable selection and predictive performance. In cross-validation mode, selection is performed on training folds, and performance is evaluated on held-out test data. Results are also computed based on forward excursion, in combination with LOO-CV and 5-fold-CV.

Finally, output visualizations report the proportion of times each variable was selected. If `cumulate = FALSE`, proportions reflect selection at specific ranking positions. If `cumulate = TRUE`, they capture cumulative inclusion up to each rank, as annotated in the "CV ranking proportions matrix" plot.

Before inspecting the results, the primary search strategy was set to LOO-L1 (leave-one-out cross-validation with L1-regularized search), as this most closely aligns with the original SAP. To assess the robustness of predictor selection, forward selection and 5-fold cross-validation were additionally computed, not as alternatives, but to evaluate whether they yield different sets of selected variables, thus to access the internal robustness.

**Predictive Projections: Forward search, LOO (internal robustness)**

```{r predictive projection, results='hide', cache=TRUE}
set.seed(2341)
vs_cvf <- cv_varsel(object = ref_fit, method = 'forward', cv_method = 'loo', validate_search = TRUE, verbose = FALSE)

save(vs_cvf,file = file.path(PPath,"vs_cvf"))
#load(file = file.path(PPath,"vs_cvf"))

plot(vs_cvf, deltas = TRUE, stats =  c('elpd'))
```

**Predictive Projection: Forward search, five-fold (internal robustness)**

```{r 5-fold cross validation, results='hide', cache=TRUE}
set.seed(2341)
vs_cvf5 <- cv_varsel(object = ref_fit,  method = "forward",cv_method = "kfold",      
  K = 5,validate_search = TRUE, verbose = FALSE)

save(vs_cvf5,file = file.path(PPath,"vs_cvs"))
#load(file = file.path(PPath,"vs_cvs"))

plot(vs_cvf5, deltas = TRUE, stats =  c('elpd'))
```

**Predictive Projections: L1 search,LOO (key results)**

```{r predictive lasso projection, results='hide', cache=TRUE}
# Generate model path: By varying lambda, a sequence of submodels with an increasing number of (non-zero) coefficients is created. ➤ The order in which coefficients “appear” serves as a ranking of the variables.

set.seed(2341)
vs_cvl <- cv_varsel(object = ref_fit, method = 'L1',  cv_method = 'loo', validate_search = TRUE, verbose = FALSE)

save(vs_cvl,file = file.path(PPath,"vs_cvl"))
#load(file = file.path(PPath,"vs_cvl"))

plot(vs_cvl, deltas = TRUE, stats =  c('elpd'))
```

**Predictive Projections: L1 search, 5-fold (internal robustness)**

```{r predictive lasso projection k-fold, results='hide', cache=TRUE}
set.seed(2341)
vs_cvl5 <- cv_varsel( object = ref_fit, method = 'L1',
  cv_method = "kfold",K = 5, verbose = FALSE)

save(vs_cvl5,file = file.path(PPath,"vs_cvl5"))
#load(file = file.path(PPath,"vs_cvl5"))

# Plot the variable selection 

plot(vs_cvl5, deltas = TRUE, stats =  c('elpd'))
```

#### Model stability: Stability of predictive projection

Forward Search (LOO)

```{r forward selection LOO results}
inspect_variable_selection(vs_cvf, top_n = 5)
```

Forward Search (5-fold)

```{r forward selection 5-fold results}
inspect_variable_selection(vs_cvf5, top_n = 5)
```

L1 Search (LOO)

```{r lasso selection LOO results}
key_results <- inspect_variable_selection(vs_cvl, top_n = 5)
solterms_final_fivel <- key_results$top_terms
```

L1 Search (5-fold)

```{r lasso selection 5-fold results}
inspect_variable_selection(vs_cvl5, top_n = 5)
```

# **Compute Projected Posterior: Post-selection-inference for L1 Search (LOO)**

Here post-selection inference starts, thus the results retain the uncertainty from the reference model posterior.

```{r post-selection inference, results='hide'}
# https://cran.r-project.org/web/packages/projpred/vignettes/projpred.html#post-selection-inference
proj_ref_fit <- project(vs_cvl, predictor_terms = solterms_final_fivel, ns = 4000)
```

#### Model description

```{r model summary table}

# assemble data
results_beta <- matrix(0, nrow = 6 , ncol  = 5)
rownames(results_beta) <- c("(Intercept)",solterms_final_fivel)
colnames(results_beta) <- c("Regression coefficient", "Standardized beta", "Cum. proportions <=5", "5% PD-quantile", "95% PD-quantile")

mat <- as.matrix(proj_ref_fit)
beta <- mat[, 1:(ncol(mat) - 1)]

# Transform matrix to draws format (optional, for posterior-friendly summaries)
draws1 <- posterior::as_draws_matrix(beta)

summary_tbl <- summarise_draws(draws1, mean,  median, ~ quantile(.x, probs = c(0.05, 0.95)))
# Compute standard deviations (excluding the first two columns) for women dataset
sd_women <- apply(women[, -c(1,2)], 2, sd)
vars_to_select <- c("BMI", "Waist", "Hips", "Knee", "Biceps")
sd_selected <- sd_women[intersect(vars_to_select, names(sd_women))]

std_coef <- summary_tbl$mean[2:6] * (sd_selected)
std_coef <- c(summary_tbl$mean[1],std_coef )

results_beta[,1] <- round(summary_tbl$mean,3)
results_beta[,2] <- round(std_coef,3)
results_beta[,3] <- round(c(1.0,key_results$cumprop[1:5]),3)
results_beta[,4] <- round(summary_tbl$`5%`,3)
results_beta[,5] <- round(summary_tbl$`95%`,3)


results_beta<-results_beta[order(-abs(results_beta[,3])),]
kable(results_beta, format="html", caption="Regression coefficients and stability measures")
```

Global model descriptors:

```{r R-squared and Sigma}
y_linpred_female <- proj_linpred(proj_ref_fit, newdata = women)
women2 <- cbind(women, linpred = colMeans(y_linpred_female$pred))
R.square <- (cor(women2$linpred, women2$Fat, method = "pearson"))^2
#R.square

draws1b <- posterior::as_draws_matrix(mat)

summary_tbl_1b <- summarise_draws(draws1b, mean,  median, ~ quantile(.x, probs = c(0.05, 0.95)))
summary_sigma <- summary_tbl_1b %>% slice(7)
sigma_mean <- summary_sigma %>% pull(mean)

global_desc <- c(residual_error = round(sigma_mean,3) , Rsq = round(R.square,3))
names(global_desc) <- c("Residual error",  "R-squared ")

kable(global_desc, caption="Global model descriptors")
kable(summary_sigma , digits = 3, caption = "Residual Error ") %>%
  kable_styling(full_width = FALSE)

r2_values  <- bayes_r2(ypred = y_linpred_female$pred, y = women2$Fat)

summary_r2 <- data.frame(
  mean   = round(mean(r2_values),3),
  median = round(median(r2_values),3),
  p05   = round(quantile(r2_values, 0.05),3),
  p95   = round(quantile(r2_values, 0.95),3)
)
#colnames(summary_r2)
rownames(summary_r2) <- "R²"
colnames(summary_r2)[colnames(summary_r2) == "p05" ] <- "5%"
colnames(summary_r2)[colnames(summary_r2) ==  "p95" ] <- "95%"

kable(summary_r2 ,digits = 3 , caption = "R² as defined by Gelman et al. (2019)") %>%
  kable_styling(full_width = FALSE)
```

#### Assumptions and diagnostics: residual analysis

```{r Extended set of active predictors}
print("Five active predictors")
active_predictors <- solterms_final_fivel
active_predictors
print("Eight ext. active predictors")
active_predictors_ext <- key_results$full_ranking[1:8]
active_predictors_ext

p <- length(predictors)

x <- as.matrix(women[, predictors])
y <- as.numeric(women$Fat)
n <- nrow(x)
```

```{r residual diagnostics I}
library(ggplot2)

predictions <- women2$linpred
y <- women2$Fat
# Compute residuals
prj_predictW <- proj_predict(proj_ref_fit)

qres.m1 = createDHARMa(simulatedResponse = t(prj_predictW), 
                          observedResponse = women2$Fat, 
                          fittedPredictedResponse = apply(prj_predictW, 2, median), integerResponse = F, seed = 123, method = "PIT")
# uniform distributed residuals
print("Uniform residuals")
summary(residuals(qres.m1))
# transform to normal scale
residuals <- qnorm(residuals(qres.m1))
print("Normal residuals")
summary(residuals)


ggplot(data=data.frame(residuals), aes(x=residuals)) + geom_histogram() + ggtitle(label="Histogram of residuals")

sqrt_abs_res <- sqrt(abs(residuals))

ggplot(data=data.frame(residuals, predictions, y), aes(x=predictions, y=residuals)) + geom_point() + geom_smooth() + ggtitle(label="Residals vs. predictions")

ggplot(data=data.frame(sqrt_abs_res, predictions, y), aes(x=predictions, y=sqrt_abs_res)) + geom_point() + geom_smooth() + ggtitle(label="Assessment of heteroskedasticity: square root of absolute residuals vs. predictions")
```

In the following residual plots, the nearly-selected predictors are denoted by red dots, the selected predictors by blue dots.

```{r residual diagnostics II}

for(v in active_predictors_ext){
  cols=ifelse(v %in% active_predictors, "blue", "red")
  print(ggplot(data=data.frame(residuals, variable=x[,v]), aes(x=variable, y=residuals), xlab(v)) + geom_point(color=cols) + geom_smooth() + xlab(v) + ggtitle(label=paste("Residuals vs ",v, collapse=""))) 
  }
```

### Objective 2: Model validation

```{r Overall performance measures}
# Computing predictions in the men's data set

xmen <- as.matrix(men[, predictors])
ymen <- as.numeric(men$Fat)
ymean_men <- mean(ymen)
nmen <- nrow(xmen)

y_linpred_male <- proj_linpred(proj_ref_fit,  newdata = men)
ypred_men <- colMeans(y_linpred_male$pred)
res_men <- ymen - ypred_men

r2_val <- 1 - sum((ymen - ypred_men)^2)/sum((ymen - ymean_men)^2)
R.squareM <- (cor(ypred_men,ymen, method = "pearson"))^2



rmspe <- sqrt(mean(res_men^2))

glob_pm <- c(r2_val, rmspe,R.squareM)
names(glob_pm) <- c("Validation R-squared", "Root mean squared prediction error", "R-squared")

kable(glob_pm, caption="Overall performance measures")

r2_values_men  <- bayes_r2(ypred = y_linpred_male$pred, y = men$Fat)

summary_r2_men  <- data.frame(
  mean   = round(mean(r2_values_men),3),
  median = round(median(r2_values_men),3),
  p05   = round(quantile(r2_values_men, 0.05),3),
  p95   = round(quantile(r2_values_men, 0.95),3)
)

rownames(summary_r2_men) <- "R²"
colnames(summary_r2_men)[colnames(summary_r2_men) == "p05" ] <- "5%"
colnames(summary_r2_men)[colnames(summary_r2_men) ==  "p95" ] <- "95%"

kable(summary_r2_men  ,digits = 3 , caption = "R² as defined by Gelman et al. (2019) for men") %>%
  kable_styling(full_width = FALSE)


r2_values_men_ext  <- bayes_r2_ext(ypred = y_linpred_male$pred, y = men$Fat)


summary_r2_men_ext  <- data.frame(
  mean   = round(mean(r2_values_men_ext),3),
  median = round(median(r2_values_men_ext),3),
  p05   = round(quantile(r2_values_men_ext, 0.05),3),
  p95   = round(quantile(r2_values_men_ext, 0.95),3)
)

rownames(summary_r2_men_ext) <- "R²"
colnames(summary_r2_men_ext)[colnames(summary_r2_men_ext) == "p05" ] <- "5%"
colnames(summary_r2_men_ext)[colnames(summary_r2_men_ext) ==  "p95" ] <- "95%"

kable(summary_r2_men_ext  ,digits = 3 , caption = "Validation R²  for men") %>%
  kable_styling(full_width = FALSE)
```

Calibration plot:

```{r Calibration plot}
# Calibration plot

ggplot(data.frame(ypred_men, ymen), aes(x=ypred_men, y=ymen)) + geom_point() + geom_smooth(col="red") + geom_abline(intercept=0, slope=1, linetype="dashed")
```

Calibration model:

```{r compute Calibration intercept and slope (external), results='hide'}
observed <- men$Fat
predicted <- ypred_men
calib_data <- data.frame(observed = observed, predicted = predicted)
# Compute calibration intercept and slope
calibration_model <- stan_glm(observed ~ predicted, family = gaussian(), data = calib_data)
calibration_model0 <- stan_glm(observed ~ offset(predicted), family = gaussian(), data = calib_data)
```

```{r Calibration intercept and slope (external)}
summary(calibration_model)
summary(calibration_model0)

posterior_samples <- as.matrix(calibration_model)
posterior_samples0 <- as.matrix(calibration_model0)
intercepts <- posterior_samples0[, "(Intercept)"]
slopes <- posterior_samples[, "predicted"]

cal_slope <- mean(slopes) 
conf_slope <- quantile(posterior_samples[, "predicted"], probs = c(0.025,  0.975))
cal_int <- mean(intercepts)
conf_int <- quantile(posterior_samples0[, "(Intercept)"], probs = c(0.025,  0.975))

cal_stats <- c(cal_int, conf_int, cal_slope, conf_slope)

names(cal_stats) <- c("Calibration intercept", "Calibration intercept, lower 95% PD", "Calibration intercept, upper 95% PD", "Calibration slope", "Calibration slope, lower 95% PD", "Calibration slope, upper 95% PD")

kable(cal_stats, caption="Calibration statistics")




# Dataframe  for regression lines
regression_lines <- data.frame(
  intercept = intercepts,
  slope = slopes
)

n_lines <- nrow(as.matrix(calibration_model))
print(n_lines)
sampled_lines <- regression_lines[sample(nrow(regression_lines), size = 100), ]
# plot
ggplot(calib_data, aes(x = predicted, y = observed)) +
  geom_point(alpha = 0.9) +
  geom_abline(data = sampled_lines, aes(intercept = intercept, slope = slope), alpha = 0.1, color = "blue") +
  labs(title = "Calibration Plot with regression lines",
       x = "Predicted",
       y = "Observed") +
  theme_minimal()
color_scheme_set("orange")


# Intercept
mcmc_areas(
  as.matrix(calibration_model0),
  pars = "(Intercept)",
  prob = 0.95
) +
  ggtitle("Calibration Intercept (external)")

# Slope
mcmc_areas(
  as.matrix(calibration_model),
  pars = "predicted",
  prob = 0.95
) +
  ggtitle("Calibration Slope (external)")



mcmc_areas(as.matrix(calibration_model0),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model for intercept (external)")

mcmc_areas(as.matrix(calibration_model),
           pars = "sigma",
           prob = 0.95) + 
  ggtitle("Sigma of Calibration model for slope (external)")

r2_values_men_extK  <- bayes_r2_ext(ypred = y_linpred_male$pred + intercepts, y = men$Fat)


summary_r2_men_extK  <- data.frame(
  mean   = round(mean(r2_values_men_extK),3),
  median = round(median(r2_values_men_extK),3),
  p05   = round(quantile(r2_values_men_extK, 0.05),3),
  p95   = round(quantile(r2_values_men_extK, 0.95),3)
)

rownames(summary_r2_men_extK) <- "R²"
colnames(summary_r2_men_extK)[colnames(summary_r2_men_extK) == "p05" ] <- "5%"
colnames(summary_r2_men_extK)[colnames(summary_r2_men_extK) ==  "p95" ] <- "95%"

kable(summary_r2_men_extK, digits = 3 , caption = "Cor. Validation R²  for men") %>%
  kable_styling(full_width = FALSE)

```

The calibration slope is approximately 1.0, exactly `r cal_slope`, suggesting good predictive scaling. However, the intercept indicates a systematic overestimation of body fat in men by `r -cal_int` percentage points.

To address this bias, a re-calibrated model adjusting the intercept by `r cal_int`. could be proposed. This adjusted model yields an apparent ( R\^2 ) of `r summary_r2_men_extK$mean`. The uncertainty of the ( R\^2 )estimate is expressed through a posterior predictive interval, reflecting the range of plausible values based on the model and the observed data.
