---
title: "School belonging data analysis, logistic model, Bayes"
author: 
  - Mariana Nold
date: now
date-format: iso
title-block-banner: "#828282"
format:
  html:
    code-fold: true
    code-tools: true
    theme: default
    toc: true
    number-sections: true
    toc-location: right
    embed-resources: true
    self-contained: true
    df-print: kable
params: 
  folder_analysis: Data/Processed/School_FSAP
---

# Background 

This Bayesian analysis for the school belonging analysis task follows the 
statistical analysis plan by Mariana Nold dated 5.2025 (and referring
to the task description for the simplified school belonging task with the same 
date that addresses a binary outcome variable). For a very brief initial data
analysis we refer to `analysis_school-logistic-ida`.

```{r load package, message = FALSE, warning = FALSE}
#| code-fold: true
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task_Truancy"
source(file.path(PPath,"RFiles/packages.r"))
# Load original PISA data, single imputed data set, to avoid missing data as statistical issue here
load(file = file.path(PPath,"Orgdata/IMPDAT_PISA.RData"))
# Prepare data, thus build PISA scales 
# Pole so that high values indicate a higher sense of belonging based on PIM
source(file.path(PPath,"RFiles/01_datapreperation.r"))
to.rm <- objects()[!objects() %in% c("PPath","pisa18")]
# There should be at least ten students per school. Schools with fewer observations are excluded from the analysis. 
pisa18$school.id <- as.factor(pisa18$school.id)
help <- split(pisa18,pisa18$school.id)
nst <- lapply(help, function(x) length(x[[1]]))
tf <- lapply(nst, function(x) x[[1]] > 9)
utf <- unlist(tf)
large_school <- help[utf]
pisa2018 <- list.rbind(large_school)
pisa2018$school.id <- droplevels(pisa2018$school.id)
#table(pisa2018$school.id)

rm(pisa18,tf,nst,help,large_school,utf)

save(pisa2018,file = file.path(PPath,"Files/pisa2018.RData")) 

```
## Prepare focal variables

```{r  define second focal variable and outcome}
#| code-fold: true
# Compute quartiles of being bullied at school level
pisa2018$ATT4 <- cut_number(pisa2018$ATT01,4)
levels(pisa2018$ATT4)
levels(pisa2018$ATT4) <- 1:4
threshold <- quantile(pisa2018$belong, 0.25)
threshold
# Focal variable is 1, if student has a low sense of school-belonging
pisa2018$bel01 <- ifelse(pisa2018$belong <= threshold, 1, 0)
#print("Focal variables")
#print("ATT4")
#table(pisa2018$ATT4)
#print("Bel01")
#table(pisa2018$bel01)
print("Joint distribution of ATT4 and Bel01")

library(janitor)

pisa2018 %>%
  tabyl(ATT4, bel01) %>%
  adorn_totals(where = c("row", "col"))

```



## Helper functions

Code for helper functions.

```{r helper functions}
#| code-fold: true

source(file.path(PPath,"RFiles/02_functions_KD.R"))
```

## Variables

As outlined in Table 1 in the task description. 

```{r non-focal variables}

VARIABLES = list(
    outcome = "bel01", 
    focal = c("bull", "ATT4"), 
    nonfocal = c("female", "nld", "aca", "val", "nfof", "native", "joyread", 
                 "goal", "mot", "res", "ndis_clim", "GYM", "UNI", "NDISH", "N_SC")
)

glue("{length(VARIABLES$nonfocal)} non-focal variables used for modeling according to SAP: 
     {VARIABLES$nonfocal |> paste0(collapse = ', ')}")
```

# Analysis

## Define model two models, first full model, second background model

```{r  define model,results='hide', message=FALSE, warning=FALSE}
models<-list()
# Full model, variable selection based on PIM  including focal variables
models[[1]] <- as.formula(bel01 ~ bull*ATT4 + female + nld +  aca  + val  + nfof  + native+ joyread + goal + mot + res +  ndis_clim    + GYM + UNI  +  NDISH + N_SC + (1 | school.id))
# Background model, variable selection based on PIM without focal variables
models[[2]] <- as.formula(bel01 ~ female + nld +  aca  + val  + nfof  + native + joyread + goal + mot + res +  ndis_clim    + GYM + UNI  + NDISH + N_SC + (1 | school.id))
```


## Check the convergence criteria 

```{r fit modesl and do MCMC diagnostic,results='hide', message = FALSE, warning = FALSE, cache=TRUE}
a.seed <- 12345 
a.iter <-  2000 
a.chains <- 4  

fits <- list()
# full model
fits[[1]] <- stan_glmer(models[[1]], data = pisa2018, seed = a.seed,
                      iter = a.iter, chains = a.chains,family = binomial()) 
# background model
fits[[2]] <- stan_glmer(models[[2]], data = pisa2018, seed = a.seed,
                      iter = a.iter, chains = a.chains,family = binomial()) 

res.path <- file.path(PPath,"Result")
save(fits, file = file.path(res.path,"fits.RData"))
summary_list <- lapply(fits, summary) # mcse Rhat n_eff
# launch_shinystan(fits[[1]])
# launch_shinystan(fits[[2]])
```
## Summary of results
```{r Summary of results}

# Calculate summaries for all models in fits
summaryFixed <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "fixed"))
summarySchools <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "ran_vals"))
summaryVar <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "ran_pars"))
summaryPrior <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "ran_pars"))



# Fixed effects

print("## Fixed Effects for Model")
summaryFixed[[1]]
summaryFixed[[2]]

# Optional: Random effects (school-level)
#print("## Random Effects (Schools) for Model")
#summarySchools[[1]]
#summarySchools[[2]]

# Random effect variances
print("## Random Effect Variances for Model")
summaryVar[[1]]
summaryVar[[2]]

prior_summaries <- lapply(fits, prior_summary)
print(prior_summaries[[1]])
print(prior_summaries[[2]])

# Prior summaries
# show prior_summaries 
#for (i in seq_along(prior_summaries)) {
#  cat("Prior Summary for model", i, ":\n")
#  print(prior_summaries[[i]])
#  cat("\n----------------------\n")
#}
```






## LOOIC

```{r looic, cache = TRUE}
loo_list <- list()
# Compute loo

loo_list <- lapply(fits, loo, cores = 1)
# Save loo

save("loo_list", file = file.path(res.path,"loo_list.RData"))

loo_comparison <- loo_compare(loo_list[[1]], loo_list[[2]])

loo_df <- as.data.frame(loo_comparison)
rownames(loo_df) <- c("full", "bg")

kable(loo_df, digits = 2, caption = "LOO Model Comparison")
```

## Randomized Quantile Residuals

```{r Randomized Quantile Residuals}
#source(file.path(PPath,"RFiles/02_functions_KD.R"))
# Compute posterior predictions
posterior_predictions <- lapply(fits, posterior_predict)

qres.m1 = createDHARMa(simulatedResponse = t(posterior_predictions[[1]]), 
                          observedResponse = pisa2018$bel01, 
                          fittedPredictedResponse = apply(posterior_predictions[[1]], 2, median), integerResponse = T, seed = 123, method = "PIT") 
summary(residuals(qres.m1))
# QQ-Plot 
res <- qnorm(residuals(qres.m1))

# QQ-Plot mit ggplot2
ggplot(data.frame(res), aes(sample = res)) +
  stat_qq(color = "steelblue", size = 2) +
  stat_qq_line(color = "firebrick", linetype = "dashed") +
  labs(
     title = "QQ Plot of Transformed Residuals",
    x = "Theoretical Quantiles",
    y = "Observed Quantiles"
  ) +
  theme_minimal(base_size = 14)

plots1 <- plot_violin_residuals(qres.m1, pisa2018, "Residual Diagnostics Full model")
print(plots1)

qres.m2=createDHARMa(simulatedResponse = t(posterior_predictions[[2]]), 
                          observedResponse = pisa2018$bel01, 
                          fittedPredictedResponse = apply(posterior_predictions[[2]], 2, function(x) ifelse(median(x) == 0.5, 1, median(x))), integerResponse = T, seed = 123, method = "PIT") 
summary(residuals(qres.m2))

# QQ-Plot 
res2 <- qnorm(residuals(qres.m2))

# QQ-Plot mit ggplot2
ggplot(data.frame(res2), aes(sample = res2)) +
  stat_qq(color = "steelblue", size = 2) +
  stat_qq_line(color = "firebrick", linetype = "dashed") +
  labs(
    title = "QQ Plot of Transformed Residuals",
    x = "Theoretical Quantiles",
    y = "Observed Quantiles"
  ) +
  theme_minimal(base_size = 14)

plots2 <- plot_violin_residuals(qres.m2, pisa2018, "Residual Diagnostics Background model")
print(plots2)
```

## PPN based on Kolmogorov distance
```{r KDs}

# Compute Kolmogorov distance
kds <- KD(ypredCore = posterior_predictions[[1]], yrep = posterior_predictions[[2]])



# Tabelle 1: Summary
kable(kds[[1]], digits = 3, caption = "Bull & ATT04")

# Tabelle 2: Details
kable(kds[[2]], digits = 3, caption = "Bull")

# Tabelle 3: Diagnostics
kable(kds[[3]], digits = 3, caption = "ATT04")


```

## Posterior predictive checks
```{r ppcs}
color_scheme_set(scheme = "pink")
ppcModelFull <- visualPPC(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[1]])

ppcModelFull


color_scheme_set(scheme = "blue")
#color_scheme_set(scheme = "gray")
ppcModelBG <- visualPPC(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[2]])

ppcModelBG
```

## Compute Bayesian p-values
```{r Bayesian p value}
# https://search.r-project.org/CRAN/refmans/HuraultMisc/html/post_pred_pval.html
#library("kableExtra")

num <- 3

pppvModelFull <- two.sided.pppv(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[1]], alternative = "two.sided")
#kable(kds[[3]], digits = 3, caption = "ATT04")

kable(matrix(pppvModelFull[[1]],nrow = 2,byrow=TRUE),  digits = 3, caption = "Bull x ATT04")
kable(matrix(pppvModelFull[[2]],nrow = 2,byrow=TRUE),  digits = 3, caption = "Bull")
kable(matrix(pppvModelFull[[3]],nrow = 2,byrow=TRUE),  digits = 3, caption = "ATT04")


pppvModelBG <- two.sided.pppv(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[2]], alternative = "two.sided")


kable(matrix(pppvModelBG[[1]],nrow = 2,byrow=TRUE), digits = 3, caption = "Bull x ATT04")
kable(matrix(pppvModelBG[[2]],nrow = 2,byrow=TRUE), digits = 3, caption = "Bull")
kable(matrix(pppvModelBG[[3]],nrow = 2,byrow=TRUE), digits = 3, caption = "ATT04")

```

## Compute Brier scores

```{r Brier scores of cands}
# Brier Scores berechnen
br <- list()
for (i in 1:2) {
  br[[i]] <- brierS(pisa2018$bel01, colMeans(posterior_predictions[[i]]))
}

# In Data Frame umwandeln
brier_df <- data.frame(
  Model = c("full", "bg"),
  BrierScore = unlist(br)
)

# Ausgabe als Tabelle
kable(brier_df, digits = 3, caption = "Brier Scores for Models full and bg")


```

# Summarizing results based on the posterior distribution of full model

```{r Summarizing results}
# For each quartile compute the median ATT01.
pisa18condPr1 <- pisa2018 %>%
  group_by(ATT4) %>%
  summarise(mATT01 = median(ATT01), sd = sd(ATT01),min = min(ATT01),max = max(ATT01))

meancondPr1 <- pisa18condPr1$mATT01

meancondPr1

squcondPr <- split(pisa2018,f = list(pisa2018$ATT4,pisa2018$bull))


pint_condPr25 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.25)

pint_condPr50 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.5)

pint_condPr75 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.75)


brE <- 4

df_condPr25 <- data.frame(x = rep(round(meancondPr1,2),2),
                          group = c(rep("nbull.",brE),rep("bull.",brE)),
                          mean =  pint_condPr25[,2],low = pint_condPr25[,1], 
                          up = pint_condPr25[,3])

df_condPr50 <- data.frame(x = rep(round(meancondPr1,2),2),
                        group = c(rep("nbull.",brE),rep("bull.",brE)),
                        mean =  pint_condPr50[,2],low = pint_condPr50[,1], 
                        up = pint_condPr50[,3])

df_condPr75 <- data.frame(x = rep(round(meancondPr1,2),2),
                          group = c(rep("nbull.",brE),rep("bull.",brE)),
                          mean =  pint_condPr75[,2],low = pint_condPr75[,1], 
                          up = pint_condPr75[,3])

lowQ <- c(pint_condPr25[,1],pint_condPr50[,1],pint_condPr75[,1])
med <-  c(pint_condPr25[,2],pint_condPr50[,2],pint_condPr75[,2])
upQ <-  c(pint_condPr25[,3],pint_condPr50[,3],pint_condPr75[,3])


df_condPr1 <- data.frame(x = rep(round(meancondPr1,2),6),
                            group = c(rep("bv = 25%, nb.",brE),rep("bv = 25%, b.",brE),
                                  rep("bv = 50%, nb.",brE),rep("bv = 50%, b.",brE),
                                  rep("bv = 75%, nb.",brE),rep("bv = 75%, b.",brE)),
                                  mean =  med,low = lowQ, 
                                   up = upQ)


knitr::kable(df_condPr1, format = "simple", digits = 3)



p_condPr <- plot.result(df_condPr1)


p_condPr
```

