---
title: "School belonging data analysis, short IDA"
author: 
  - Michael Kammer
date: now
date-format: iso
title-block-banner: "#828282"
format:
  html:
    code-fold: true
    code-tools: true
    theme: default
    toc: true
    number-sections: true
    toc-location: right
    embed-resources: true
    self-contained: true
    df-print: kable
params: 
  folder_analysis: Data/Processed/School_FSAP
---

# Background 

This frequentist analysis for the school belonging analysis task follows the 
statistical analysis plan by Michael Kammer dated 30.6.2025 (and referring
to the task description for the simplified school belonging task with the same 
date that addresses a binary outcome variable). Here we present a very short 
initial data analysis common to the Frequentist and Bayesian workflows.

# Setup and data

Data were provided by Mariana Nold by mail on 3.1.2025 (`pisa2018.RData`). This
dataset is already pre-processed with one small exception.

- ATT4 is computed by the `ggplot2::cut_number()` function, as outlined by 
    Mariana in her SAP. 
- bel01 is computed as the lowest 25% of students in terms of sense of belonging
    in `belong`.

```{r message=FALSE, warning=FALSE}
#| code-fold: true

T_START = Sys.time()
SUFFIX = ""

# data handling and printing
library(tidyverse)
library(glue)
library(magrittr)
library(patchwork)
library(gtsummary)
library(gt)

# plotting
library(GGally)

rm(list = c("pisa2018"))
load("Data/Raw/pisa2018.Rdata")

pisa2018 %<>% mutate(
        ATT4 = cut_number(ATT01, 4), 
        bel01 = belong <= quantile(belong, probs = 0.25, na.rm = TRUE)
    )
```

## Helper functions

Code for helper functions.

```{r}
#| code-fold: true

plot_density_and_box <- function(plotdf, v) {
    list(
        ggplot(plotdf, aes(x = .data[[v]])) + 
            geom_density() + 
            theme_bw() +
            theme(axis.title.x = element_blank(), 
                  axis.text.x = element_blank(), 
                  axis.ticks.x = element_blank(),
                  plot.margin = margin()), 
        ggplot(plotdf, aes(x = .data[[v]], y = 0)) + 
            geom_boxplot(outlier.shape = NA) +
            geom_jitter(height = 0.25, width = 0, alpha = 0.1) +
            theme_bw() + 
            theme(axis.text.y = element_blank(), 
                  axis.title.y = element_blank(),
                  axis.ticks.y = element_blank(),
                  panel.grid.minor.y = element_blank(), 
                  panel.grid.major.y = element_blank(), 
                  plot.margin = margin(0))
    ) %>% 
        wrap_plots(nrow = 2, heights = c(0.8, 0.2))
}
```

# Variables

As outlined in Table 1 in the task description. 

```{r}
VARIABLES = list(
    outcome = "bel01", 
    focal = c("bull", "ATT4"), 
    nonfocal = c("female", "nld", "aca", "val", "nfof", "native", "joyread", 
                 "goal", "mot", "res", "ndis_clim", "GYM", "UNI", "NDISH", "N_SC")
)

glue("{length(VARIABLES$nonfocal)} non-focal variables used for modeling according to SAP: 
     {VARIABLES$nonfocal |> paste0(collapse = ', ')}")
```

# IDA

Elementary overview of distributions and correlations. 

```{r}
pisa2018 %>% 
    select(bel01, one_of(unlist(VARIABLES))) %>% 
    tbl_summary(
        type = all_continuous() ~ "continuous2",
        statistic = list(all_continuous() ~ c("{median} ({p25}, {p75})",
                                              "{mean} ({sd})",
                                             "{min}, {max}"))
    )
```

```{r fig.width=8, fig.height=8}
cormat = pisa2018 %>% 
    select(one_of(unlist(VARIABLES))) %>% 
    mutate(across(where(is.factor), as.numeric)) %>% 
    cor(use = "pairwise.complete.obs", method = "spearman")

ggcorrplot::ggcorrplot(cormat)
```

Focal predictors - important for target estimands.

```{r}
tbl_cross(pisa2018, bull, ATT4)
```

# Sessioninfo

To build this notebook:

- install `quarto`
- make sure the data is in the correct folder (either in global Data folder, 
    or as subfolder of the folder containing this notebook).
- build using the `make.R` script (global Data), or using the Render button in 
    RStudio (subfolder Data).

```{r}
params

glue("Runtime {difftime(Sys.time(), T_START, units = 'm') / 60} hours.")

sessioninfo::session_info()
```
