---
title: "School belonging data analysis, logistic model"
author: 
  - Michael Kammer
date: now
date-format: iso
title-block-banner: "#828282"
format:
  html:
    code-fold: true
    code-tools: true
    theme: default
    toc: true
    number-sections: true
    toc-location: right
    embed-resources: true
    self-contained: true
    df-print: kable
params: 
  folder_analysis: Data/Processed/School_FSAP
---

# Background 

This frequentist analysis for the school belonging analysis task follows the 
statistical analysis plan by Michael Kammer dated 20.5.2025 (and referring
to the task description for the simplified school belonging task with the same 
date that addresses a binary outcome variable).

# Setup and data

Data were provided by Mariana Nold by mail on 3.1.2025 (`pisa2018.RData`). This
dataset is already pre-processed with one small exception.

- ATT4 is computed by the `ggplot2::cut_number()` function, as outlined by 
    Mariana in her SAP. 
- bel01 is computed as the lowest 25% of students in terms of sense of belonging
    in `belong`.

```{r message=FALSE, warning=FALSE}
#| code-fold: true

T_START = Sys.time()
SUFFIX = ""

# data handling and printing
library(tidyverse)
library(glue)
library(magrittr)
library(patchwork)
library(gtsummary)
library(gt)

# modelling
library(lme4)
library(MuMIn)
library(DHARMa)

# plotting
library(GGally)

# parallel
library(doFuture)

# required bot not loaded
# ranger
# pracma

rm(list = c("pisa2018"))
load("Data/Raw/pisa2018.Rdata")

pisa2018 %<>% mutate(
        ATT4 = cut_number(ATT01, 4), 
        bel01 = belong <= quantile(belong, probs = 0.25, na.rm = TRUE)
    )
```

## Helper functions

Code for helper functions.

```{r}
#| code-fold: true

logit = binomial()$linkfun
expit = binomial()$linkinv

plot_density_and_box <- function(plotdf, v) {
    list(
        ggplot(plotdf, aes(x = .data[[v]])) + 
            geom_density() + 
            theme_bw() +
            theme(axis.title.x = element_blank(), 
                  axis.text.x = element_blank(), 
                  axis.ticks.x = element_blank(),
                  plot.margin = margin()), 
        ggplot(plotdf, aes(x = .data[[v]], y = 0)) + 
            geom_boxplot(outlier.shape = NA) +
            geom_jitter(height = 0.25, width = 0, alpha = 0.1) +
            theme_bw() + 
            theme(axis.text.y = element_blank(), 
                  axis.title.y = element_blank(),
                  axis.ticks.y = element_blank(),
                  panel.grid.minor.y = element_blank(), 
                  panel.grid.major.y = element_blank(), 
                  plot.margin = margin(0))
    ) %>% 
        wrap_plots(nrow = 2, heights = c(0.8, 0.2))
}
```

# Variables

As outlined in Table 1 in the task description. 

```{r}
VARIABLES = list(
    outcome = "bel01", 
    focal = c("bull", "ATT4"), 
    nonfocal = c("female", "nld", "scie_std", "aca", "val", "comp", "ndiff", 
                 "nfof", "native", "nfewbooks", "joyread", "goal", "mot", 
                 "res", "swbp", "mean", "parent_sup", "ndis_clim", "GYM", "UNI")
)

glue("{length(VARIABLES$nonfocal)} non-focal variables used for modeling according to SAP: 
     {VARIABLES$nonfocal |> paste0(collapse = ', ')}")
```

## Overview

Elementary overview of distributions and correlations. 

```{r}
pisa2018 %>% 
    select(bel01, one_of(unlist(VARIABLES))) %>% 
    tbl_summary(
        type = all_continuous() ~ "continuous2",
        statistic = list(all_continuous() ~ c("{median} ({p25}, {p75})",
                                              "{mean} ({sd})",
                                             "{min}, {max}"))
    )
```

```{r fig.width=8, fig.height=8}
cormat = pisa2018 %>% 
    select(one_of(unlist(VARIABLES))) %>% 
    mutate(across(where(is.factor), as.numeric)) %>% 
    cor(use = "pairwise.complete.obs", method = "spearman")

ggcorrplot::ggcorrplot(cormat)
```

Focal predictors - important for target estimands.

```{r}
tbl_cross(pisa2018, bull, ATT4)
```

# Analysis

## Primary model 

```{r}
f_m = sprintf("bel01 ~ bull * ATT4 + %s + (1 | school.id)", 
              paste0(VARIABLES$nonfocal, collapse = " + ")) %>% 
    as.formula()

m = glmer(f_m, data = pisa2018, family = binomial)
summary(m)
```

### Diagnostics

Calibration including and excluding random effects, respectively. 

```{r}
rms::val.prob(logit = predict(m, re.form = NA), y = pisa2018$bel01)
rms::val.prob(logit = predict(m, re.form = NULL), y = pisa2018$bel01)
```

Randomized quantile residuals. Note that outlier tests may have inflated
error rates and are likely not relevant here. 

```{r}
s = simulateResiduals(m, seed = 26, plot = TRUE)
plotResiduals(s, pisa2018$ATT4)
plotResiduals(s, pisa2018$bull)
```

Random effect distribution (qq-plot for normality).

```{r}
ggplot(ranef(m)$school.id, aes(sample = `(Intercept)`)) + 
        stat_qq() + 
        stat_qq_line() + 
        theme_bw()
```

### Target estimands

Follow SAP to create representative students. We compute this for the 
full model only.

```{r}
preddf = pisa2018 %>% 
    select(bull, ATT4, one_of(VARIABLES$nonfocal)) %>% 
    # 1) create subsets of data by focal predictors
    group_by(bull, ATT4) %>% 
    # 2) compute 1st, 2nd, 3rd quartiles in subsets
    reframe(
        across(everything(), ~ quantile(.x, probs = c(0.25, 0.5, 0.75))),
        quantile = 1:3
    )

# 3) apply models
preddf$pred_m = predict(m, newdata = preddf, re.form = NA)
preddf$pred_m_prob = expit(preddf$pred_m)

# 4) marginalize over random effect distribution
rf = ranef(m)$school.id[, 1]
preddf %<>% 
    rowwise() %>% 
    mutate(
        pred_m_marg = mean(pred_m + rf), 
        pred_m_marg_prob = mean(expit(pred_m + rf))
    ) %>% 
    ungroup()
```

Example: estimated probability and whole distribution due to random effect 
variation for a single representative student. All the estimated distributions 
have a common shape around their point estimate. 

```{r}
#| message: false
#| warning: false
i = 1
preddf_1 = data.frame(
    id = 1, pred_m = preddf$pred_m[i], 
    pred_m_rf = preddf$pred_m[i] + rf, 
    pred_m_prob_rf = expit(preddf$pred_m[i] + rf)
)

ggplot(preddf_1, aes(x = pred_m_prob_rf)) +
    geom_vline(xintercept = expit(preddf_1$pred_m[1]), linetype = "dashed") + 
    geom_density() + 
    theme_bw()
```

## Secondary model 

```{r}
f_m2 = sprintf("bel01 ~ %s + (1 | school.id)", 
              paste0(VARIABLES$nonfocal, collapse = " + ")) %>% 
    as.formula()

m2 = glmer(f_m2, data = pisa2018, family = binomial)
summary(m2)
```

### Diagnostics

Calibration including and excluding random effects, respectively. 

```{r}
rms::val.prob(logit = predict(m2, re.form = NA), y = pisa2018$bel01)
rms::val.prob(logit = predict(m2, re.form = NULL), y = pisa2018$bel01)
```

Randomized quantile residuals. Note that outlier tests may have inflated
error rates and are likely not relevant here. 

```{r}
s = simulateResiduals(m2, seed = 26, plot = TRUE)
```

Random effect distribution (qq-plot for normality).

```{r}
ggplot(ranef(m2)$school.id, aes(sample = `(Intercept)`)) + 
        stat_qq() + 
        stat_qq_line() + 
        theme_bw()
```

## Model comparison

Compare models in terms of a likelihood ratio test to assess if the secondary
model would describe the data as well as the primary model.

The test is significant, indicating the the larger model "better" describes the
data based on frequentist likelihood theory. Also the AIC and BIC criteria are
smaller for the larger model, indicating it is to be preferred despite the 
larger number of fixed effects. 

```{r}
anova(m, m2)
```

## Bootstrap

Simulate 1000 bootstraps for full model. The bootstrap is done at cluster
level (i.e. schools).

### Bootstrap indices

```{r}
set.seed(13376)
b = 1000
sids = unique(pisa2018$school.id)
bootstraps = map(1:b, ~ sample(sids, length(sids), replace = TRUE))

glue("Statistics of relative frequency of unique schools per bootstrap:")
(map_dbl(bootstraps, n_distinct) / length(sids)) %>% summary()
```

### Run bootstrap

```{r}
f_b = sprintf("bel01 ~ bull * ATT4 + %s + (1 | school.id)", 
              paste0(VARIABLES$nonfocal, collapse = " + ")) %>% 
    as.formula()

do_bootstrap <- function(df, ind, f_b, VARIABLES) {
    # construct bootstrap sample
    # first take schools in bootstrap, then give them unique ids
    newids = set_names(ind, 1:length(ind))
    df_bi = map_df(1:length(newids), 
                   ~ df %>% 
                       filter(school.id == newids[.x]) %>% 
                       mutate(school.id = names(newids)[.x]))
    
    # fit model
    m_bi = glmer(f_b, data = df_bi, family = "binomial")
    
    preddf_bi = df_bi %>% 
        select(bull, ATT4, one_of(VARIABLES$nonfocal)) %>% 
        # 1) create subsets of data by focal predictors
        group_by(bull, ATT4) %>% 
        # 2) compute 1st, 2nd, 3rd quartiles in subsets
        reframe(
            across(everything(), ~ quantile(.x, probs = c(0.25, 0.5, 0.75))),
            quantile = 1:3
        )

    preddf_bi$pred_bi = predict(m_bi, newdata = preddf_bi, re.form = NA)
    preddf_bi$pred_bi_prob = expit(preddf_bi$pred_bi)

    rf = ranef(m_bi)$school.id[, 1]
    preddf_bi %<>% 
        rowwise() %>% 
        mutate(pred_bi_marg = mean(pred_bi + rf), 
               pred_bi_marg_prob = mean(expit(pred_bi + rf))) %>% 
        ungroup() 

    list(
        preddf_bi = preddf_bi %>% select(bull, ATT4, quantile, 
                                         pred_bi, pred_bi_prob, 
                                         pred_bi_marg, pred_bi_marg_prob), 
        optinfo = m_bi@optinfo, 
        nrow = nrow(df_bi), 
        rf = rf
    )
}
```

```{r warning=FALSE}
#| message: false
file_bootstrap = file.path(params$folder_analysis, 
                           sprintf("bootstrap-logistic%s.rds", SUFFIX))
if (!file.exists(file_bootstrap)) {
    glue("Running bootstraps and saving to {file_bootstrap}") %>% print()
    
    plan(multisession, workers = 12)
    res_b = foreach(
        i = seq_along(bootstraps), 
        .errorhandling = "pass", 
        .options.future = list(seed = 15)
    ) %dofuture%  {
        t_start = Sys.time()
        res = do_bootstrap(pisa2018, bootstraps[[i]], f_b, VARIABLES)
        res$runtime = difftime(now(), t_start, units = "s") %>% as.numeric()
        res
    }
    plan(sequential)
    saveRDS(res_b, file_bootstrap)    
} else {
    glue("Reading bootstrap results from {file_bootstrap}") %>% print()
    res_b = readRDS(file_bootstrap)
}

glue("Runtime per bootstrap in hours:")
summary(map_dbl(res_b, 'runtime')) / 3600
```

### Convergence

```{r}
#| message: false
#| warning: false
res_b %>% map_chr(~ .x$optinfo$message) %>% table()
```

### Insights

Random effect variances. 

```{r}
res_b %>% 
    map_dfr( ~ data.frame(rf = .x$rf), .id = "bootstrap") %>% 
    group_by(bootstrap) %>% 
    summarise(
        var_rf = var(rf)
    ) %>% 
    ggplot(aes(x = var_rf)) + 
    geom_boxplot() + 
    theme_bw()
```

# Final result

Display target estimands and variability as table. We sort by bullying, 
to ease comparison across levels of ATT and quantiles. 

These confidence intervals reflect the variability of the marginalized
probability estimates.

```{r rows.print=50}
#| message: false
#| warning: false

res_b_summary = bind_rows(map(res_b, "preddf_bi")) %>% 
    group_by(bull, ATT4, quantile) %>% 
    reframe(
        pred_marg_prob = quantile(pred_bi_marg_prob, probs = c(0.025, 0.25, 0.75, 0.975)), 
        pred_quantile = c("q2.5", "q25", "q75", "q97.5")
    ) %>% 
    pivot_wider(id_cols = c(bull, ATT4, quantile), 
                names_from = pred_quantile, 
                values_from = pred_marg_prob)

res = preddf %>% 
    select(bull, ATT4, quantile, pred_marg_prob = pred_m_marg_prob) %>% 
    left_join(res_b_summary, by = c("bull", "ATT4", "quantile"))

res$result = sprintf("%s<br>50%% [%s, %s]<br>95%% [%s, %s]", 
                   round(res$pred_marg_prob, 2), 
                   round(res$q25, 2), round(res$q75, 2), 
                   round(res$q2.5, 2), round(res$q97.5, 2))

res %>% 
    select(bull, ATT4, quantile, result) %>% 
    pivot_wider(id_cols = c(ATT4, quantile), 
                names_from = bull, values_from = result) %>% 
    arrange(ATT4, quantile) %>%
    gt() %>% 
    tab_style(
        style = list(cell_fill("#EEEEEE")), 
        locations = cells_body(rows = ATT4 %in% c(2, 4))
    ) %>% 
    fmt_markdown(columns = everything()) %>%
    cols_align("center", c("0", "1")) %>% 
    tab_spanner(label = "bull", columns = c("0", "1")) %>% 
    tab_options(quarto.disable_processing = TRUE)

ggplot(res, aes(group = ATT4, y = pred_marg_prob, color = fct_rev(factor(bull)), 
                x = factor(quantile))) + 
    geom_linerange(aes(ymin = q2.5, ymax = q97.5), position = position_dodge(0.5), 
                   linewidth = 1.5) +
    geom_linerange(aes(ymin = q25, ymax = q75), position = position_dodge(0.5), 
                   linewidth = 3) +
    geom_point(position = position_dodge(0.5), size = 5, shape = 3) + 
    scale_color_brewer(palette = "Set1", name = "bull") +
    coord_cartesian(ylim = c(0, 1)) +
    ylab("Probability of low sense of belonging") +
    xlab("Covariate quantile (dodged by increasing categories of ATT4)") + 
    theme_bw() + 
    theme(legend.position = "bottom")
```

# Further notes

- Convergence: the models had some trouble with convergence as the random 
    intercepts lead to models that are (close to being) singular. In particular
    within the bootstrap this leads to the situation that the optimiser says
    it converged, while lme4 specific diagnostics say it didn't. Likely 
    that does not matter, as we also did some case studies with `allFit` which
    indicated that all optimisers say they converged. In any case we switched
    back to the default "Nelder Mead" optimiser after some testing for 
    the analysis, as it seemed the most robust. 
- RF variance: because it is so low overall we do not provide further CIs 
    that reflect RF variance. 

# Conclusions

Overall model convergence seemed satisfactory across the bootstrap samples given
the insights above. Random intercepts for schools reflect the sampling 
strategy but do not capture much of the residual variance. In general the
presented modeling strategy should provide an adequate estimate for the 
desired target estimands. 

Bullying consistently leads to a drop in sense of belonging across all levels
of school truancy and all representative student values (i.e. subset quartiles).
This is expressed by a higher probability of being in group of students with 
sense of belonging lower than the 25% quantile.
These differences are are also demonstrated by non-overlapping confidence
intervals in all scenarios.

# Sessioninfo

To build this notebook:

- install `quarto`
- make sure the data is in the correct folder (either in global Data folder, 
    or as subfolder of the folder containing this notebook).
- build using the `make.R` script (global Data), or using the Render button in 
    RStudio (subfolder Data).

```{r}
params

glue("Runtime {difftime(Sys.time(), T_START, units = 'm') / 60} hours.")

sessioninfo::session_info()
```
