

---
title: "Students with attendance problems"
author: "Mariana Nold"
output: html_document
---
##### Load R packages
```{r  eval = FALSE}
source("RFiles/packages.r")
```

##### Load single imputed PISA data.
```{r  eval = FALSE}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task_Truancy"
load(file = file.path(PPath,"Orgdata/IMPDAT_PISA.RData"))
```

##### Prepare data, thus build PISA scales 
```{r  eval = FALSE}
source(file.path(PPath,"RFiles/01data_preperation.r"))
to.rm <- objects()[!objects() %in% c("PPath","pisa18","countries")]
rm(list = to.rm)
 #head(pisa18)
# dim(pisa18)
```


##### There should be at least ten students per school. Schools with fewer observations are excluded from the analysis. The variable nb.stud informs about the number of stundets per school and save final data set

```{r  eval = FALSE}
tab.id <- table(pisa18$school.id)
id.frame <- as.data.frame(tab.id)

nb.stud <- rep(id.frame$Freq,id.frame$Freq)
pisa18$nb.stud <- nb.stud
rm(tab.id,id.frame,nb.stud)
pisa.data <- subset(pisa18, pisa18$nb.stud > 9)
pisa18 <- pisa.data

rm(pisa.data)
summary(pisa18$nb.stud)
save(pisa18,file = file.path(PPath,"Files/pisa18.RData")) #file = file.path(PPath,"File/fits0.RData")
#load(file.path(PPath,"Files/pisa18.RData"))
```



##### Data sets for schools: Data set contains one row for each school and only school level variables
```{r  eval = FALSE}
sc.level <- names(pisa18) == toupper(names(pisa18))
sc.level[1] <- TRUE
pisa18_sc <- pisa18[,sc.level]
dim(pisa18_sc)
pisa18_sc <- pisa18_sc[!duplicated(pisa18_sc$school.id), ]
dim(pisa18_sc)
names(pisa18_sc)

save(pisa18,file = file.path(PPath,"Files/pisa18_sc.RData")) 
load(file.path(PPath,"Files/pisa18_sc.RData"))
```

##### Two models, one to check the other, model 2 to check model 1


```{r  eval = FALSE}
#
# theoretical model
model1 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL  + (1 | school.id))
# composition effect model
model2 <- as.formula(attp ~  immig  + female + escs  + RUR + DISH 
                     + GYM + STAFFSHORT + EDUSHORT + FEMFRAC + SN_SC + NN_SC + (1 | school.id))

models <- vector(mode = "list", length = 2)

models[[1]] <- model1
models[[2]] <- model2

```
##### fit models

```{r  eval = FALSE}
a.seed <- 12345 
a.iter <-  1000 
a.chains <- 4  
warmup <- 1000

#pisa18$WITHOUT  <- as.numeric(pisa18$FAILED >= 0.02) # above median

fits0 <- list()

fits0[[1]] <- brm(bf(models[[1]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits0[[2]] <- brm(bf(models[[2]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits1 <- list()

fits1[[1]] <- brm(bf(models[[1]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits1[[2]] <- brm(bf(models[[2]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 


save(fits0, file = file.path(PPath,"Files/fits0.RData"))
save(fits1, file = file.path(PPath,"Files/fits1.RData"))
```


##### load fits

```{r  eval = FALSE}
load(file.path(PPath, "Files/fits0.RData"))
load(file.path(PPath, "Files/fits1.RData"))
```


##### Study model parameters

```{r  eval = FALSE}
n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par0.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits0[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par0.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits0[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)


n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par1.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits1[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par1.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits1[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)

```
##### load functions
```{r  eval = FALSE}
source(file.path(PPath, "RFiles/02functions.r"))
```
##### compute predictions and Brier score
```{r  eval = FALSE}
#n01.models <- length(models)-1
ypredCand0 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand0[[i]] <- posterior_predict(fits0[[i]])
}
save("ypredCand0", file = file.path(PPath, "Files/ypredCand0.RData"))


br.candL20 <- list()
for (i in 1:n01.models) {
  colMeansM <- colMeans(ypredCand0[[i]])
  br.candL20[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 0],colMeansM)
 }

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores0.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL20[[i]])
  
}
sink(file = NULL)


ypredCand1 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand1[[i]] <- posterior_predict(fits1[[i]])
}
save("ypredCand1", file = file.path(PPath, "Files/ypredCand1.RData"))

## Brier Score 

br.candL21 <- list()
for (i in 1:n01.models) {
  br.candL21[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 1],colMeans(ypredCand1[[i]]))
 
}

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores1.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL21[[i]])
  
}
sink(file = NULL)
```

##### Posterior predictive checks, predictors of candidate model 1, not of key interest
```{r  eval = FALSE}
## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand20_out <- visualPPC1(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_out.pdf"))
ppcCand20_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand21_out <- visualPPC1(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_out.pdf"))
ppcCand21_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand10_in <- visualPPC1(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_in.pdf"))
ppcCand10_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand11_in <- visualPPC1(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_in.pdf"))
ppcCand11_in$ppc.plots
dev.off()

num <- length(ppcCand10_in$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand2W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand1W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


```
##### Posterior predictive checks, predictors of candidate model 2, this is of focal interest, if the model doas not pass this check, we need to include further predictors, or if not possible the stop and rethink the theoretical model


```{r  eval = FALSE}

## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand10_out <- visualPPC2(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_out.pdf"))
ppcCand10_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand11_out <- visualPPC2(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_out.pdf"))
ppcCand11_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand20_in <- visualPPC2(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_in.pdf"))
ppcCand20_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand21_in <- visualPPC2(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_in.pdf"))
ppcCand21_in$ppc.plots
dev.off()

num <- length(ppcCand21_in$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand1W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand2W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

```


##### more flexibility for important predictors, each model is of substantive interest, improve theoretical model, more flexibility for important predictors,
###### interaction based plausibility
```{r  eval = FALSE}
# theoretical model, possible improvements, WITHOUT = 0
imp0 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL + ld:goal + native:ld  + (1 + bull| school.id))
# theoretical model, possible improvements, WITHOUT = 0
imp1 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL + ld:fewbooks + native:fewbooks + goal:ld + (1 + bull| school.id))


models.imp<- vector(mode = "list", length = 2)

models.imp[[1]] <- imp0
models.imp[[2]] <- imp1

```
##### fit improved models, for pisa18$WITHOUT == 1

```{r  eval = FALSE}
fits.imp1 <- vector(mode = "list", length = 1)
a.seed <- 12345 
a.iter <-  3000 
a.chains <- 4  

fits.imp1[[1]] <- brm(bf(models.imp[[2]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit"))

save(fits.imp1, file = file.path(PPath,"Files/fits.imp1.RData"))
#load(file.path(PPath,"Files/fits.imp1.RData"))
```
##### fit improved models, for pisa18$WITHOUT == 0

```{r  eval = FALSE}
 
fits.imp0 <- vector(mode = "list", length = 1)
a.seed <- 12345 
a.iter <-  3000 
a.chains <- 4  

fits.imp0[[1]] <- brm(bf(models.imp[[1]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit"))

save(fits.imp0, file = file.path(PPath,"Files/fits.imp0.RData"))

#load(file.path(PPath,"Files/fits.imp0.RData"))

```

##### loo-compare for the improved theoretical model, WITHOUT == 1
```{r  eval = FALSE}

fits.imp1[[2]] <- fits1[[1]]
n01.models <- length(fits.imp1)
# Compute loo

loo_list1 <- lapply(fits.imp1, loo, cores = 1)
loo_compare(loo_list1)

save("loo_list1", file = file.path(PPath, "Files","loo_list1.RData"))
#load(file.path(PPath, "Files","loo_list0.RData"))

sink(file = file.path(PPath, "Files","loo_info1.txt"))
for (i in 1:n01.models) {
  print(paste("******** "))
  print(paste("model ", i))
  print(loo_list1[[i]])
  print(paste("******** "))
}
#print(loo_list[[1]])
sink(file = NULL) # All models are similar, first model is best

#wtsStacking1 <- list()

#wtsStacking1 <- loo_model_weights(loo_list1, method = "stacking")
#print(wtsStacking1)


#sink(file = file.path(PPath, "Results","wtsStacking1.txt"))
#print(wtsStacking1)
#print(paste("******** "))
#sink(file = NULL)

```

##### Stacking, only if continous model extension is not possible
##### https://jrnold.github.io/bayesian_notes/model-comparison.html
##### loo-compare for the improved theoretical model, WITHOUT == 0
```{r  eval = FALSE}
fits.imp0[[2]] <- fits0[[1]]
n01.models <- length(fits.imp0)

loo_list0 <- lapply(fits.imp0, loo, cores = 1)
loo_compare(loo_list0)

save("loo_list0", file = file.path(PPath, "Files","loo_list0.RData"))
#load(file.path(PPath, "Files","loo_list0.RData"))

sink(file = file.path(PPath, "Files","loo_info0.txt"))
for (i in 1:n01.models) {
  print(paste("******** "))
  print(paste("model ", i))
  print(loo_list0[[i]])
  print(paste("******** "))
}
#print(loo_list[[1]])
sink(file = NULL)

#wtsStacking0 <- list()

#wtsStacking0 <- loo_model_weights(loo_list0, method = "stacking")
#print(wtsStacking0)


#sink(file = file.path(PPath, "Results","wtsStacking0.txt"))
#print(wtsStacking0)
#print(paste("******** "))
#sink(file = NULL)

```




##### PPC for improved model, WITHOUT = 0
```{r  eval = FALSE}
set.seed(a.seed)
ypred0 <- posterior_predict(fits.imp0[[1]]) 
save("ypred0", file = file.path("Files/ypred0.RData"))

color_scheme_set(scheme = "pink")
ppc0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypred0)
pdf(file.path(PPath, "Results/ppc0.pdf"))
ppc0
dev.off()

```

##### PPC for improved model, WITHOUT = 0
```{r  eval = FALSE}
set.seed(a.seed)

ypred1 <- posterior_predict(fits.imp1[[1]]) 
save("ypred1", file = file.path("Files/ypred1.RData"))

color_scheme_set(scheme = "purple")
ppc0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypred1, WITHOUT = 1)
pdf(file.path(PPath, "Results/ppc1.pdf"))
ppc0
dev.off()



```


###### marginaeffects code
###### https://www.andrewheiss.com/blog/2022/11/29/conditional-marginal-marginaleffects/



##### result of focal interest, WITHOUT = 0
##### https://marginaleffects.com/vignettes/brms.html
```{r eval = FALSE}
conditional_preds0 <- avg_predictions(
  fits.imp0[[1]], 
  by = "ld", 
  re_formula = NA
) %>% 
  posteriordraws()
```

##### result of focal interest, WITHOUT = 0, plot
```{r eval = FALSE}
clrs <- NatParksPalettes::natparks.pals("BryceCanyon")
p_conditional_preds0 <- conditional_preds0 %>% 
  ggplot(aes(x = draw, fill = factor(ld))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "ld",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
pdf(file.path(PPath, "Results/result0_ld.pdf"))
p_conditional_preds0
dev.off()
```
#####  Cluster-specific average treatment effect (when intercept is 0)
```{r eval = FALSE}
conditional_ate0 <- avg_comparisons(
   fits.imp0[[1]], 
  variables = "ld",
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r eval = FALSE}
p_conditional_ate0 <- conditional_ate0 %>% 
  ggplot(aes(x = draw)) +
  stat_halfeye(fill = clrs[3]) +
 # scale_x_continuous(labels = label_dollar(style_negative = "minus")) +
  labs(x = "(ld = 1) − (ld = 0)", y = "Density", 
       title = "Conditional cluster-specific ATE",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(-0.2,0.3)) #+
  #theme_nice() #+
  #theme(plot.subtitle = element_markdown())
pdf(file.path(PPath, "Results/result0_ld_ate.pdf"))
p_conditional_ate0
dev.off()
```

##### result of focal interest, WITHOUT = 1
```{r eval = FALSE}
conditional_preds1 <- avg_predictions(
  fits.imp1[[1]], 
  #newdata = datagrid(ld = c(0, 1)), 
  by = "ld", 
  re_formula = NA
) %>% 
  posteriordraws()
```

##### result of focal interest, WITHOUT = 1, plot
```{r eval = FALSE}

p_conditional_preds1 <- conditional_preds1 %>% 
  ggplot(aes(x = draw, fill = factor(ld))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "ld",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
pdf(file.path(PPath, "Results/result1_ld.pdf"))
p_conditional_preds1
dev.off()
```
#####  Cluster-specific average treatment effect (when intercept is 0)
```{r eval = FALSE}
conditional_ate1 <- avg_comparisons(
   fits.imp1[[1]], 
  variables = "ld",
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r eval = FALSE}
p_conditional_ate1 <- conditional_ate1 %>% 
  ggplot(aes(x = draw)) +
  stat_halfeye(fill = clrs[3]) +
 # scale_x_continuous(labels = label_dollar(style_negative = "minus")) +
  labs(x = "(ld = 1) − (ld = 0)", y = "Density", 
       title = "Conditional cluster-specific ATE",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(-0.2,0.3)) #+
  #theme_nice() #+
  #theme(plot.subtitle = element_markdown())
pdf(file.path(PPath, "Results/result1_ld_ate.pdf"))
p_conditional_ate1
dev.off()
```

##### bull, WITHOUT = 0
```{r  eval = FALSE}
conditional_preds0_b <- avg_predictions(
  fits.imp0[[1]], 
 # newdata = datagrid(ld = c(0, 1)), 
  by = "bull", 
  re_formula = NA
) %>% 
  posteriordraws()
```
##### plot
```{r  eval = FALSE}
p_conditional_preds0_b <- conditional_preds0_b %>% 
  ggplot(aes(x = draw, fill = factor(bull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "bull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
pdf(file.path(PPath, "Results/result0_bull.pdf"))
p_conditional_preds0_b
dev.off()
```
##### bull, WITHOUT = 1
```{r  eval = FALSE}
conditional_preds1_b <- avg_predictions(
  fits.imp1[[1]], 
 # newdata = datagrid(ld = c(0, 1)), 
  by = "bull", 
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r  eval = FALSE}
p_conditional_preds1_b <- conditional_preds1_b %>% 
  ggplot(aes(x = draw, fill = factor(bull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "bull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
pdf(file.path(PPath, "Results/result1_bull.pdf"))
p_conditional_preds1_b
dev.off()
```
##### ATE of bull, WITHOUT = 0
```{r  eval = FALSE}
conditional_ate0_b <- avg_comparisons(
   fits.imp0[[1]], 
  variables = "bull",
  re_formula = NA
) %>% 
  posteriordraws()


```

```{r  eval = FALSE}
p_conditional_ate0_b <- conditional_ate0_b %>% 
  ggplot(aes(x = draw)) +
  stat_halfeye(fill = clrs[3]) +
 # scale_x_continuous(labels = label_dollar(style_negative = "minus")) +
  labs(x = "(bull = 1) − (bull = 0)", y = "Density", 
       title = "Conditional cluster-specific ATE",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(-0.2,0.3)) #+
  #theme_nice() #+
  #theme(plot.subtitle = element_markdown())
pdf(file.path(PPath, "Results/result0_bull_ate.pdf"))
p_conditional_ate0_b
dev.off()
```


##### ATE of bull, WITHOUT = 1
```{r  eval = FALSE}
conditional_ate1_b <- avg_comparisons(
   fits.imp1[[1]], 
  variables = "bull",
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r  eval = FALSE}
p_conditional_ate1_b <- conditional_ate1_b %>% 
  ggplot(aes(x = draw)) +
  stat_halfeye(fill = clrs[3]) +
 # scale_x_continuous(labels = label_dollar(style_negative = "minus")) +
  labs(x = "(bull = 1) − (bull = 0)", y = "Density", 
       title = "Conditional cluster-specific ATE",
       subtitle = "Typical cluster where school intercept, slope bull = 0") +
  coord_cartesian(xlim = c(-0.2,0.3)) #+
  #theme_nice() #+
  #theme(plot.subtitle = element_markdown())
pdf(file.path(PPath, "Results/result1_bull_ate.pdf"))
p_conditional_ate1_b
dev.off()
```
