

---
title: "Students with attendance problems"
author: "Mariana Nold"
output: html_document
---
##### Load R packages
```{r  eval = FALSE}
source("RFiles/packages.r")
```

##### Load single imputed PISA data.
```{r  eval = FALSE}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task_Truancy"
load(file = file.path(PPath,"Orgdata/IMPDAT_PISA.RData"))
```

##### Prepare data, thus build PISA scales 
```{r  eval = FALSE}
source(file.path(PPath,"RFiles/01data_preperation.r"))
to.rm <- objects()[!objects() %in% c("PPath","pisa18","countries")]
rm(list = to.rm)
 #head(pisa18)
# dim(pisa18)
```


##### There should be at least ten students per school. Schools with fewer observations are excluded from the analysis. The variable nb.stud informs about the number of stundets per school and save final data set

```{r  eval = FALSE}
tab.id <- table(pisa18$school.id)
id.frame <- as.data.frame(tab.id)

nb.stud <- rep(id.frame$Freq,id.frame$Freq)
pisa18$nb.stud <- nb.stud
rm(tab.id,id.frame,nb.stud)
pisa.data <- subset(pisa18, pisa18$nb.stud > 9)
pisa18 <- pisa.data

rm(pisa.data)
summary(pisa18$nb.stud)
save(pisa18,file = file.path(PPath,"Files/pisa18.RData")) #file = file.path(PPath,"File/fits0.RData")
#load(file.path(PPath,"Files/pisa18.RData"))
```



##### Data sets for schools: Data set contains one row for each school and only school level variables
```{r  eval = FALSE}
sc.level <- names(pisa18) == toupper(names(pisa18))
sc.level[1] <- TRUE
pisa18_sc <- pisa18[,sc.level]
dim(pisa18_sc)
pisa18_sc <- pisa18_sc[!duplicated(pisa18_sc$school.id), ]
dim(pisa18_sc)
names(pisa18_sc)

save(pisa18,file = file.path(PPath,"Files/pisa18_sc.RData")) 
load(file.path(PPath,"Files/pisa18_sc.RData"))
```

##### Two models, one to check the other, model 2 to check model 1


```{r  eval = FALSE}
#
# theoretical model
model1 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL  + (1 | school.id))
# composition effect model
model2 <- as.formula(attp ~  immig  + female + escs  + RUR + DISH 
                     + GYM + STAFFSHORT + EDUSHORT + FEMFRAC + SN_SC + NN_SC + (1 | school.id))

models <- vector(mode = "list", length = 2)

models[[1]] <- model1
models[[2]] <- model2

```
##### fit models

```{r  eval = FALSE}
a.seed <- 12345 
a.iter <-  1000 
a.chains <- 4  
warmup <- 1000

#pisa18$WITHOUT  <- as.numeric(pisa18$FAILED >= 0.02) # above median

fits0 <- list()

fits0[[1]] <- brm(bf(models[[1]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits0[[2]] <- brm(bf(models[[2]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits1 <- list()

fits1[[1]] <- brm(bf(models[[1]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 

fits1[[2]] <- brm(bf(models[[2]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit")) 


save(fits0, file = file.path(PPath,"Files/fits0.RData"))
save(fits1, file = file.path(PPath,"Files/fits1.RData"))
```


##### load fits

```{r  eval = FALSE}
load(file.path(PPath, "Files/fits0.RData"))
load(file.path(PPath, "Files/fits1.RData"))
```


##### Study model parameters

```{r  eval = FALSE}
n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par0.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits0[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par0.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits0[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)


n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par1.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits1[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par1.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits1[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)

```
##### load functions
```{r  eval = FALSE}
source(file.path(PPath, "RFiles/02functions.r"))
```
##### compute predictions and Brier score
```{r  eval = FALSE}
#n01.models <- length(models)-1
ypredCand0 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand0[[i]] <- posterior_predict(fits0[[i]])
}
save("ypredCand0", file = file.path(PPath, "Files/ypredCand0.RData"))


br.candL20 <- list()
for (i in 1:n01.models) {
  colMeansM <- colMeans(ypredCand0[[i]])
  br.candL20[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 0],colMeansM)
 }

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores0.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL20[[i]])
  
}
sink(file = NULL)


ypredCand1 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand1[[i]] <- posterior_predict(fits1[[i]])
}
save("ypredCand1", file = file.path(PPath, "Files/ypredCand1.RData"))

## Brier Score 

br.candL21 <- list()
for (i in 1:n01.models) {
  br.candL21[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 1],colMeans(ypredCand1[[i]]))
 
}

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores1.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL21[[i]])
  
}
sink(file = NULL)
```

##### Posterior predictive checks, predictors of candidate model 1, not of key interest
```{r  eval = FALSE}
## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand20_out <- visualPPC1(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_out.pdf"))
ppcCand20_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand21_out <- visualPPC1(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_out.pdf"))
ppcCand21_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand10_in <- visualPPC1(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_in.pdf"))
ppcCand10_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand11_in <- visualPPC1(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_in.pdf"))
ppcCand11_in$ppc.plots
dev.off()

num <- length(ppcCand10_in$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand2W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand1W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


```
##### Posterior predictive checks, predictors of candidate model 2, this is of focal interest, if the model doas not pass this check, we need to include further predictors, or if not possible the stop and rethink the theoretical model


```{r  eval = FALSE}

## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand10_out <- visualPPC2(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_out.pdf"))
ppcCand10_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand11_out <- visualPPC2(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_out.pdf"))
ppcCand11_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand20_in <- visualPPC2(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_in.pdf"))
ppcCand20_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand21_in <- visualPPC2(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_in.pdf"))
ppcCand21_in$ppc.plots
dev.off()

num <- length(ppcCand21_in$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand1W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand2W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

```


##### more flexibility for important predictors, each model is of substantive interest, improve theoretical model, more flexibility for important predictors,
###### interaction based plausibility
```{r  eval = FALSE}
# theoretical model, possible improvements, WITHOUT = 0
imp0 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL + ld:goal + native:ld  + (1 + bull| school.id))
# theoretical model, possible improvements, WITHOUT = 1
imp1 <- as.formula(attp ~ aca + native + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot + RES + VAL + ld:fewbooks + native:fewbooks + goal:ld + (1 + bull| school.id))


models.imp<- vector(mode = "list", length = 2)

models.imp[[1]] <- imp0
models.imp[[2]] <- imp1

```
##### fit improved models, for pisa18$WITHOUT == 1

```{r  eval = FALSE}
fits.imp1 <- vector(mode = "list", length = 1)
a.seed <- 12345 
a.iter <-  3000 
a.chains <- 4  

fits.imp1[[1]] <- brm(bf(models.imp[[2]]), data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit"))

save(fits.imp1, file = file.path(PPath,"Files/fits.imp1.RData"))
#load(file.path(PPath,"Files/fits.imp1.RData"))
```
##### fit improved models, for pisa18$WITHOUT == 0

```{r  eval = FALSE}
 
fits.imp0 <- vector(mode = "list", length = 1)
a.seed <- 12345 
a.iter <-  3000 
a.chains <- 4  

fits.imp0[[1]] <- brm(bf(models.imp[[1]]), data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, warmup = warmup, bernoulli(link="logit"))

save(fits.imp0, file = file.path(PPath,"Files/fits.imp0.RData"))

#load(file.path(PPath,"Files/fits.imp0.RData"))

```

##### loo-compare for the improved theoretical model, WITHOUT == 1
```{r  eval = FALSE}

fits.imp1[[2]] <- fits1[[1]]
n01.models <- length(fits.imp1)
# Compute loo

loo_list1 <- lapply(fits.imp1, loo, cores = 1)
loo_compare(loo_list1)

save("loo_list1", file = file.path(PPath, "Files","loo_list1.RData"))
#load(file.path(PPath, "Files","loo_list0.RData"))

sink(file = file.path(PPath, "Files","loo_info1.txt"))
for (i in 1:n01.models) {
  print(paste("******** "))
  print(paste("model ", i))
  print(loo_list1[[i]])
  print(paste("******** "))
}
#print(loo_list[[1]])
sink(file = NULL) # All models are similar, first model is best

#wtsStacking1 <- list()

#wtsStacking1 <- loo_model_weights(loo_list1, method = "stacking")
#print(wtsStacking1)


#sink(file = file.path(PPath, "Results","wtsStacking1.txt"))
#print(wtsStacking1)
#print(paste("******** "))
#sink(file = NULL)

```

##### Stacking, only if continous model extension is not possible
##### https://jrnold.github.io/bayesian_notes/model-comparison.html
##### loo-compare for the improved theoretical model, WITHOUT == 0
```{r  eval = FALSE}
fits.imp0[[2]] <- fits0[[1]]
n01.models <- length(fits.imp0)

loo_list0 <- lapply(fits.imp0, loo, cores = 1)
loo_compare(loo_list0)

save("loo_list0", file = file.path(PPath, "Files","loo_list0.RData"))
#load(file.path(PPath, "Files","loo_list0.RData"))

sink(file = file.path(PPath, "Files","loo_info0.txt"))
for (i in 1:n01.models) {
  print(paste("******** "))
  print(paste("model ", i))
  print(loo_list0[[i]])
  print(paste("******** "))
}
#print(loo_list[[1]])
sink(file = NULL)

#wtsStacking0 <- list()

#wtsStacking0 <- loo_model_weights(loo_list0, method = "stacking")
#print(wtsStacking0)


#sink(file = file.path(PPath, "Results","wtsStacking0.txt"))
#print(wtsStacking0)
#print(paste("******** "))
#sink(file = NULL)

```




##### PPC for improved model, WITHOUT = 0
```{r  eval = FALSE}
set.seed(a.seed)
ypred0 <- posterior_predict(fits.imp0[[1]]) 
save("ypred0", file = file.path("Files/ypred0.RData"))

color_scheme_set(scheme = "pink")
ppc0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypred0)
pdf(file.path(PPath, "Results/ppc0.pdf"))
ppc0
dev.off()

```

##### PPC for improved model, WITHOUT = 0
```{r  eval = FALSE}
set.seed(a.seed)

ypred1 <- posterior_predict(fits.imp1[[1]]) 
save("ypred1", file = file.path("Files/ypred1.RData"))

color_scheme_set(scheme = "purple")
ppc0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypred1, WITHOUT = 1)
pdf(file.path(PPath, "Results/ppc1.pdf"))
ppc0
dev.off()



```


###### marginaeffects code
###### https://www.andrewheiss.com/blog/2022/11/29/conditional-marginal-marginaleffects/



##### result of focal interest, WITHOUT = 0
##### https://marginaleffects.com/vignettes/brms.html
```{r eval = FALSE}
conditional_preds0 <- avg_predictions(
  fits.imp0[[1]], 
  by = "ld", 
  re_formula = NA
) %>% 
  posteriordraws()
```

##### result of focal interest, WITHOUT = 0, plot
```{r eval = FALSE}
clrs <- NatParksPalettes::natparks.pals("BryceCanyon")
p_conditional_preds0 <- conditional_preds0 %>% 
  ggplot(aes(x = draw, fill = factor(ld))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "ld",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
#pdf(file.path(PPath, "Results/result0_ld.pdf"))
p0ld <- p_conditional_preds0
#dev.off()
```

```{r eval = FALSE}
conditional_preds1 <- avg_predictions(
  fits.imp1[[1]], 
  by = "ld", 
  re_formula = NA
) %>% 
  posteriordraws()
```

##### result of focal interest, WITHOUT = 1, plot
```{r eval = FALSE}

p_conditional_preds1 <- conditional_preds1 %>% 
  ggplot(aes(x = draw, fill = factor(ld))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "ld",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 1") +
  coord_cartesian(xlim = c(0, 0.4)) 
#pdf(file.path(PPath, "Results/result1_ld.pdf"))
p1ld <- p_conditional_preds1
#dev.off()
```


##### bull, WITHOUT = 0
```{r  eval = FALSE}
conditional_preds0_b <- avg_predictions(
  fits.imp0[[1]], 
  by = "bull", 
  re_formula = NA
) %>% 
  posteriordraws()
```
##### plot
```{r  eval = FALSE}
p_conditional_preds0_b <- conditional_preds0_b %>% 
  ggplot(aes(x = draw, fill = factor(bull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "bull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 0") +
  coord_cartesian(xlim = c(0, 0.4)) 
#pdf(file.path(PPath, "Results/result0_bull.pdf"))
p0bull <- p_conditional_preds0_b
#dev.off()
```
##### bull, WITHOUT = 1





```{r  eval = FALSE}
conditional_preds1_b <- avg_predictions(
  fits.imp1[[1]], 
  by = "bull", 
  re_formula = NA
) %>% 
  posteriordraws()
```



```{r  eval = FALSE}
p_conditional_preds1_b <- conditional_preds1_b %>% 
  ggplot(aes(x = draw, fill = factor(bull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1])) +
  labs(x = "Attendance problem", y = "Density", fill = "bull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 1") +
  coord_cartesian(xlim = c(0, 0.4)) 
#pdf(file.path(PPath, "Results/result1_bull.pdf"))
p1bull <- p_conditional_preds1_b
#dev.off()
```

##### plot to pdf
```{r  eval = FALSE}
pdf(file.path(PPath,"Results/bull.pdf"))
grid.arrange(p0bull, p1bull, nrow = 2)
dev.off()

pdf(file.path(PPath,"Results/ld.pdf"))
grid.arrange(p0ld, p1ld, nrow = 2)
dev.off()
```
##### final plot, WITHOUT = 1, bull and ld in one plot
```{r eval = FALSE}
conditional_preds1_both <- avg_predictions(
  fits.imp1[[1]], 
  by = c("bull", "ld"),
  re_formula = NA
) %>% 
  posteriordraws()
```
##### predictors not used in by-option are as observed
##### before posteriordraws() is used
```{r eval = FALSE}
#dat <- attr(conditional_preds1_both,"newdata_modeldata")
# table(pisa18[pisa18$WITHOUT==1,]$belong)
# table(dat$belong)
```


```{r eval = FALSE}
conditional_preds1_both$ldbull <- interaction(conditional_preds1_both$ld,conditional_preds1_both$bull)
```

```{r eval = FALSE}
# library(ggdist)
p_conditional_preds1_both <- conditional_preds1_both %>% 
  ggplot(aes(x = draw, fill = factor(ldbull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1],clrs[3], clrs[4])) +
  labs(x = "Attendance problem", y = "Density", fill = "ldbull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 1") +
  coord_cartesian(xlim = c(0.1, 0.5)) 
#pdf(file.path(PPath, "Results/result1_bull.pdf"))
p1ldbull <- p_conditional_preds1_both
```

##### final plot, WITHOUT = 0
```{r eval = FALSE}
conditional_preds0_both <- avg_predictions(
  fits.imp0[[1]], 
 # newdata = datagrid(ld = c(0, 1)), 
  by = c("bull", "ld"),
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r eval = FALSE}
conditional_preds0_both$ldbull <- interaction(conditional_preds0_both$ld,conditional_preds0_both$bull)
```

```{r eval = FALSE}
p_conditional_preds0_both <- conditional_preds0_both %>% 
  ggplot(aes(x = draw, fill = factor(ldbull))) +
  stat_halfeye() +
  scale_fill_manual(values = c(clrs[5], clrs[1],clrs[3], clrs[4])) +
  labs(x = "Attendance problem", y = "Density", fill = "ldbull",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 0") +
  coord_cartesian(xlim = c(0.1, 0.5)) 
#pdf(file.path(PPath, "Results/result1_bull.pdf"))
p0ldbull <- p_conditional_preds0_both
```
##### plot to pdf
```{r  eval = FALSE}
pdf(file.path(PPath,"Results/ldbull.pdf"))
grid.arrange(p0ldbull, p1ldbull, nrow = 2)
dev.off()

```
##### goal, WITHOUT = 0
```{r eval = FALSE}
conditional_preds0_goal <- avg_predictions(
  fits.imp0[[1]], 
 # newdata = datagrid(ld = c(0, 1)), 
  by = c("goal"),
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r eval = FALSE}
p_conditional_preds0_goal <- conditional_preds0_goal %>% 
  ggplot(aes(x = draw, fill = factor(goal))) +
  stat_halfeye() +
  #scale_fill_manual(values = c(clrs[5], clrs[1],clrs[3], clrs[4])) +
  labs(x = "Attendance problem", y = "Density", fill = "goal",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 0") +
  coord_cartesian(xlim = c(0.0, 0.5)) 
#pdf(file.path(PPath, "Results/result1_bull.pdf"))
p0goal <- p_conditional_preds0_goal
```
##### goal, WITHOUT = 1
```{r eval = FALSE}
conditional_preds1_goal <- avg_predictions(
  fits.imp1[[1]], 
 # newdata = datagrid(ld = c(0, 1)), 
  by = c("goal"),
  re_formula = NA
) %>% 
  posteriordraws()
```

```{r eval = FALSE}
p_conditional_preds1_goal <- conditional_preds1_goal %>% 
  ggplot(aes(x = draw, fill = factor(goal))) +
  stat_halfeye() +
  #scale_fill_manual(values = c(clrs[5], clrs[1],clrs[3], clrs[4])) +
  labs(x = "Attendance problem", y = "Density", fill = "goal",
       title = "Conditional cluster-specific means",
       subtitle = "Typical cluster, WITHOUT = 1") +
  coord_cartesian(xlim = c(0.0, 0.5)) 
#pdf(file.path(PPath, "Results/result1_bull.pdf"))
p1goal <- p_conditional_preds1_goal
```

##### plot to pdf
```{r  eval = FALSE}
pdf(file.path(PPath,"Results/goal.pdf"))
grid.arrange(p0goal, p1goal, nrow = 2)
dev.off()

```
### Students within schools, combination with VAL and RES

##### RES and VAL for WITHOUT = 1
```{r  eval = FALSE}
fits.imp1[[1]]$data$VAL3 <- cut_number(fits.imp1[[1]]$data$VAL,3)
fits.imp1[[1]]$data$RES3 <- cut_number(fits.imp1[[1]]$data$RES,3)


foc.grid <- fits.imp1[[1]]$data |> 
  summarise(
    aca = mean(aca),native=mean(native), ld=mean(ld),uni= mean(uni), bull = mean(bull),  
    fewbooks = mean(fewbooks),  belong  = mean(belong),   joyread = mean(joyread),
    goal  = mean(goal),   mot =mean(mot), RES = mean(RES),  VAL = mean(VAL),
    .by = c(VAL3,RES3)
  )

foc.grid.cnt <- fits.imp1[[1]]$data  |> count(VAL3,RES3) |>
  mutate(Prop = n / sum(n))

RESVAL1 <- predictions(
  fits.imp1[[1]], 
  newdata = foc.grid[,3:14], 
  re_formula = NA
)

RESVAL_W1 = cbind(foc.grid[,1:2],RESVAL1)
RESVAL_W1
```
##### Write to file
```{r eval = FALSE}
sink(file = file.path(PPath, "Results/resval1.txt"))
RESVAL_W1[,1:18]
sink(file = NULL)
```




##### RES and VAL for WITHOUT = 0
```{r  eval = FALSE}
fits.imp0[[1]]$data$VAL3 <- cut_number(fits.imp0[[1]]$data$VAL,3)
fits.imp0[[1]]$data$RES3 <- cut_number(fits.imp0[[1]]$data$RES,3)


foc.grid0 <- fits.imp0[[1]]$data |> 
  summarise(
    aca = mean(aca),native=mean(native), ld=mean(ld),uni= mean(uni), bull = mean(bull),  
    fewbooks = mean(fewbooks),  belong  = mean(belong),   joyread = mean(joyread),
    goal  = mean(goal),   mot =mean(mot), RES = mean(RES),  VAL = mean(VAL),
    .by = c(VAL3,RES3)
  )

foc.grid.cnt0 <- fits.imp0[[1]]$data  |> count(VAL3,RES3) |>
  mutate(Prop = n / sum(n))

RESVAL0 <- predictions(
  fits.imp0[[1]], 
  newdata = foc.grid0[,3:14], 
  re_formula = NA
)

RESVAL_W0 = cbind(foc.grid0[,1:2],RESVAL0)
RESVAL_W0
```
##### Write to file
```{r eval = FALSE}
sink(file = file.path(PPath, "Results/resval0.txt"))
RESVAL_W0[,1:18]
sink(file = NULL)
```
#### Result analogous Figure 13.19 https://sites.google.com/view/statistical-modeling-with-r/part-iii
##### cannot find 'slo.eff.m2.brms'
```{r eval = FALSE}
library("ggbreak")
library("reshape2")
DF <- fits.imp0[[1]]$data
pop.ef.m2.brms=mcmc_areas(fits.imp0[[1]], prob=0.62, prob_outer = 0.95, regex_pars = "b")+
  geom_density(lwd=1.2, alpha=0.9)+ 
  theme_bw()+
  scale_x_break(c(5, 50),scales = "fixed")+ 
#  scale_y_discrete(labels=c("Int.boys", "Slope.boys", "diff.int","diff.slope"))+
  theme(axis.text.y=element_text(size = 14, angle=90, hjust=0.4),
        axis.text.x=element_text(size = 16, angle=90),  
        axis.title.y = element_blank(),
        strip.background=element_rect(colour = "black", fill = "white"),
        strip.text = element_text(size = 14)) 
# Gathering the posterior distributions of the group-level effects (random effects in the Frequentist framework)
gr.eff.m2.brms=as_draws_df(fits.imp0[[1]], variable = "^r_", regex = TRUE)
gr.eff.m2.brms=gr.eff.m2.brms[,-c(199,200,201)] # delete the last three columns: .chain, .draw, and .iteration
names(gr.eff.m2.brms)=paste(rep(c("int", "slope"),each=length(unique(DF$school.id))), 
                            rep(unique(DF$school.id),times=2)) # shortening the names
# Reformatting gr.eff.m2.brms as "long format": stacked columns
gr.eff.m2.brms=melt(gr.eff.m2.brms,value.name = "gr.eff.school") # stores posterior distr of all group level effects effects from 
int.eff.m2.brms=gr.eff.m2.brms[grep(pattern="^int", gr.eff.m2.brms$variable),] # posterior distr of int group level effects 
slopes.eff.m2.brms=gr.eff.m2.brms[grep(pattern="^slope", gr.eff.m2.brms$variable),] # posterior distr of slope group level effects

# Posterior distribution plots of intercepts and slopes group effects
m2.brms.int.schools=ggplot(data=int.eff.m2.brms,aes(x=gr.eff.school, col=variable))+
  geom_density()+
  theme_bw()+
  geom_vline(xintercept = 0, linetype=2)+
  labs(x="Group effect in Intercept", y="Probability density")+
  theme(legend.position="none",
        axis.text=element_text(size = 16),  
        axis.title = element_text(size = 18))
m2.brms.slopes.schools=ggplot(data=slo.eff.m2.brms,aes(x=gr.eff.school, col=variable))+
  geom_density()+
  theme_bw()+
  geom_vline(xintercept = 0, linetype=2)+
  labs(x="Group effect in Slope", y="Probability density")+
  theme(legend.position="none",
        axis.text=element_text(size = 16),  
        axis.title = element_text(size = 18))
# Obtains means, SE and 95% Credible Intervals of Group effects for m2.brms
stats.ranef.m2.brms=as.data.frame(ranef(m2.brms)) 
names(stats.ranef.m2.brms)=c("mean.int", "se.int", "Q2.5.int", "Q97.5.int",
                             "mean.slope", "se.slope", "Q2.5.slope", "Q97.5.slope") # put shorter names in the data frame
stats.ranef.m2.brms$school=unique(DF$school.id) # adds schoolid  
CrInt95.int.m2.brms=ggplot(data=stats.ranef.m2.brms, aes(x=reorder(school, mean.int), y=mean.int))+
  geom_point()+
  coord_flip()+
  geom_linerange(aes(ymin=Q2.5.int,ymax=Q97.5.int))+
  geom_hline(yintercept=0, size=1, linetype=2)+
  scale_y_continuous(breaks=seq(from=-30, to=40, by=10))+
  theme_bw()+
  labs(x="School", y="Group intercept")+
  theme(axis.title.x = element_text(size=18),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.text.x = element_text(size=16))
CrInt95.slopes.m2.brms=ggplot(data=stats.ranef.m2.brms, aes(x=reorder(school, mean.int), y=mean.slope))+
  geom_point()+
  coord_flip()+
  geom_linerange(aes(ymin=Q2.5.slope,ymax=Q97.5.slope))+
  geom_hline(yintercept=0, size=1, linetype=2)+
  scale_y_continuous(breaks=seq(from=-30, to=40, by=10))+
  theme_bw()+
  labs(x="School", y="Group slope")+
  theme(axis.title.x = element_text(size=18),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.text.x = element_text(size=16))
cor.sl.int.m2.brms=ggplot(data=stats.ranef.m2.brms, aes(x=mean.int, y=mean.slope))+
  geom_point()+
  geom_hline(yintercept=0, size=1, linetype=2)+
  geom_vline(xintercept=0, size=1, linetype=2)+
  theme_bw()+
  labs(x="Mean group Intercept", y="Mean group slope")+
  theme(axis.title = element_text(size=18), 
        axis.text = element_text(size=16))
plot_grid(pop.ef.m2.brms, m2.brms.int.schools,m2.brms.slopes.schools,
          CrInt95.int.m2.brms,CrInt95.slopes.m2.brms,cor.sl.int.m2.brms,
          nrow=2,labels = LETTERS[1:6], align="hv",label_x=0.85, label_y=0.95) #Fig. 13.19
```
##### plot to pdf, does not yet work
```{r eval = FALSE}
pdf(file.path(PPath,"Results/test2.pdf"))
pop.ef.m2.brms
dev.off()
```

