##### Von Interesse: ppcCand1W0_out und ppcCand1W1_out, Modell 2 ist die
##### Messlatte für Modell 1, es soll die Outcome-Region von Modell 1 voerhersagen.
##### 

---
title: "Students with attendance problems"
author: "Mariana Nold"
output: html_document
---
### Load R packages
```{r  eval = FALSE}
source("RFiles/packages.r")
```

### Load  single imputed PISA data.
```{r  eval = FALSE}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task_Truancy"
load(file = file.path(PPath,"Orgdata/IMPDAT_PISA.RData"))
```

### Prepare data, thus build PISA scales, 6802 students 
```{r  eval = FALSE}
source(file.path(PPath,"RFiles/01data_preperation.r"))
to.rm <- objects()[!objects() %in% c("PPath","pisa18","countries")]
rm(list = to.rm)
 #head(pisa18)
# dim(pisa18)
```


### There should be at least ten students per school. Schools with fewer observations are excluded from the analysis. The variable nb.stud informs about the number of stundets per school and save final data set, 6489 students

```{r  eval = FALSE}
tab.id <- table(pisa18$school.id)
id.frame <- as.data.frame(tab.id)

nb.stud <- rep(id.frame$Freq,id.frame$Freq)
pisa18$nb.stud <- nb.stud
rm(tab.id,id.frame,nb.stud)
pisa.data <- subset(pisa18, pisa18$nb.stud > 9)
pisa18 <- pisa.data

rm(pisa.data)
summary(pisa18$nb.stud)
save(pisa18,file = file.path(PPath,"Files/pisa18.RData")) #file = file.path(PPath,"File/fits0.RData")
#load(file.path(PPath,"Files/pisa18.RData"))
```



### Data sets for schools: Data set contains one row for each school and only school level variables
```{r  eval = FALSE}
sc.level <- names(pisa18) == toupper(names(pisa18))
sc.level[1] <- TRUE
pisa18_sc <- pisa18[,sc.level]
dim(pisa18_sc)
pisa18_sc <- pisa18_sc[!duplicated(pisa18_sc$school.id), ]
dim(pisa18_sc)
names(pisa18_sc)

save(pisa18,file = file.path(PPath,"Files/pisa18_sc.RData")) 
load(file.path(PPath,"Files/pisa18_sc.RData"))
```

### Stack models: Use all relevant predictors, hold out some to check for predictive validity


```{r  eval = FALSE}


# models to stack
# theoretical model
stack1 <- as.formula(attp ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + (1 | school.id))
# composition effect
stack2 <- as.formula(attp ~  immig + native + female + escs  + RUR + DISH 
                     + GYM + STAFFSHORT + EDUSHORT + FEMFRAC + SN_SC + NN_SC + (1 | school.id))

models <- vector(mode = "list", length = 2)

models[[1]] <- stack1
models[[2]] <- stack2

```
##### fit model

```{r  eval = FALSE}
a.seed <- 12345 
a.iter <-  2000 
a.chains <- 4  

#pisa18$WITHOUT  <- as.numeric(pisa18$FAILED >= 0.02) # above median

fits0 <- list()

fits0[[1]] <- stan_glmer(models[[1]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 

fits0[[2]] <- stan_glmer(models[[2]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 

fits1 <- list()

fits1[[1]] <- stan_glmer(models[[1]], data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 

fits1[[2]] <- stan_glmer(models[[2]], data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 


save(fits0, file = file.path(PPath,"Files/fits0.RData"))
save(fits1, file = file.path(PPath,"Files/fits1.RData"))
```


###load fits

```{r  eval = FALSE}
load(file.path(PPath, "Files/fits0.RData"))
load(file.path(PPath, "Files/fits1.RData"))
```


##### Study model parameters

```{r  eval = FALSE}
n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par0.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits0[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par0.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits0[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)


n01.models <- length(models)
sink(file = file.path(PPath, "Results/fixed_par1.txt"))
for (i in 1:n01.models) {
  print(kable(tidy(fits1[[i]],conf.int  = T, conf.level = .95, 
                   effects = "fixed"),"simple",digits = 4, 
              caption = paste("model ",i," *****")))
}
sink(file = NULL)

sink(file = file.path(PPath, "Results/ran_par1.txt"))
for (i in 1:n01.models){
  print(kable(tidy(fits1[[i]],conf.int=T, conf.level=.95, effects = "ran_pars"),"simple",digits=4, caption = paste("model ",i," *****")))
  
  }
sink(file = NULL)

```
### Load functions
```{r  eval = FALSE}
source(file.path(PPath, "RFiles/02functions.r"))
```
### Compute Predictions and Brier Score
```{r  eval = FALSE}
#n01.models <- length(models)-1
ypredCand0 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand0[[i]] <- posterior_predict(fits0[[i]])
}
save("ypredCand0", file = file.path(PPath, "Files/ypredCand0.RData"))


br.candL20 <- list()
for (i in 1:n01.models) {
  colMeansM <- colMeans(ypredCand0[[i]])
  br.candL20[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 0],colMeansM)
 }

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores0.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL20[[i]])
  
}
sink(file = NULL)


ypredCand1 <- list()
for (i in 1:n01.models) {
  set.seed(a.seed)
  ypredCand1[[i]] <- posterior_predict(fits1[[i]])
}
save("ypredCand1", file = file.path(PPath, "Files/ypredCand1.RData"))

## Brier Score 

br.candL21 <- list()
for (i in 1:n01.models) {
  br.candL21[[i]] <- brierS(pisa18$attp[pisa18$WITHOUT == 1],colMeans(ypredCand1[[i]]))
 
}

# Write Brier score to file.
sink(file = file.path(PPath, "Results/brier.scores1.txt"))

for (i in 1:n01.models) {
  print(paste("model ", i))
  print(br.candL21[[i]])
  
}
sink(file = NULL)
```

### Posterior predictive checks, predictors of candidate model 1
```{r  eval = FALSE}
## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand20_out <- visualPPC1(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_out.pdf"))
ppcCand20_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand21_out <- visualPPC1(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_out.pdf"))
ppcCand21_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand10_in <- visualPPC1(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_in.pdf"))
ppcCand10_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand11_in <- visualPPC1(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_in.pdf"))
ppcCand11_in$ppc.plots
dev.off()

num <- length(ppcCand20$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand2W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand1W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


```
#### Posterior predictive checks, predictors of candidate model 2, this is of focal #### interest
```{r}

## hold out ppc
color_scheme_set(scheme = "blue")
ppcCand10_out <- visualPPC2(yrepM = ypredCand0[[1]])
pdf(file.path(PPath, "Results/ppcCand1W0_out.pdf"))
ppcCand10_out$ppc.plots
dev.off()

color_scheme_set(scheme = "green")
ppcCand11_out <- visualPPC2(yrepM = ypredCand1[[1]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand1W1_out.pdf"))
ppcCand11_out$ppc.plots
dev.off()

color_scheme_set(scheme = "yellow")
ppcCand20_in <- visualPPC2(yrepM = ypredCand0[[2]])
pdf(file.path(PPath, "Results/ppcCand2W0_in.pdf"))
ppcCand20_in$ppc.plots
dev.off()

color_scheme_set(scheme = "red")
ppcCand21_in <- visualPPC2(yrepM = ypredCand1[[2]], WITHOUT = 1)
pdf(file.path(PPath, "Results/ppcCand2W1_in.pdf"))
ppcCand21_in$ppc.plots
dev.off()

num <- length(ppcCand21_in$ppc.plots)

sink(file = file.path(PPath,"Results/ppcCand1W0_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand10_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand1W1_out.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand11_out$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

sink(file = file.path(PPath,"Results/ppcCand2W0_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand20_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)


sink(file = file.path(PPath,"Results/ppcCand2W1_in.txt"))
for(i in 1:num){
  print(paste("Table   ",i))
  print(kable(ppcCand21_in$pppvals[i], format = "simple", digits = 2))
  print("                 ")
}
sink(file=NULL)

```

#### improve theoretical model
```{r  eval = FALSE}


# theoretical model, immig
imp1 <- as.formula(immig ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + (1 | school.id))
# theoretical model, native
imp2 <- as.formula(native ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + (1 | school.id))

# theoretical model, native
imp3 <- as.formula(NN_SC ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + (1 | school.id))

models.imp <- vector(mode = "list", length = 3)

models.imp[[1]] <- imp1
models.imp[[2]] <- imp2
models.imp[[3]] <- imp3

```
##### Fit new models

```{r  eval = FALSE}
fits.imp0 <- vector(mode = "list", length = 3)
a.seed <- 12345 
a.iter <-  2000 
a.chains <- 4  

fits.imp0[[1]] <- stan_glmer(models.imp[[1]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 

fits.imp0[[2]] <- stan_glmer(models.imp[[2]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial()) 

fits.imp0[[3]] <- stan_glmer(models.imp[[3]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,
                      iter = a.iter, chains = a.chains, family = binomial())


save(fits.imp0, file = file.path(PPath,"Files/fits.imp0.RData"))


#load(file.path(PPath,"Files/fits.imp0.RData"))

```
#### ld and fewbooks are very important, moreover aca, bul and uni
```{r}
# theoretical model, immig
imp_post1 <- as.formula(attp ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + ld:fewbooks + uni:aca + goal:ld +(1 + fewbooks| school.id))

imp_post2 <- as.formula(attp ~ aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + ld:uni + fewbooks:aca + belong:fewbooks +(1 + ld| school.id))

knots <- quantile(pisa18$goal, p = c(0.25, 0.5, 0.75))
imp_post3 <- as.formula(attp ~ aca + ld + uni  + bull + fewbooks + joyread +  mot  +  bs(goal, knots = knots) + (1 | school.id))

models.imp_post <- vector(mode = "list", length = 2)

models.imp_post[[1]] <- imp_post1
models.imp_post[[2]] <- imp_post2
models.imp_post[[3]] <- imp_post3
```
##### fit improved models
```{r}
fits.imp0_post <- vector(mode = "list", length = 2)
a.seed <- 12345 
a.iter <-  2000 
a.chains <- 4  

fits.imp0_post[[1]] <- stan_glmer(models.imp_post[[1]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed, iter = a.iter, chains = a.chains, family = binomial()) 

fits.imp0_post[[2]] <- stan_glmer(models.imp_post[[2]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,iter = a.iter, chains = a.chains, family = binomial()) 

fits.imp0_post[[3]] <- stan_glmer(models.imp_post[[3]], data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,iter = a.iter, chains = a.chains, family = binomial())


save(fits.imp0_post, file = file.path(PPath,"Files/fits.imp0_post.RData"))
```


##### Stacking, only if continous model extension is not possible
##### https://jrnold.github.io/bayesian_notes/model-comparison.html
```{r  eval = FALSE}
n01.models <- length(fits.imp0_post)
# Compute loo

loo_list0 <- lapply(fits.imp0_post, loo, cores = 1)
loo_compare(loo_list0)

save("loo_list0", file = file.path(PPath, "Files","loo_list0.RData"))
#load(file.path(PPath, "Files","loo_list0.RData"))

sink(file = file.path(PPath, "Files","loo_info0.txt"))
for (i in 1:n01.models) {
  print(paste("******** "))
  print(paste("model ", i))
  print(loo_list0[[i]])
  print(paste("******** "))
}
#print(loo_list[[1]])
sink(file = NULL)

wtsStacking0 <- list()

wtsStacking0 <- loo_model_weights(loo_list0, method = "stacking")
print(wtsStacking0)


sink(file = file.path(PPath, "Results","wtsStacking0.txt"))
print(wtsStacking0)
print(paste("******** "))
sink(file = NULL)

```
#### PPC stack for stack, indicates that the information is missing, thus 
#### the theoretical model cannot prdict the outcom region of model 2
```{r  eval = FALSE}
set.seed(a.seed)
ypredStack0 <- ypred_Stacking(fits.imp0_post,loo_list0, method = "stacking") 
save("ypredStack0", file = file.path("Files/ypredStack0.RData"))

color_scheme_set(scheme = "pink")
ppcStack0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypredStack0)
pdf(file.path(PPath, "Results/ppcStack0.pdf"))
ppcStack0
dev.off()

```
#### final model
```{r}
final <- as.formula(attp ~ native + immig + aca + ld + uni  + bull + fewbooks + belong +  joyread + goal + mot  + DISCIPLINE + TEACHBAD + VAL + NN_SC  + (1| school.id))

final.fit0 <- stan_glmer(final, data = pisa18[pisa18$WITHOUT == 0,], seed = a.seed,iter = a.iter, chains = a.chains, family = binomial())

save(final.fit0, file = file.path(PPath,"Files/final.fit0.RData"))

final.fit1 <- stan_glmer(final, data = pisa18[pisa18$WITHOUT == 1,], seed = a.seed,iter = a.iter, chains = a.chains, family = binomial())

save(final.fit1, file = file.path(PPath,"Files/final.fit1.RData"))
```
##### PPCs again
```{r}
set.seed(a.seed)
ypredfinal0 <- posterior_predict(final.fit0) 
save("ypredfinal0", file = file.path("Files/ypredfinal0.RData"))

color_scheme_set(scheme = "purple")
ypredfinal0 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypredfinal0)
pdf(file.path(PPath, "Results/ypredfinal0.pdf"))
ypredfinal0
dev.off()

set.seed(a.seed)
ypredfinal1 <- posterior_predict(final.fit1) 
save("ypredfinal1", file = file.path("Files/ypredfinal1.RData"))

color_scheme_set(scheme = "darkgray")
ypredfinal1 <- visualPPC2(y.obs = pisa18$attp, yrepM = ypredfinal1, WITHOUT = 1)
pdf(file.path(PPath, "Results/ypredfinal1.pdf"))
ypredfinal1
dev.off()
```


### Compute results, WITHOUT == 0
```{r  eval = FALSE}

pisa18_W0 <- pisa18[pisa18$WITHOUT == 0,]

pisa18_W0$DP_QU <- as.factor(cut_number(pisa18_W0$goal,4))
levels_DP_W0 <- levels(pisa18_W0$DP_QU)
ds_DP_W0 <- pisa18_W0 %>% group_by(DP_QU) %>% summarise(median = median(goal), Freq = n())
levels(pisa18_W0$DP_QU) <- 1:4

sgr <- split(pisa18_W0 ,f = list(pisa18_W0$DP_QU,pisa18_W0$bull))
t<-lapply(sgr, function(x) length(x[[1]]))
pint0 <- comp.int.stack.core01(sgr = sgr,fits = fits0,
                                     wts = wtsStacking0,
                                     cnt.path = file.path(PPath,"Files"), p = 0.5)
```
### Summarize results DH_QU
```{r  eval = FALSE}
dfDH_QU_W0 <- data.frame(x = rep(round(ds_DP_W0$median,2),2),
                        group = c(rep("no",4),rep("yes",4)),
                        mean =  pint0[,2],low = pint0[,1], 
                        up = pint0[,3])
sink(file = file.path(PPath,"Results/dfDH_QU_W0.txt"))
dfDH_QU_W0
sink(file = NULL)
```

### Compute results, WITHOUT == 1
```{r  eval = FALSE}

pisa18_W1 <- pisa18[pisa18$WITHOUT == 1,]

pisa18_W1$DP_QU <- as.factor(cut_number(pisa18_W1$goal,4))
levels_DP_W1 <- levels(pisa18_W1$DP_QU)
ds_DP_W1 <- pisa18_W1 %>% group_by(DP_QU) %>% summarise(median = median(goal), Freq = n())
levels(pisa18_W1$DP_QU) <- 1:4

sgr1 <- split(pisa18_W1 ,f = list(pisa18_W1$DP_QU,pisa18_W1$bull))
t1<-lapply(sgr1, function(x) length(x[[1]]))
pint1 <- comp.int.stack.core01(sgr = sgr1,fits = fits1,
                                     wts = wtsStacking1,
                                     cnt.path = file.path(PPath,"Files"), p = 0.5)
```
### Summarize results DH_QU
```{r  eval = FALSE}
dfDH_QU_W1 <- data.frame(x = rep(round(ds_DP_W1$median,2),2),
                        group = c(rep("no",4),rep("yes",4)),
                        mean =  pint1[,2],low = pint1[,1], 
                        up = pint1[,3])
sink(file = file.path(PPath,"Results/dfDH_QU_W1.txt"))
dfDH_QU_W1
sink(file = NULL)
```

### Plot results
```{r  eval = FALSE}

pdf(file.path(PPath,"Results/dfDH_QU_W0.pdf"))
plot.result(dfDH_QU_W0)
dev.off()

pdf(file.path(PPath,"Results/dfDH_QU_W1.pdf"))
plot.result(dfDH_QU_W1)
dev.off()
`
