---
output:
  html_document: default
  pdf_document:  default
  word_document: default
  html_notebook: default
  
---
---
title: "Exploring School Belonging Through the Lens of the Participation-Identification Model"
author: "Mariana Nold" 
date: "June 12, 2025"
---
##### Load R packages 
```{r load package, message = FALSE, warning = FALSE}
source("RFiles/packages.r")
```

##### Load single imputed PISA data and define Project Path
```{r  read path and load data}
PPath <- "C:/Users/zo95yup/Documents/GitHub/Bayes_for_STRATOS/Task_Truancy"
load(file = file.path(PPath,"Orgdata/IMPDAT_PISA.RData"))
```

##### Prepare data, thus build PISA scales 
```{r prepare data, message = FALSE, warning = FALSE}
# Pole so that high values indicate a higher sense of belonging based on PIM
source(file.path(PPath,"RFiles/01_datapreperation.r"))
to.rm <- objects()[!objects() %in% c("PPath","pisa18")]
rm(list = to.rm)
# head(pisa18)
# dim(pisa18)
```


##### There should be at least ten students per school. Schools with fewer observations are excluded from the analysis. 

```{r  exclude small schools}
pisa18$school.id <- as.factor(pisa18$school.id)
help <- split(pisa18,pisa18$school.id)
nst <- lapply(help, function(x) length(x[[1]]))
tf <- lapply(nst, function(x) x[[1]] > 9)
utf <- unlist(tf)
large_school <- help[utf]
pisa2018 <- list.rbind(large_school)
pisa2018$school.id <- droplevels(pisa2018$school.id)
#table(pisa2018$school.id)

rm(pisa18,tf,nst,help,large_school,utf)

save(pisa2018,file = file.path(PPath,"Files/pisa2018.RData")) 
```


##### Prepare focal variables

```{r  define second focal variable and outcome}
# Compute quartiles of being bullied at school level
pisa2018$ATT4 <- cut_number(pisa2018$ATT01,4)
levels(pisa2018$ATT4)
levels(pisa2018$ATT4) <- 1:4
threshold <- quantile(pisa2018$belong, 0.25)
threshold
# Focal variable is 1, if student has a low sense of school-belonging
pisa2018$bel01 <- ifelse(pisa2018$belong <= threshold, 1, 0)

table(pisa2018$ATT4)
table(pisa2018$bel01)

```

##### Define model two models, first full model, second background model

```{r  fit model,results='hide', message=FALSE, warning=FALSE}
models<-list()
# Full model, variable selection based on PIM  including focal variables
models[[1]] <- as.formula(bel01 ~ female + nld +  aca  + val  + nfof  + native   +  joyread + goal + mot + res +  ndis_clim    + GYM + UNI  +  NDISH + N_SC + bull + ATT4 + bull:ATT4 +(1 | school.id))
# Background model, variable selection based on PIM without focal variables
models[[2]] <- as.formula(bel01 ~ female + nld +  aca  + val  + nfof  + native   +  joyread + goal + mot + res +  ndis_clim    + GYM + UNI  + NDISH + N_SC + (1 | school.id))
```
##### Check the convergence criteria (Step 5 in SAP)



```{r fit modesl and do MCMC diagnostic,results='hide', message = FALSE, warning = FALSE, cache=TRUE}
a.seed <- 12345 
a.iter <-  2000 
a.chains <- 4  

fits <- list()
# full model
fits[[1]] <- stan_glmer(models[[1]], data = pisa2018, seed = a.seed,
                      iter = a.iter, chains = a.chains,family = binomial()) 
# background model
fits[[2]] <- stan_glmer(models[[2]], data = pisa2018, seed = a.seed,
                      iter = a.iter, chains = a.chains,family = binomial()) 

res.path <- file.path(PPath,"Result")
save(fits, file = file.path(res.path,"fits.RData"))
summary_list <- lapply(fits, summary) # mcse Rhat n_eff
# launch_shinystan(fits[[1]])
# launch_shinystan(fits[[2]])
```
#### Summary of results
```{r Summary of results}
library(broom)  
library(purrr)  

# Calculate summaries for all models in fits
summaryFixed <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "fixed"))
summarySchools <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "ran_vals"))
summaryVar <- map(fits, ~ tidy(.x, conf.int = TRUE, conf.level = .95, effects = "ran_pars"))

# Outputting results
walk(summaryFixed, ~ print(.x, digits = 2, n = 28))
walk(summarySchools, ~ print(.x, digits = 2))
walk(summaryVar, ~ print(.x, digits = 2))

```

##### LOOIC
```{r looic, cache=TRUE}
loo_list <- list()
# Compute loo

loo_list <- lapply(fits, loo, cores = 1)
# Save loo

save("loo_list", file = file.path(res.path,"loo_list.RData"))

print(loo_compare(loo_list[[1]], loo_list[[2]]), digits = 2)


```
##### Randomized Quantile Residuals
```{r Randomized Quantile Residuals}
source(file.path(PPath,"RFiles/02_functions_KD.R"))
# Compute posterior predictions
posterior_predictions <- lapply(fits, posterior_predict)
library("DHARMa")
qres.m1=createDHARMa(simulatedResponse = t(posterior_predictions[[1]]), 
                          observedResponse = pisa2018$bel01, 
                          fittedPredictedResponse = apply(posterior_predictions[[1]], 2, median), integerResponse = T, seed = 123, method = "PIT") 
summary(residuals(qres.m1))
# QQ-Plot 
qqnorm(residuals(qres.m1))
qqline(residuals(qres.m1), col = "red")

plots1 <- plot_violin_residuals(qres.m1, pisa2018, "Residual Diagnostics Full model")
print(plots1)

qres.m2=createDHARMa(simulatedResponse = t(posterior_predictions[[2]]), 
                          observedResponse = pisa2018$bel01, 
                          fittedPredictedResponse = apply(posterior_predictions[[2]], 2, function(x) ifelse(median(x) == 0.5, 1, median(x))), integerResponse = T, seed = 123, method = "PIT") 
summary(residuals(qres.m2))

# QQ-Plot 
qqnorm(residuals(qres.m2))
qqline(residuals(qres.m2), col = "red")

plots2 <- plot_violin_residuals(qres.m2, pisa2018, "Residual Diagnostics Background model")
print(plots2)
```

##### PPN based on Kolmogorov distance
```{r KDs}

# Compute Kolmogorov distance
KD(ypredCore = posterior_predictions[[1]], yrep = posterior_predictions[[2]])

```

##### Posterior predictive checks
```{r ppcs}
color_scheme_set(scheme = "pink")
ppcModelFull <- visualPPC(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[1]])
#pdf(file.path(cnt.path,"ppcCand1.pdf"))
ppcModelFull
#dev.off()

color_scheme_set(scheme = "blue")
#color_scheme_set(scheme = "gray")
ppcModelBG <- visualPPC(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[2]])
#pdf(file.path(cnt.path,"ppcCand2.pdf"))
ppcModelBG
#dev.off()


```

##### Compute Bayesian p-values
```{r Bayesian p value}
# https://search.r-project.org/CRAN/refmans/HuraultMisc/html/post_pred_pval.html
#library("kableExtra")

num <- 3

pppvModelFull <- two.sided.pppv(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[1]], alternative = "two.sided")
#sink(file = file.path(cnt.path,"pppv_Cand1.txt"))
for(i in 1:num){
  print(knitr::kable(matrix(pppvModelFull[[i]],nrow=2,byrow=TRUE), format = "simple", digits = 3))
}
#sink(file=NULL)

pppvModelBG <- two.sided.pppv(y.obs = pisa2018$bel01, yrepM = posterior_predictions[[2]], alternative = "two.sided")
#sink(file = file.path(cnt.path,"pppv_Cand2.txt"))
for(i in 1:num){
    print(knitr::kable(matrix(pppvModelBG[[i]],nrow=2,byrow=TRUE), format = "simple", digits = 3))
}
#sink(file=NULL)


```
##### Compute Brier scores
```{r Brier scores of cands}
br <- list()
for (i in 1:2) {
  br[[i]] <- brierS(pisa2018$bel01,colMeans(posterior_predictions[[i]]))
 
}

#sink(file = file.path(cnt.path,"brier.scores.cand.txt"))
for (i in 1:2) {
  print(paste("model ", i))
  print(br[[i]])
  }
#sink(file = NULL)
```
##### Summarizing results based on the posterior distribution of full model
```{r Summarizing results}
library(dplyr)

# For each quartile compute the median ATT01.
pisa18condPr1 <- pisa2018 %>%
  group_by(ATT4) %>%
  summarise(mATT01 = median(ATT01), sd = sd(ATT01),min = min(ATT01),max = max(ATT01))

meancondPr1 <- pisa18condPr1$mATT01

meancondPr1

squcondPr <- split(pisa2018,f = list(pisa2018$ATT4,pisa2018$bull))


pint_condPr25 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.25)

pint_condPr50 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.5)

pint_condPr75 <- drawPosteriorInequalityGroup(sgr = squcondPr,fit = fits[[1]], p = 0.75)


brE <- 4

df_condPr25 <- data.frame(x = rep(round(meancondPr1,2),2),
                          group = c(rep("nbull.",brE),rep("bull.",brE)),
                          mean =  pint_condPr25[,2],low = pint_condPr25[,1], 
                          up = pint_condPr25[,3])

df_condPr50 <- data.frame(x = rep(round(meancondPr1,2),2),
                        group = c(rep("nbull.",brE),rep("bull.",brE)),
                        mean =  pint_condPr50[,2],low = pint_condPr50[,1], 
                        up = pint_condPr50[,3])

df_condPr75 <- data.frame(x = rep(round(meancondPr1,2),2),
                          group = c(rep("nbull.",brE),rep("bull.",brE)),
                          mean =  pint_condPr75[,2],low = pint_condPr75[,1], 
                          up = pint_condPr75[,3])

lowQ <- c(pint_condPr25[,1],pint_condPr50[,1],pint_condPr75[,1])
med <-  c(pint_condPr25[,2],pint_condPr50[,2],pint_condPr75[,2])
upQ <-  c(pint_condPr25[,3],pint_condPr50[,3],pint_condPr75[,3])

df_condPr1 <- data.frame(x = rep(round(meancondPr1,2),6),
                            group = c(rep("disad. nbull.",brE),rep("disad. bull.",brE),
                                  rep("aver. nbull.",brE),rep("aver bull.",brE),
                                  rep("adv. nbull.",brE),rep("adv. bull.",brE)),
                                  mean =  med,low = lowQ, 
                                   up = upQ)

#sink(file = file.path(cnt.path,"interval.txt"))
 # kable(df_condPr1,format = "latex", digits = 3)
knitr::kable(df_condPr1, format = "simple", digits = 3)
#sink(file = NULL)


p_condPr <- plot.result(df_condPr1)

#pdf(file.path(cnt.path,"plot_condPr.pdf"))
p_condPr
#dev.off()

```