---
title: "Bacteremia data Bayesian reanalysis"
author: 
  - Aliaksandr Hubin
date: now
date-format: iso
title-block-banner: "#828282"
format:
  html:
    code-fold: true
    code-tools: true
    theme: default
    toc: true
    number-sections: true
    toc-location: right
    embed-resources: true
    self-contained: true
    df-print: kable
params: 
  folder_analysis: Data/Processed/Bacteremia_SAP
  set: 4K
---

## **Overview**

This analysis evaluates a binary classification model for predicting BloodCulture (bacteremia presence) using the FBMS package. We use a bootstrap approach to estimate model performance, including **C-index,** and **calibration slope**. The bootstrap loop is parallelized using mclapply to improve efficiency. We compute the .632+ bootstrap estimates and 95% confidence intervals for each metric as well as the relative over-fitting rate.

The dataset is sourced from [Zenodo](https://zenodo.org/records/7554815/files/Bacteremia_public_S2.csv). The analysis includes data preprocessing, model fitting, performance evaluation, and bootstrap resampling.

**Step 1: Load Libraries and Set Seed**

We load required packages: Metrics for C-index, FBMS for model fitting, and parallel for parallelized bootstrapping. The random seed is set for reproducibility.

```{r}
# Install FBMS if needed
# devtools::install_github("https://github.com/jonlachmann/GMJMCMC/tree/stratos")

library(Metrics)  # For C-index
library(FBMS)     # For model fitting
library(parallel) # For mclapply
set.seed(123)     # For reproducibility
read.results <- T # If we should read saved results here
```

## **Define Performance Metric Functions**

We use the following metrics:

-   **C-index**;

-   **Calibration Slope**;

-   **Calibration Intercept**.

**Step 3: Load and Preprocess Data**

We load the bacteremia dataset, remove unnecessary columns, and convert the BloodCulture variable to a binary (0/1) format.

```{r}
# Load dataset
df <- read.csv("https://zenodo.org/records/7554815/files/Bacteremia_public_S2.csv", 
               header = TRUE, sep = ",", dec = ".")

# Remove unnecessary columns
df <- df[, !(names(df) %in% c("ID", "MONOR", "LYMR", "NEUR", "EOSR", "BASOR"))]

# Convert BloodCulture to binary
df$BloodCulture <- ifelse(df$BloodCulture == "yes", 1, 0)


df.train <- df[1:4000,]
df.valid<- df[-c(1:4000),]

```

## **Fit Full Models**

We fit two models using the FBMS package:

-   A **nonlinear model** using gmjmcmc.parallel with transformations (sigmoid, sin, cos, exp_dbl).

Predictions are made on the training set (optimistic) using the nonlinear model, and we compute C-index, and calibration slope.

```{r, echo=FALSE, results='hide'}
# Fit nonlinear model

params.gmj <- gen.params.gmjmcmc(94)

params.gmj$feat$pop.max <- 150

result.nonlinear <- fbms(formula = BloodCulture ~ 1 + ., family = "binomial", beta_prior =   list(type = "Jeffreys-BIC"), params = params.gmj,
data = df.train, impute = TRUE, method = "gmjmcmc.parallel", 
P = 10, cores = 6, runs = 6, 
transforms = c("sigmoid", "sin", "cos", "exp_dbl"))

```

get summary

```{r}
summary(result.nonlinear)

```

and predictions

```{r}


# Predictions on test data (using nonlinear model)
prob_full <- predict(result.nonlinear, df.train[, -48], link = sigmoid)$aggr$mean
prob_full_test <- predict(result.nonlinear, df.valid[, -48], link = sigmoid)$aggr$mean

# Metrics on full model
cindex_full <- rms::val.prob(prob_full, df.train$BloodCulture, pl = FALSE)["C (ROC)"]
calib_slope_full <- rms::val.prob(prob_full, df.train$BloodCulture, pl = FALSE)["Slope"]
calib_inter_full <- rms::val.prob(prob_full, df.train$BloodCulture, pl = FALSE)["Intercept"]
calib_brier_full <- rms::val.prob(prob_full, df.train$BloodCulture, pl = FALSE)["Brier"]



# Metrics on full model
cindex_full_test <- rms::val.prob(prob_full_test, df.valid$BloodCulture, pl = FALSE)["C (ROC)"]
calib_slope_full_test <- rms::val.prob(prob_full_test, df.valid$BloodCulture, pl = FALSE)["Slope"]
calib_inter_full_test <- rms::val.prob(prob_full_test, df.valid$BloodCulture, pl = FALSE)["Intercept"]
calib_brier_full_test <- rms::val.prob(prob_full_test, df.valid$BloodCulture, pl = FALSE)["Brier"]


cat("train cindex_full: ", cindex_full," train calib_slope_full: ",calib_slope_full," train calib_intercept_full: ",calib_inter_full," train calib_brier_full: ",calib_brier_full)
cat("test cindex_full: ", cindex_full_test," test calib_slope_full: ",calib_slope_full_test," test calib_intercept_full: ",calib_inter_full_test," test calib_brier_full: ",calib_brier_full_test)
```

Only linear terms are present above, hence we continue with linear models as discussed in the reanalysis plan.

## **Define Theoretical Performance**

We define baseline performance for a model with no predictive power (absence of effect):

-   **C-index (AUC)**: 0.5 (no discrimination).

-   **Calibration Slope**: 0.

-   **Calibration Intercept**: intercept only model.

```{r}
# Theoretical performance (no effect)
p_M0_cindex <- 0.5  # For C-index (equivalent to AUC)
p_M0_calib_slope <- 0.0  # For calibration slope (perfect calibration)
p_M0_calib_intercept <- glm(BloodCulture ~ 1, data = df.train, family = "binomial")$coefficients[1] #For calibration intercept
```

## **Bootstrap Procedure with Parallelization**

We perform n_bootstrap = 1000 bootstrap iterations to estimate model performance. Each iteration:

-   Resamples the training data with replacement.

-   Fits a model using mjmcmc.

-   Computes C-index, calibration intercept, and calibration slope for in-sample (bootstrap) and out-of-bag (test) predictions.

The loop is parallelized using mclapply with multiple cores to improve computational efficiency.

```{r}
# Number of bootstrap iterations
n_bootstrap <- 1000
# Define bootstrap iteration function
bootstrap_iter <- function(i) {
  # Bootstrap resample from training data
  boot_indices <- sample(1:nrow(df.train), replace = TRUE)
  df_boot <- df.train[boot_indices, ]
  df.test <- df.train[-boot_indices, ]
  # Fit model on bootstrap sample
  result_boot <- fbms(formula = BloodCulture ~ 1 + ., family = "binomial", beta_prior =   list(type = "Jeffreys-BIC"),
  data = df_boot, impute = TRUE, method = "mjmcmc", N = 1000)
  
  # Predictions on bootstrap sample (in-sample performance)
  prob_boot <- predict(result_boot, df_boot[, -48], link = sigmoid)$mean
  #prob_boot <- sigmoid(preds_boot$mean)
  
  # Predictions on test data (out-of-bag performance)
  prob_oob <- predict(result_boot, df.test[, -48], link = sigmoid)$mean
  #prob_oob <- sigmoid(preds_oob$mean)
  
  # Calculate metrics
  
  #boot
  cindex_boot <- rms::val.prob(prob_boot, df_boot$BloodCulture, pl = FALSE)["C (ROC)"]
  calib_slope_boot <- rms::val.prob(prob_boot, df_boot$BloodCulture, pl = FALSE)["Slope"]
  calib_inter_boot <- rms::val.prob(prob_boot, df_boot$BloodCulture, pl = FALSE)["Intercept"]
  calib_brier_boot <- rms::val.prob(prob_boot, df_boot$BloodCulture, pl = FALSE)["Brier"]
  
  #oob
  cindex_oob <- rms::val.prob(prob_oob, df.test$BloodCulture, pl = FALSE)["C (ROC)"]
  calib_slope_oob <- rms::val.prob(prob_oob, df.test$BloodCulture, pl = FALSE)["Slope"]
  calib_inter_oob <- rms::val.prob(prob_oob, df.test$BloodCulture, pl = FALSE)["Intercept"]
  calib_brier_oob <- rms::val.prob(prob_oob, df.test$BloodCulture, pl = FALSE)["Brier"]


  
  return(c(cindex_boot = cindex_boot, cindex_oob = cindex_oob, calib_inter_boot = calib_inter_boot, calib_inter_oob = calib_inter_oob,
           calib_slope_boot = calib_slope_boot, calib_slope_oob = calib_slope_oob,
           calib_brier_boot = calib_brier_boot, calib_brier_oob = calib_brier_oob))
}

```

bootstrap results or read them in

```{r}

# Run bootstrap iterations in parallel with 9 cores
if(!read.results)
{
  results <- mclapply(1:n_bootstrap, bootstrap_iter, mc.cores = 5)
  results_matrix <- do.call(rbind, results)
  saveRDS(object = results,file = "resultmat1.rds")
}else
{
  results_matrix <- readRDS("resultmat.rds")
}
```

proceed with analyzing the output of bootstrap

```{r}

# Extract metrics
cindex_boot <- results_matrix[, "cindex_boot.C (ROC)"]
cindex_oob <- results_matrix[, "cindex_oob.C (ROC)"]
calib_slope_boot <- results_matrix[, "calib_slope_boot.Slope"]
calib_slope_oob <- results_matrix[, "calib_slope_oob.Slope"]
calib_brier_boot <- results_matrix[, "calib_brier_boot.Brier"]
calib_brier_oob <- results_matrix[, "calib_brier_oob.Brier"]
calib_inter_boot <- results_matrix[, "calib_inter_boot.Intercept"]
calib_inter_oob <- results_matrix[, "calib_inter_oob.Intercept"]
```

## **Calculate Overfitting Rates and .632+ Estimates**

We compute the **relative overfitting rate (R)** for each metric to quantify the degree of overfitting. Using these, we calculate the **.632+ bootstrap estimates**, which combine full model performance with bootstrap performance to correct for bias.

```{r}
# Calculate overfitting rate R
R_cindex <- mean(cindex_oob - cindex_full) / (p_M0_cindex - cindex_full)

R_calib_slope <- mean(calib_slope_oob - calib_slope_full) / (p_M0_calib_slope - calib_slope_full)

# .632+ weights
w_cindex <- 0.632 / (1 - 0.368 * R_cindex)
w_calib_slope <- 0.632 / (1 - 0.368 * R_calib_slope)

# .632+ estimates
cindex_632plus <- (1 - w_cindex) * cindex_full + w_cindex * mean(cindex_oob)
calib_slope_632plus <- (1 - w_calib_slope) * calib_slope_full + w_calib_slope * mean(calib_slope_oob)
```

**Compute Confidence Intervals**

We calculate 95% confidence intervals for each metric using the 2.5th and 97.5th percentiles of the bootstrap distributions.

```{r}
# Confidence intervals (95% CI using bootstrap percentiles)
cindex_ci <- quantile(cindex_oob, probs = c(0.025, 0.975))
calib_slope_ci <- quantile(calib_slope_oob, probs = c(0.025, 0.975))
calib_inter_ci <- quantile(calib_inter_oob, probs = c(0.025, 0.975))
calib_brier_ci <- quantile(calib_brier_oob, probs = c(0.025, 0.975))
```

## **Display Results**

Finally, we print the full model performance, .632+ estimates, and 95% confidence intervals for accuracy, C-index (AUC), and calibration slope.

```{r}
# Print results
cat("Full model Train C-index (AUC):", cindex_full, "\n")
cat("Full model Test C-index (AUC):", cindex_full_test, "\n")
cat(".632+ C-index (AUC) estimate:", cindex_632plus, "\n")
cat("C-index (AUC) 95% CI from bootstrap:", cindex_ci, "\n\n")

cat("Full model Train Calibration Slope:", calib_slope_full, "\n")
cat("Full model Test Calibration Slope:", calib_slope_full_test, "\n")
cat(".632+ Calibration Slope estimate:", calib_slope_632plus, "\n")
cat("Calibration Slope 95% CI from bootstrap:", calib_slope_ci, "\n")

cat("Full model Train Calibration Intercept:", calib_inter_full, "\n")
cat("Full model Test Calibration Intercept:", calib_inter_full_test, "\n")
#cat(".632+ Calibration Intercept estimate:", calib_inter_632plus, "\n")
cat("Calibration Intercept 95% CI from bootstrap:", calib_inter_ci, "\n")

cat("Full model Train Brier Score:", calib_brier_full, "\n")
cat("Full model Test Brier Score:", calib_brier_full_test, "\n")
cat("C-index Brier Score  95% CI from bootstrap:", calib_brier_ci, "\n\n")
```

## **Conclusion**

This analysis provides a robust evaluation of the classification model using bootstrap resampling. The .632+ estimates account for overfitting, and the calibration slope ensures the reliability of predicted probabilities. Parallelization with 9 cores improves computational efficiency. Future work could include visualizing the calibration curve or comparing linear and nonlinear models.
