---
title: "Kepler IDA"
author: 
  - Aliaksandr Hubin
date: now
date-format: iso
title-block-banner: "#828282"
format:
  html:
    code-fold: true
    code-tools: true
    theme: default
    toc: true
    number-sections: true
    toc-location: right
    embed-resources: true
    self-contained: true
    df-print: kable
editor: 
  visual
---

# Load the libraries

```{r}
library(tidyverse) 
library(DataExplorer)
```

# IDA per 02.07.2025 version of the database

Firs the data will be updated from the most recent version of the database (OpenExoplanetCatalogue)

```{r}
data = read.csv("https://raw.githubusercontent.com/OpenExoplanetCatalogue/oec_tables/master/comma_separated/open_exoplanet_catalogue.txt")

head(data)
```

The data has 5414 observations over 25 variables.

```{r}
DataExplorer::plot_str(data = data)
```

The data has 19 continuous columns and 5 categorical columns and in total 34% missing observations.

```{r}
DataExplorer::plot_intro(data = data) 
```

Item missingness is described below

```{r}
DataExplorer::plot_missing(data = data)
```

We shall keep data with less than 60% of missing values and further remove planetary system specific variables and irrelevant for the analysis ones such as last-update, list, name, and discovery method. We assume missingness at random at hence observation units with missing data in any predictor variable or the dependent variable are removed.

```{r}
data <- data %>% select(semimajoraxis,mass, radius, period, eccentricity, hoststar_mass, hoststar_radius, hoststar_metallicity,hoststar_temperature,binaryflag) %>% na.omit()
```

Univariate distributions of all predictors and the dependent variables are produced below

```{r}

DataExplorer::plot_density(data,nrow = 3,ncol = 3)
```

Same on the log scale

```{r}
DataExplorer::plot_density(log(data[,c("eccentricity","mass","period","semimajoraxis","hoststar_radius","hoststar_mass")]),nrow = 2,ncol = 3)

```

A matrix of Spearman correlations is produced below

```{r}
DataExplorer::plot_correlation(data,type = "continuous",cor_args = list(use = "pairwise.complete.obs", method = "spearman"))
```

# Splitting into the training and hold out test sets and saving

The data set was split into training and test data set. 300 randomly chosen planets are kept for evaluation.

```{r}

summary(data)

set.seed(1)
te.ind <- sample.int(n = 939,size = 300,replace = F)
data.train <- data[-te.ind,]
data.test <- data[te.ind,]

saveRDS(data.train,file = "train.RDS")
saveRDS(data.test,file = "test.RDS")
```
