---
title: "Analysis Plan for Kepler Data"
author: "Aliaksandr Hubin"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
#devtools::install_github("https://github.com/jonlachmann/GMJMCMC/tree/stratos")
library(FBMS) 
library(tidyverse) 
library(DataExplorer)
#setwd("Task_Kepler")
```

# Load the data and perform EDA

```{r}
data = read.csv("https://raw.githubusercontent.com/OpenExoplanetCatalogue/oec_tables/master/comma_separated/open_exoplanet_catalogue.txt")


head(data)
DataExplorer::plot_str(data = data)
DataExplorer::plot_intro(data = data) 
DataExplorer::plot_missing(data = data)
DataExplorer::plot_density(data[,-2],nrow = 2,ncol = 3)
DataExplorer::plot_density(log(data[,c("eccentricity","mass","period","semimajoraxis","hoststar_radius","hoststar_mass")]),nrow = 2,ncol = 3)
DataExplorer::plot_correlation(data[,-2],type = "continuous",cor_args = list(use = "pairwise.complete.obs", method = "pearson"))

```

# Select relevant columns for analysis

We shall keep 300 observations as a hold out test set and the rest as the training set

```{r}
data <- data %>% select(semimajoraxis,mass, radius, period, eccentricity, hoststar_mass, hoststar_radius, hoststar_metallicity,hoststar_temperature,binaryflag) %>% na.omit()
summary(data)

set.seed(1)
te.ind <- sample.int(n = 939,size = 300,replace = F)
data.train = data[-te.ind,]
data.test = data[te.ind,]
```

### Bayesian generalized nonlinear models

We shall use BGNLM with Jeffreys priors for inference run with 32 parallel chains for 20 populations of GMJMCMC algorithm

```{r}
transforms <- c("sin_deg","exp_dbl","p0","troot","p3")
probs <- gen.probs.gmjmcmc(transforms)
params <- gen.params.gmjmcmc(ncol(data.train)-1)
params$loglik$var = "unknown"
set.seed(1)
bgnlm <- FBMS::fbms(semimajoraxis ~ ., data = data.train,method = "gmjmcmc.parallel",beta_prior = list(type = "Jeffreys-BIC"),transforms = transforms,runs = 32,cores = 8,P = 20, probs = probs, params = params)
plot(bgnlm)
summary(bgnlm,labels = names(data.train)[-1])
```

check convergence

```{r}
diagn_plot(bgnlm,window = 10,FUN = median)
diagn_plot(bgnlm,window = 10,FUN = max)
```

we see possible convergence for the later generations of GMJMCMC, but let us increase the compute to check if it is indeed so. The limitation of the convergence statistics is that it may show perfect convergence upon algorithm stuck in a good mode and not mixing further across the modes.

```{r}
set.seed(2)
bgnlm <- FBMS::fbms(semimajoraxis ~ ., data = data.train,method = "gmjmcmc.parallel",beta_prior = list(type = "Jeffreys-BIC"),transforms = transforms,runs = 64,cores = 8,P = 50,probs = probs, params = params)
plot(bgnlm)
summary(bgnlm,labels = names(data.train)[-1])
```

```{r}
diagn_plot(bgnlm,window = 10,FUN = median)
diagn_plot(bgnlm,window = 10,FUN = max)
```

Much better indication of convergence here.\
Median probability model:

```{r}
confint(lm(semimajoraxis ~ 1 +I(troot(period*period*hoststar_mass)),data = data.train))
```

```{r}
preds.train.bgnlm <- predict(bgnlm, data.train[,-1])
preds.test.bgnlm <- predict(bgnlm, data.test[,-1])
r.bgnlm <- round(c(cor(data.train[,1],preds.train.bgnlm$aggr$mean)^2,cor(data.test[,1],preds.test.bgnlm$aggr$mean)^2),3)
plot(x = data.train[, 1], preds.train.bgnlm$aggr$mean, xlab = "data", ylab = "prediction", main = "Title of the Plot", col = "blue", pch = 16)
points(x = data.test[, 1], preds.test.bgnlm$aggr$mean, col = "red", pch = 17)
legend("topright", legend = c(paste0("Training Data, R.sqr = ",r.bgnlm[1]), paste0("Test Data, R.sqr = ",r.bgnlm[2])), col = c("blue", "red"), pch = c(16, 17))
```

\
In the Bayesian Generalized Nonlinear Model (BGNLM) analysis, you obtained the following results:

**Predictor: `troot(((period*hoststar_mass)*period))`** **Inclusion Probability: 1.000000**

This result indicates a perfect inclusion probability (1.0) for the predictor `troot(((period*hoststar_mass)*period))`, suggesting that this complex non-linear interaction term is crucial for predicting the semimajor axis.

### Interpretation

#### 1. **Complex Interaction Term**:

-   The predictor `troot(((period*hoststar_mass)*period))` combines `period` and `hoststar_mass` in a multiplicative form, followed by a transformation.

-   `troot` likely represents a specific non-linear transformation, such as a root function, which means the predictor is a transformed version of the product of `period`, `hoststar_mass`, and `period` again.

### Discussion with Relation to Kepler's Third Law

#### 1. **Kepler's Third Law**

Kepler's Third Law states that the square of the orbital period of a planet is proportional to the cube of the semimajor axis of its orbit.

#### 2. **Implications and Interpretation**

-   **Alignment with Physical Laws**: The inclusion of `troot(((period*hoststar_mass)*period))` in the BGNLM model supports Kepler’s Third Law by explicitly incorporating the cubic root transformation of the complicated interaction. This suggests that the model correctly captures the underlying (known in this case as Kepler's Third Law) physical relationship between the orbital period and the semimajor axis.

-   **Predictive Power and Physical Meaning**: The perfect inclusion probability of this term indicates that the model effectively leverages the cubic root relationship to predict the semimajor axis. This reinforces the physical validity of the model and emphasizes the importance of incorporating both the period and the host star’s mass in a non-linear fashion.

### Predictions

```{r}
preds.train.bgnlm <- predict(bgnlm, data.train[,-1])
preds.test.bgnlm <- predict(bgnlm, data.test[,-1])
r.bgnlm <- round(c(cor(data.train[,1],preds.train.bgnlm$aggr$mean)^2,cor(data.test[,1],preds.test.bgnlm$aggr$mean)^2),3)
plot(x = data.train[, 1], preds.train.bgnlm$aggr$mean, xlab = "data", ylab = "prediction", main = "Title of the Plot", col = "blue", pch = 16)
points(x = data.test[, 1], preds.test.bgnlm$aggr$mean, col = "red", pch = 17)
legend("topright", legend = c(paste0("Training Data, R.sqr = ",r.bgnlm[1]), paste0("Test Data, R.sqr = ",r.bgnlm[2])), col = c("blue", "red"), pch = c(16, 17))
```

The predictions are expectidely perfect on both training and testing sets, as inclusion of `troot(((period*hoststar_mass)*period))` in the BGNLM model effectively captures the essence of Kepler’s Third Law, incorporating the orbital period and the host star’s mass in a cubic root transformation. This predictor reflects the key physical relationship between these variables, demonstrating that the model is not only statistically sound but also physically meaningful. The high inclusion probability underscores the importance of this non-linear interaction in accurately predicting the semimajor axis.

And use the g-prior and redo the inference.

```{r}
set.seed(3)
bgnlm <- FBMS::fbms(semimajoraxis ~ ., data = data.train,method = "gmjmcmc.parallel",beta_prior = list(type = "g-prior", g = nrow(data.train)),transforms = transforms,runs = 64,cores = 8,P = 50,probs = probs, params = params)
plot(bgnlm)
summary(bgnlm,labels = names(data.train)[-1])

```

Here we perfectly recover the true law! Also good convergence for g - prior.

```{r}
diagn_plot(bgnlm,window = 10,FUN = median)
diagn_plot(bgnlm,window = 10,FUN = max)
```

And as expected perfect predictions

```{r}
preds.train.bgnlm <- predict(bgnlm, data.train[,-1])
preds.test.bgnlm <- predict(bgnlm, data.test[,-1])
r.bgnlm <- round(c(cor(data.train[,1],preds.train.bgnlm$aggr$mean)^2,cor(data.test[,1],preds.test.bgnlm$aggr$mean)^2),3)
plot(x = data.train[, 1], preds.train.bgnlm$aggr$mean, xlab = "data", ylab = "prediction", main = "Title of the Plot", col = "blue", pch = 16)
points(x = data.test[, 1], preds.test.bgnlm$aggr$mean, col = "red", pch = 17)
legend("topright", legend = c(paste0("Training Data, R.sqr = ",r.bgnlm[1]), paste0("Test Data, R.sqr = ",r.bgnlm[2])), col = c("blue", "red"), pch = c(16, 17))
```

We see that variation is possible in the results and that ideally maximal possible compute resources should be used to reduce the variation. But given enough resources GMJMCMC converges to being able to recover the true underlying physical law.
